### 2020-11-09 proposal.tex
init

\documentclass{proposal}
\usepackage{graphicx}

% alternative font if you prefer
%\usepackage{times}

% for alternative page numbering use the following package
% and see documentation for commands
%\usepackage{fancyheadings}

% other potentially useful packages
%\uspackage{amssymb,amsmath}
%\usepackage{url}
%\usepackage{fancyvrb}
%\usepackage[final]{pdfpages}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Title of project proposal placed here}
\author{Name of author placed here}
\date{Date of submission placed here}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{intro}

briefly explain the context of the project problem

\subsection{A subsection}
Please note your proposal need not follow the included section headings - this is only a suggested structure. Also add subsections etc as required

example references: \cite{BK08}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Statement of Problem}

clearly state the problem to be addressed in your forthcoming project. Explain why it would be worthwhile to solve this problem.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background Survey}

present an overview of relevant previous work including articles, books, and existing software products. Critically evaluate the strengths and weaknesses of the previous work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proposed Approach}

state how you propose to solve the software development problem. Show that your proposed approach is feasible, but identify any risks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Work Plan}

show how you plan to organize your work, identifying intermediate deliverables and dates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% it is fine to change the bibliography style if you want
\bibliographystyle{plain}
\bibliography{proposal}
\end{document}

### 2020-11-10 src/CaptureTracking.cpp
DefaultMaxUses became registered twice in LLVM11, need new header for getDefaultTargetTriple

@@ -33,11 +33,7 @@ using namespace llvm;
 /// TODO: we should probably introduce a caching CaptureTracking analysis and
 /// use it where possible. The caching version can use much higher limit or
 /// don't have this cap at all.
-static cl::opt<unsigned> DefaultMaxUsesToExplore("capture-tracking-max-uses-to-explore", cl::Hidden,
-                                                 cl::desc("Maximal number of uses to explore."),
-                                                 cl::init(20));
-
-unsigned llvm::getDefaultMaxUsesToExploreForCaptureTracking() { return DefaultMaxUsesToExplore; }
+unsigned llvm::getDefaultMaxUsesToExploreForCaptureTracking() { return 20; }
 
 CaptureTracker::~CaptureTracker() = default;
 
@@ -106,7 +102,7 @@ void llvm::PointerMayBeCaptured(const Value *V, CaptureTracker *Tracker,
                                 unsigned MaxUsesToExplore) {
   assert(V->getType()->isPointerTy() && "Capture is for pointers only!");
   if (MaxUsesToExplore == 0)
-    MaxUsesToExplore = DefaultMaxUsesToExplore;
+    MaxUsesToExplore = getDefaultMaxUsesToExploreForCaptureTracking();
 
   SmallVector<const Use *, 20> Worklist;
   Worklist.reserve(getDefaultMaxUsesToExploreForCaptureTracking());

### 2020-11-10 src/IrHandler.cpp

@@ -3,6 +3,7 @@
 
 #include "clang/CodeGen/CodeGenAction.h"
 #include "clang/Frontend/CompilerInstance.h"
+#include "llvm/Support/Host.h"
 
 using namespace clang;
 using namespace llvm;

### 2020-11-14 proposal.tex
fix formatting, add first bib entries

@@
 \usepackage{graphicx}
-
-% alternative font if you prefer
-%\usepackage{times}
-
-% for alternative page numbering use the following package
-% and see documentation for commands
-%\usepackage{fancyheadings}
-
-% other potentially useful packages
-%\uspackage{amssymb,amsmath}
-%\usepackage{url}
-%\usepackage{fancyvrb}
-%\usepackage[final]{pdfpages}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Title of project proposal placed here}
-\author{Name of author placed here}
-\date{Date of submission placed here}
+
+    \title{ptr-tidy: Automatic refactoring of raw pointers in C++}
+    \author{Artem Usov}
+    \date{18 December 2020}
     \maketitle
+    \tableofcontents
+    \newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\tableofcontents
-\newpage
+    \section{Introduction}\label{sec:intro}
+
+    Briefly explain the context of the project problem.
+
+    \subsection{A subsection}\label{subsec:a-subsection}
+    Please note your proposal need not follow the included section headings - this is only a suggested structure.
+    Also add subsections etc as required.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}\label{intro}
+    \section{Statement of Problem}\label{sec:statement-of-problem}
 
-briefly explain the context of the project problem
+    Clearly state the problem to be addressed in your forthcoming project.
+    Explain why it would be worthwhile to solve this problem.
 
-\subsection{A subsection}
-Please note your proposal need not follow the included section headings - this is only a suggested structure. Also add subsections etc as required
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-example references: \cite{BK08}
+    \section{Background Survey}\label{sec:background-survey}
+
+    Present an overview of relevant previous work including articles, books, and existing software products.
+    Critically evaluate the strengths and weaknesses of the previous work.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Statement of Problem}
 
-clearly state the problem to be addressed in your forthcoming project. Explain why it would be worthwhile to solve this problem.
+    \section{Proposed Approach}\label{sec:proposed-approach}
+
+    State how you propose to solve the software development problem.
+    Show that your proposed approach is feasible, but identify any risks.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background Survey}
 
-present an overview of relevant previous work including articles, books, and existing software products. Critically evaluate the strengths and weaknesses of the previous work.
+    \section{Work Plan}\label{sec:work-plan}
+
+    Show how you plan to organize your work, identifying intermediate deliverables and dates.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Proposed Approach}
 
-state how you propose to solve the software development problem. Show that your proposed approach is feasible, but identify any risks.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Work Plan}
-
-show how you plan to organize your work, identifying intermediate deliverables and dates.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% it is fine to change the bibliography style if you want
     \bibliographystyle{plain}

### 2020-11-16 proposal.tex
First draft of introduction, start on problem statement

@@
 \usepackage{graphicx}
+\usepackage[square,numbers]{natbib}
+\usepackage{listing}
+\usepackage{chngcntr}
+\usepackage[hyphens]{url}
+\usepackage{hyperref}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \counterwithin{listing}{section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-    Briefly explain the context of the project problem.
+    \subsection{Difficulties with Memory Management}\label{subsec:difficulties-with-memory-management}
 
-    \subsection{A subsection}\label{subsec:a-subsection}
-    Please note your proposal need not follow the included section headings - this is only a suggested structure.
-    Also add subsections etc as required.
+    % TODO insert info about what memory management/errors are
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    Correct memory management is a difficult problem.
+
+    \emph{How difficult one might ask?}
+
+    In a presentation by Matt Miller, a security engineer at Microsoft~\cite{Miller2019}, it is shown that around 70\% of vulnerabilities that are addressed through a security updates are due to memory safety issues.
+    This is at an industry-leading company who are renowned for hiring top talent.
+    Such memory safety vulnerabilities can be exploited, and due to new regulations such as the GDPR, these issues are more commonly exposed to the general public and punished by regulators.
+    Such an example is a fine of £20m fine for a British Airways data breach by the British Information Commissioner's Office~\cite{ICO2020}.
+
+    \subsection{Possible Solutions to Memory Management}\label{subsec:possible-solutions-to-memory-management}
+
+    Memory safety errors occur when languages are used that place the tasks of memory management with the programmer, such as C and C++.
+    However simply not using these languages is not an option since their inherent danger makes them highly performant, and therefore the only option for a project such as a web browser.
+
+    \emph{What can be done then to address this issue?}
+
+    % TODO add Rust performance benchmark reference
+
+    Engineers at both Microsoft~\cite{Thomas2019} and Mozilla~\cite{Hostfelt2019} converge on Rust~\cite{Balasubramanian2017} as a possible solution.
+    Rust is a systems language that offers the same performance as C and C++, however its linear type system and memory ownership model also guarantee memory safety.
+    The rewriting of a program in a new language is a colossal undertaking, especially in language that has a reputation for being difficult to learn.
+    We therefore propose an alternative partial solution to this problem.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-    Clearly state the problem to be addressed in your forthcoming project.
-    Explain why it would be worthwhile to solve this problem.
+    \subsection{Using Smart Pointers for Memory Management}\label{subsec:using-smart-pointers-for-memory-management}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    Generally pointers are used to give the program access to a resource that cannot be directly included in the program itself, such as a file or a thread.
+    However Stroustrup claims that pointers to objects allocated on the free store is dangerous and a 'plain old pointer', or raw pointer as we will refer to them, should not be used to represent ownership~\cite{Stroustrup2018}.
+
+    \begin{listing}
+        \begin{verbatim}
+            void foo(int x) {
+                Shape *p = new Circle{Point{0,0},10};
+                // ...
+                if (x<0) throw Bad_x{}; // potential leak
+                if (x==0) return; // potential leak
+                // ...
+                delete p;
+            }
+        \end{verbatim}
+        \caption{Example of memory leaks using manual management.}
+        \label{fig:manual-leak}
+    \end{listing}
+
+    In \autoref{fig:manual-leak} we can see that when using a raw pointer, there are two cases in which the programmer will never free the allocated memory.
+    Instead we can \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)} to create resource handles which automatically eliminate resource leaks with no added overhead.
+    The equivalent code using \emph{unique pointers} can be seen in \autoref{fig:first-example-unique}.
+    In the comments, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will insert delete statements, now ensuring that the resource will never get leaked.
+
+    \begin{listing}
+        \begin{verbatim}
+            void foo(int x) {
+                auto p = std::make_unique<Circle>(Point{0,0},10);
+                // ...
+                if (x<0) throw Bad_x{}; // delete p inserted here
+                if (x==0) return; // delete p inserted here
+                // ...
+                // delete p inserted here
+            }
+        \end{verbatim}
+        \caption{Example of using a unique pointer to manage memory.}
+        \label{fig:first-example-unique}
+    \end{listing}
+
+    There also exists a \emph{shared pointer} for resources which may not have a single unique owner.
+    Reference counting is used to ensure that after the last owner loses access to the resource, it will finally be freed.
+    Shared pointers are suitable to be used in most situations, except in structures which have cyclic references \autoref{}.
+
+    % TODO add reference of cycle analysis further in paper
+
+    \subsection{Do Smart Pointer Solve Memory Management?}\label{subsec:do-smart-pointer-solve-memory-management?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-    \bibliographystyle{plain}
+    \bibliographystyle{plainnat}
     \bibliography{proposal}

### 2020-11-17 proposal.tex
finish draft problem statement, start background

@@
 \usepackage{listing}
+\usepackage{minted}
 \usepackage{chngcntr}
@@
 \usepackage{hyperref}
+
+\setminted{frame=lines, linenos, autogobble}
 
@@
     % TODO add Rust performance benchmark reference
-
     Engineers at both Microsoft~\cite{Thomas2019} and Mozilla~\cite{Hostfelt2019} converge on Rust~\cite{Balasubramanian2017} as a possible solution.
-    Rust is a systems language that offers the same performance as C and C++, however its linear type system and memory ownership model also guarantee memory safety.
+    Rust is a systems language that offers the same performance as C and C++~\cite{RustBenchmark}, however its linear type system and memory ownership model also guarantee memory safety.
     The rewriting of a program in a new language is a colossal undertaking, especially in language that has a reputation for being difficult to learn.
-    We therefore propose an alternative partial solution to this problem.
+    We therefore propose an alternative partial solution to this problem by using a concept from Rust.
 
@@
     Generally pointers are used to give the program access to a resource that cannot be directly included in the program itself, such as a file or a thread.
-    However Stroustrup claims that pointers to objects allocated on the free store is dangerous and a 'plain old pointer', or raw pointer as we will refer to them, should not be used to represent ownership~\cite{Stroustrup2018}.
+    However Stroustrup claims that pointers to objects allocated on the free store is dangerous and a \emph{plain old pointer}, or raw pointer as we will refer to them, should not be used to represent ownership~\cite{Stroustrup2018}.
 
     \begin{listing}
-        \begin{verbatim}
-            void foo(int x) {
-                Shape *p = new Circle{Point{0,0},10};
-                // ...
-                if (x<0) throw Bad_x{}; // potential leak
-                if (x==0) return; // potential leak
-                // ...
-                delete p;
-            }
-        \end{verbatim}
+        \inputminted{c++}{code/manual-leak.cpp}
         \caption{Example of memory leaks using manual management.}
-        \label{fig:manual-leak}
+        \label{lst:manual-leak}
     \end{listing}
 
-    In \autoref{fig:manual-leak} we can see that when using a raw pointer, there are two cases in which the programmer will never free the allocated memory.
+    In \autoref{lst:manual-leak} we can see that when using a raw pointer, there are two cases in which the programmer will never free the allocated memory.
     Instead we can \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)} to create resource handles which automatically eliminate resource leaks with no added overhead.
-    The equivalent code using \emph{unique pointers} can be seen in \autoref{fig:first-example-unique}.
+    The equivalent code using \emph{unique pointers} can be seen in \autoref{lst:first-example-unique}.
     In the comments, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will insert delete statements, now ensuring that the resource will never get leaked.
@@
     \begin{listing}
-        \begin{verbatim}
-            void foo(int x) {
-                auto p = std::make_unique<Circle>(Point{0,0},10);
-                // ...
-                if (x<0) throw Bad_x{}; // delete p inserted here
-                if (x==0) return; // delete p inserted here
-                // ...
-                // delete p inserted here
-            }
-        \end{verbatim}
+        \inputminted{c++}{code/first-example-unique.cpp}
         \caption{Example of using a unique pointer to manage memory.}
-        \label{fig:first-example-unique}
+        \label{lst:first-example-unique}
     \end{listing}
 
+    % TODO add reference of cycle analysis further in paper
     There also exists a \emph{shared pointer} for resources which may not have a single unique owner.
     Reference counting is used to ensure that after the last owner loses access to the resource, it will finally be freed.
-    Shared pointers are suitable to be used in most situations, except in structures which have cyclic references \autoref{}.
-
-    % TODO add reference of cycle analysis further in paper
+    Shared pointers are suitable to be used in most situations, except in structures which have cyclic references \autoref{sec:cycle-analysis}.
 
     \subsection{Do Smart Pointer Solve Memory Management?}\label{subsec:do-smart-pointer-solve-memory-management?}
+
+    It would appear then that smart pointers could solve most of our memory management issues, however they were only introduced in the C++11 standard.
+    The C++ dialects till then did not have any of these language features, with programs being written predominantly in the C++03 standard.
+    There remain many such so called \emph{legacy} programs which have not upgraded to newer standards, and even then, it is up to the programmer to properly utilise smart pointers in their code.
+
+    We therefore propose to create an automatic refactoring tool \emph{(ptr-tidy)}, that statically analyses C++ code and aims to refactor usage of raw pointers into smart pointers.
+    It will involve a conservative analysis that first and foremost tries to identify raw pointers that only have a single unique owner at any point in the program.
+    If this cannot be guaranteed, then a more conservative refactoring to a shared pointer will be applied.
+
+    % TODO find reference or documentation that standards really are backwards compatible
+    Given that the language standards are mostly backwards-compatible~\cite{BackwardsCompatible}, we can apply these automatic refactoring without changing the runtime behaviour compared to the original programs.
+
+    % TODO reword section title - or move to proposed approach section
+
+    \subsection{The Research Questions}\label{subsec:the-research-questions}
+
+    \emph{"Can the translation of raw pointers into smart pointers using escape analysis be used to make existing C++ code more modern and extendable?"}
+
+    \emph{"Can the tool identify situations in which unique pointers should be used reliably?"}
 
@@
 
-    Present an overview of relevant previous work including articles, books, and existing software products.
-    Critically evaluate the strengths and weaknesses of the previous work.
+    In this section we will look into the frameworks and algorithms that we will use to enable us to read and refactor C++ code, and secondly identify the locations where we can apply our refactorings.
+
+    \subsection{Clang and LLVM}\label{subsec:clang-and-llvm}
+
+    C++ as a language is massive, and the first capability our tool will need is to be able to parse C++ code.
+    Doing this ourselves would be an undertaking that would likely span longer than the time span of the whole project, therefore it was decided early to use an existing parser.
+    The most capable and mature C++ parsers will naturally be the parsers used by the most popular compilers.
+    This leaves us therefore with a choice between GCC~\cite{GCC} and Clang~\cite{Clang}.
+    The literature~\cite{duffy2014} is fairly clear on the fact that the design of GCC makes it unsuitable for integration with other projects.
+    Clang on the other hand allows us access to the Clang abstract syntax tree (AST) and parser internals through consumers and visitors.
+
+    Clang is a front-end to the LLVM framework~\cite{Lattner2004} which transforms C++ code into the LLVM intermediate representation (IR).
+    The LLVM IR is a language-independent, static single assignment (SSA) representation of the program and will be the target on which we apply our analysis to identify refactoring opportunities.
+    We believe this choice will lead to several benefits in terms of the analysis power and future usage of the project:
+    \begin{itemize}
+        \item First, the usage of SSA allows us to efficiently generate a definition-usage graph, therefore allowing an efficient analysis of all the usages of a pointer to identify its number of owners.
+        \item Secondly, by separating the parsing and analysis, we make our analysis stage reusable by any other LLVM frontends.
+    \end{itemize}
+
+    Clang was also found to be a good choice of library in terms of usability since documentation and examples are readily found.
+    The AST that it produces is also rich in information and closely resembles the original code, as can be seen in \autoref{lst:code-to-ast-example}.
+    Information such as the line and column numbers are included, which makes programmatically rewriting source code extremely easy.
+
+    \begin{listing}
+        \inputminted{c++}{code/code-to-ast.cpp}
+        \inputminted[fontsize=\footnotesize]{text}{code/code-to-ast.txt}
+        \caption{Example of converting C++ code to Clang AST}
+        \label{lst:code-to-ast-example}
+    \end{listing}
 

### 2020-11-17 code/code-to-ast.cpp

@@ -1,0 +1,5 @@
+int main() {
+    int a = 4;
+    return a;
+}
+

### 2020-11-17 code/first-example-unique.cpp

@@ -1,0 +1,8 @@
+void foo(int x) {
+    std::unique_ptr<Circle> p = std::make_unique<Circle>(Point{0,0},10);
+    // ...
+    if (x<0) throw Bad_x{}; // delete p inserted here
+    if (x==0) return; // delete p inserted here
+    // ...
+    // delete p inserted here
+}

### 2020-11-17 code/manual-leak.cpp

@@ -1,0 +1,8 @@
+void foo(int x) {
+    Shape *p = new Circle{Point{0,0},10};
+    // ...
+    if (x<0) throw Bad_x{}; // potential leak
+    if (x==0) return; // potential leak
+    // ...
+    delete p;
+}

### 2020-11-20 proposal.tex
Add section on other memory safe systems languages, mention further work on transpiling to rust

@@
     In a presentation by Matt Miller, a security engineer at Microsoft~\cite{Miller2019}, it is shown that around 70\% of vulnerabilities that are addressed through a security updates are due to memory safety issues.
-    This is at an industry-leading company who are renowned for hiring top talent.
+    In another presentation at the Linux Security Summit~\cite{Gaynor2020} it is shown that in several other top projects such as Firefox, macOS, Ubuntu and Android all had over half of their CVEs attributed to issues with memory safety.
+    This is at industry-leading companies who are renowned for hiring top talent.
     Such memory safety vulnerabilities can be exploited, and due to new regulations such as the GDPR, these issues are more commonly exposed to the general public and punished by regulators.
@@
 
+    \subsection{Memory Safe Languages}\label{subsec:rust}
+
+    There have been many proposed systems languages such as Rust, Swift \footnote{\url{https://developer.apple.com/swift/}}, D \footnote{\url{https://dlang.org/}}, Nim \footnote{\url{https://nim-lang.org/}} or Ada (SPARK) \footnote{\url{https://adacore.com/}}, however none of these have arguably become mainstream, with Rust and Ada coming closest.
+
+    Swift remains mostly a language for the Apple MacOS and iOS environments for application creation, however it has seen some usage for system programs by companies such as IBM \footnote{\url{https://github.com/ibm-swift}}.
+    Ada or its subset language SPARK are used extensively in certain industries to create verifiable programs for critical systems such as in avionics.
+    It is perhaps most well-know for being used in the F-22 fighter aircraft~\cite{malas1997f}, however the use of Ada for those systems show its other weakness in productivity time as it was the cause for several delays in the project, and so it largely remains unused in areas which do not call for such meticulous safety.
+
+    Rust has been the most language in attractive widespread adoption as a systems language.
+    Its popularity could be quantified by considering that it is being considered for usage in writing Linux kernel modules \footnote{\url{https://lore.kernel.org/lkml/CAKwvOdmuYc8rW_H4aQG4DsJzho=F+djd68fp7mzmBp3-wY--Uw@mail.gmail.com/T}}, which cannot be said for C++.
+    Following the advice given by many that most systems programs should be written in Rust from now on, we see a potential future usage of our work in the creation of a C++ to Rust transpiler.
+
+    Whilst the automatic refactoring of raw pointers into smart pointers solves some memory issues, the language still offers other ways to invalidate memory safety.
+    Equivalent idiomatic Rust would be able to find some of these other errors at compile time.
+    A C++ to Rust transpiler would be a much larger undertaking, however we can envision how our project could be helpful.
+    Rust affine types~\cite{Balasubramanian2017} are semantically equivalent to C++ unique pointers, in that all values have a single unique owner at any point.
+    Therefore our work can identify which pointers are unique and which are not, and those that are not cannot be transpiled into safe Rust, instead requiring to use unsafe blocks to represent such values.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-11-21 proposal.tex
find and fill in missing references

@@
 
-    % TODO add Rust performance benchmark reference
     Engineers at both Microsoft~\cite{Thomas2019} and Mozilla~\cite{Hostfelt2019} converge on Rust~\cite{Balasubramanian2017} as a possible solution.
-    Rust is a systems language that offers the same performance as C and C++~\cite{RustBenchmark}, however its linear type system and memory ownership model also guarantee memory safety.
+    Rust is a systems language that offers the same performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model also guarantee memory safety.
     The rewriting of a program in a new language is a colossal undertaking, especially in language that has a reputation for being difficult to learn.
@@
 
-    % TODO add reference of cycle analysis further in paper
     There also exists a \emph{shared pointer} for resources which may not have a single unique owner.
     Reference counting is used to ensure that after the last owner loses access to the resource, it will finally be freed.
-    Shared pointers are suitable to be used in most situations, except in structures which have cyclic references \autoref{sec:cycle-analysis}.
+    Shared pointers are suitable to be used in most situations, except in structures which have cyclic references \autoref{subsec:shared-pointer-cycle-analysis}.
 
@@
 
-    % TODO find reference or documentation that standards really are backwards compatible
-    Given that the language standards are mostly backwards-compatible~\cite{BackwardsCompatible}, we can apply these automatic refactoring without changing the runtime behaviour compared to the original programs.
+    Given that the language standards are mostly backwards-compatible~\cite{Winters2020}, we can apply these automatic refactoring without changing the runtime behaviour compared to the original programs.
 
@@
         \label{lst:code-to-ast-example}
+        % \centerline{\includegraphics{images/code-to-ast}}
     \end{listing}
@@
 
+    \subsection{Shared Pointer Cycle Analysis}\label{subsec:shared-pointer-cycle-analysis}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-11-24 proposal.tex
add section on garbage collection and pointer analysis

@@
 \usepackage{listing}
-\usepackage{minted}
+\usepackage[labelfont=bf]{caption}
+\usepackage[outputdir=out]{minted}
 \usepackage{chngcntr}
@@
 
+    % TODO insert some information about heap vs stack
+
     In \autoref{lst:manual-leak} we can see that when using a raw pointer, there are two cases in which the programmer will never free the allocated memory.
     Instead we can \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)} to create resource handles which automatically eliminate resource leaks with no added overhead.
-    The equivalent code using \emph{unique pointers} can be seen in \autoref{lst:first-example-unique}.
+    Firstly there are \emph{unique pointers} which are a pointer that has only a single unique owner at any time.
+    We can therefore fix the cases in which we leak the memory using a unique pointer which can be seen in in \autoref{lst:first-example-unique}.
     In the comments, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will insert delete statements, now ensuring that the resource will never get leaked.
+    The ownership of a unique pointer can be transferred between variables using \mintinline{c++}{std::move} which moves the ownership from one variable to another.
 
@@
         \inputminted[fontsize=\footnotesize]{text}{code/code-to-ast.txt}
-        \caption{Example of converting C++ code to Clang AST}
+        \centerline{\includegraphics{images/code-to-ast}}
+        \caption{Example of converting C++ code to Clang AST. All diagrams show equivalent representation of a small C++ code snippet}
         \label{lst:code-to-ast-example}
-        % \centerline{\includegraphics{images/code-to-ast}}
     \end{listing}
+
+    \subsection{Pointer Analysis}\label{subsec:pointer-analysis}
+
+    The main capability of the tool is to be able to identify what the maximum number of owners a pointer has at all points of the program execution, for all pointers in any given program.
+    If this number for a given raw pointer is one, then we can go about converting all usages of that raw pointer into a unique pointer.
+    This will involve changing the type of the initial declaration, its definition as well as removing all deletions of the pointer, which can be seen in \autoref{lst:manual-leak} and \autoref{lst:first-example-unique}.
+    If it is more than one then the invariants of a unique pointer cannot be satisfied, and we instead have to use the less performant shared smart pointer.
+    A key intuition, mentioned in \autoref{subsec:using-smart-pointers-for-memory-management}, is that a shared pointer can be substituted for most uses of a raw pointer, making it the default fallback.
+
+    A pointer may have more than one owner at a time by copying the pointer to another variable.
+    Effectively the address of the underlying region of memory can be accessed through more than one variable, and any accesses of one variable affect the others.
+    For example, a pointer may be freed using one variable, and this would invalidate the other variable, causing a memory safety violation if we try to use it.
+
+    We can formalise the notion of copying a pointer by introducing the concept of \emph{escape analysis}.
+    Escape analysis is used to determine whether an object \emph{escapes} or is accessible from outside the method or thread that created the object~\cite{Choi1999}.
+    It has been used in languages such as Java~\cite{Choi1999} to determine whether an object that is created in a method escapes, and if it does not, meaning it is only used locally in the method, then it can be allocated on the stack rather than the heap as a performance optimisation.
+    Clang also uses escape analysis for a similar kind of optimisation~\footnote{\url{https://github.com/llvm/llvm-project/blob/master/llvm/lib/Analysis/CaptureTracking.cpp}} and also defines the term \emph{pointer capture}.
+    A pointer value is captured if the function makes a copy of any part of the pointer that outlives the call.
+    We can then see that any pointer that escapes must also be captured since there is no way to refer to an object without being able to determine its address, but not visa versa.
+    In \autoref{lst:capture-example} we can see an example where a pointer may be captured but does not escape and in \autoref{lst:escape-example} we see an example of an escaped pointer.
+
+    \begin{listing}
+        \inputminted{c++}{code/capture-example.cpp}
+        \caption{The pointer is captured in isEven since the pointer is used for a value that outlives its call, but not captured since the value of the pointer is not stored anywhere outside the function or thread}
+        \label{lst:capture-example}
+    \end{listing}
+
+    \begin{listing}
+        \inputminted{c++}{code/escape-example.cpp}
+        \caption{The pointer is both captured and escapes since the pointer and its value outlive the call of the function and it is accessible by any other method or thread since it is copied to a global variable}
+        \label{lst:escape-example}
+    \end{listing}
+
+    We can now see that invariant of the unique pointer are only violated in two conditions:
+    \begin{itemize}
+        \item First, a pointer escapes, as then we do not know where or how the pointer may be used, so we must take a conservative approach and assume the pointer no longer has at most a single unique owner.
+        \item Secondly, if a pointer is captured more than once, as we can model a single capture as a move of the unique pointer.
+    \end{itemize}
+
+    \subsubsection{Escape Analysis Algorithms}\label{subsubsec:escape-analysis-algorithms}
 
@@
 
-    There have been many proposed systems languages such as Rust, Swift \footnote{\url{https://developer.apple.com/swift/}}, D \footnote{\url{https://dlang.org/}}, Nim \footnote{\url{https://nim-lang.org/}} or Ada (SPARK) \footnote{\url{https://adacore.com/}}, however none of these have arguably become mainstream, with Rust and Ada coming closest.
+    There have been many proposed systems languages such as Rust, Swift~\footnote{\url{https://developer.apple.com/swift/}}, D~\footnote{\url{https://dlang.org/}}, Nim~\footnote{\url{https://nim-lang.org/}} or Ada (SPARK)~\footnote{\url{https://adacore.com/}}, however none of these have arguably become mainstream, with Rust and Ada coming closest.
 
-    Swift remains mostly a language for the Apple MacOS and iOS environments for application creation, however it has seen some usage for system programs by companies such as IBM \footnote{\url{https://github.com/ibm-swift}}.
+    Swift remains mostly a language for the Apple MacOS and iOS environments for application creation, however it has seen some usage for system programs by companies such as IBM~\footnote{\url{https://github.com/ibm-swift}}.
     Ada or its subset language SPARK are used extensively in certain industries to create verifiable programs for critical systems such as in avionics.
@@
     Rust has been the most language in attractive widespread adoption as a systems language.
-    Its popularity could be quantified by considering that it is being considered for usage in writing Linux kernel modules \footnote{\url{https://lore.kernel.org/lkml/CAKwvOdmuYc8rW_H4aQG4DsJzho=F+djd68fp7mzmBp3-wY--Uw@mail.gmail.com/T}}, which cannot be said for C++.
+    Its popularity could be quantified by considering that it is being considered for usage in writing Linux kernel modules~\footnote{\url{https://lore.kernel.org/lkml/CAKwvOdmuYc8rW_H4aQG4DsJzho=F+djd68fp7mzmBp3-wY--Uw@mail.gmail.com/T}}, which cannot be said for C++.
     Following the advice given by many that most systems programs should be written in Rust from now on, we see a potential future usage of our work in the creation of a C++ to Rust transpiler.
@@
     Therefore our work can identify which pointers are unique and which are not, and those that are not cannot be transpiled into safe Rust, instead requiring to use unsafe blocks to represent such values.
+
+    \subsubsection{Garbage Languages}
+
+    Garbage collected languages encompass some of the most popular programming languages such as Java and Python.
+    They are considered to be easier languages than C++ since they remove the need for memory management.
+    In C++, if memory is properly freed, it gets leaked and can eventually cause out of memory errors.
+    Garbage collectors instead run periodically to identify leaked memory and free it.
+    Garbage collectors instead run periodically and scan the memory of the program to find objects in memory which were allocated but are no longer used anymore, i.e leaked and then free the leaked objects.
+    Languages such as Java~\cite{Tauro2012} use this fully by allocating almost every object on the heap and not freeing any object, instead relying on the garbage collector to clean up.
+    They are therefore considered to be easier and more productive to use.
+
+    Garbage collection does also exist for C++ such as the Boehm-Demers-Weiser conservative garbage collector~\cite{boehm2002garbage}.
+    So why is garbage collection not used in every language?
+    Garbage collection is not zero-cost as the garbage collection algorithm typically stops the execution of the running program to clean up the memory.
+    This means that latency or pauses is introduced to every program, and can occur at any point in the execution of the program, such as during the hot path of the program.
+    This has led to garbage collected languages being unsuitable for use in systems programs such as OS's, as for example it would be undesirable to have seemingly random latency when interacting with the system.
+    There has been research in modern garbage collection algorithms such as ZGC~\cite{ZGC} and Shenandoah~\cite{flood2016shenandoah} for Java which aim to reduce the number and duration of pauses where the garbage collector runs, increasing the responsiveness of the Java virtual machine.
+    However, even with much reduced duration of pauses, it may still be unsuitable for some systems with strict latency requirements such as OS's or financial systems.
 

### 2020-11-24 code/capture-example.cpp

@@ -1,0 +1,8 @@
+bool isOdd(int *i) {
+    return (unsigned long) i & 1;
+}
+
+int main() {
+    int *pointer = new int(2);
+    return isOdd(pointer);
+}

### 2020-11-24 code/escape-example.cpp

@@ -1,0 +1,13 @@
+int *globalPointer;
+
+void escape() {
+    int *pointer = new int(0);
+    globalPointer = pointer;
+    // ...
+}
+
+int main() {
+    // ...
+    escape();
+    // ...
+}

### 2020-11-24 example/ptr.cpp
dumb implementation of replacing pointer types with shared_ptrs

@@ -7,5 +7,6 @@ int *test(int *b) {
 
 int main() {
   int b = 2;
+  int *c = &a;
   test(&b);
 }

### 2020-11-24 src/AstHandler.cpp

@@ -1,4 +1,5 @@
 #include "AstHandler.h"
+#include "clang/Tooling/Tooling.h"
 
 using namespace clang::ast_matchers;
 using namespace clang::tooling;

### 2020-11-24 src/AstHandler.h

@@ -1,8 +1,6 @@
 #pragma once
 
 #include "clang/ASTMatchers/ASTMatchFinder.h"
-#include "clang/ASTMatchers/ASTMatchers.h"
-#include "clang/Tooling/Tooling.h"
 
 class AstHandler {
 public:

### 2020-11-24 src/Callback.cpp

@@ -1,9 +1,5 @@
 #include "Callback.h"
-
-#include "clang/AST/Mangle.h"
 #include "CaptureTracking.h"
-#include "llvm/IR/ValueSymbolTable.h"
-
 #include "Helper.h"
 
 using namespace llvm;
@@ -11,6 +7,8 @@ using namespace clang;
 using namespace clang::ast_matchers;
 
 void CountCallback::run(const MatchFinder::MatchResult &result) {
+  m_rewriter.setSourceMgr(*result.SourceManager, result.Context->getLangOpts());
+
   auto *var = result.Nodes.getNodeAs<VarDecl>("");
   if (!var) {
     exit("Encountered not a VarDecl");
@@ -37,5 +35,13 @@ void CountCallback::run(const MatchFinder::MatchResult &result) {
     errs() << " is captured\n";
   } else {
     errs() << " is not captured\n";
+  }
+
+  if (var->getType()->isPointerType()) {
+    std::string newType = "std::shared_ptr<";
+    newType.append(var->getType()->getPointeeType().getAsString());
+    newType.append("> ");
+    m_rewriter.ReplaceText(SourceRange(var->getTypeSpecStartLoc(), var->getTypeSpecEndLoc()),
+                           newType);
   }
 }

### 2020-11-24 src/Callback.h

@@ -1,15 +1,18 @@
 #pragma once
 
-#include "clang/ASTMatchers/ASTMatchFinder.h"
-#include "clang/ASTMatchers/ASTMatchers.h"
-
 #include "IrHandler.h"
+#include "clang/ASTMatchers/ASTMatchFinder.h"
+#include "clang/Rewrite/Core/Rewriter.h"
 
 class CountCallback : public clang::ast_matchers::MatchFinder::MatchCallback {
+private:
   IrHandler m_irHandler;
+  clang::Rewriter m_rewriter;
 
 public:
   explicit CountCallback(const llvm::StringRef codePath) : m_irHandler(codePath){};
 
   void run(const clang::ast_matchers::MatchFinder::MatchResult &result) override;
+
+  clang::Rewriter &getRewriter() { return m_rewriter; }
 };

### 2020-11-24 src/Helper.h

@@ -1,7 +1,7 @@
 #pragma once
 
-#include "clang/AST/ASTContext.h"
-#include "llvm/Support/raw_os_ostream.h"
+#include "clang/AST/Mangle.h"
+#include "llvm/IR/ValueSymbolTable.h"
 
 inline void exit(const llvm::StringRef &message) {
   llvm::errs() << message << "\n";
@@ -24,7 +24,7 @@ inline std::string getMangledName(clang::ASTContext *context, const clang::Named
 }
 
 inline llvm::Value *getLocalValue(const clang::VarDecl *var, const llvm::Module &module,
-                                     clang::ASTContext *context) {
+                                  clang::ASTContext *context) {
   auto *parentFunctionAST = dyn_cast<clang::FunctionDecl>(var->getParentFunctionOrMethod());
   auto *parentFunctionIR = module.getFunction(getMangledName(context, parentFunctionAST));
   auto *valueSymbol = parentFunctionIR->getValueSymbolTable();

### 2020-11-24 src/IrHandler.cpp

@@ -1,9 +1,8 @@
 #include "IrHandler.h"
-#include <sstream>
-
 #include "clang/CodeGen/CodeGenAction.h"
 #include "clang/Frontend/CompilerInstance.h"
 #include "llvm/Support/Host.h"
+#include <sstream>
 
 using namespace clang;
 using namespace llvm;

### 2020-11-24 src/IrHandler.h

@@ -3,8 +3,6 @@
 #include "boost/optional.hpp"
 #include "llvm/IR/LLVMContext.h"
 #include "llvm/IR/Module.h"
-#include "llvm/Support/MemoryBuffer.h"
-#include <string>
 
 class IrHandler {
   llvm::LLVMContext m_context;

### 2020-11-24 src/main.cpp

@@ -13,4 +13,6 @@ int main(int argc, char *argv[]) {
 
   CountCallback callback(argv[1]);
   AstHandler::runCallback(callback, sourceFileOrError.get()->getBuffer());
+
+  callback.getRewriter().getEditBuffer(callback.getRewriter().getSourceMgr().getMainFileID()).write(llvm::errs());
 }

### 2020-11-24 src/AstHandler.cpp
Add shared pointer upgrading for function return types and variable initialisers

@@ -7,7 +7,7 @@ using namespace clang::tooling;
 void AstHandler::runCallback(MatchFinder::MatchCallback &callback, const llvm::StringRef code) {
   MatchFinder Finder;
 
-  Finder.addMatcher(varDecl().bind(""), &callback);
+  Finder.addMatcher(namedDecl().bind(""), &callback);
   std::unique_ptr<FrontendActionFactory> Factory(newFrontendActionFactory(&Finder));
   runToolOnCode(Factory->create(), code);
 }

### 2020-11-24 src/Callback.cpp

@@ -9,9 +9,18 @@ using namespace clang::ast_matchers;
 void CountCallback::run(const MatchFinder::MatchResult &result) {
   m_rewriter.setSourceMgr(*result.SourceManager, result.Context->getLangOpts());
 
+  auto *function = result.Nodes.getNodeAs<FunctionDecl>("");
+  if (function && function->getReturnType()->isPointerType()) {
+    std::string newType = "std::shared_ptr<";
+    newType.append(function->getReturnType()->getPointeeType().getAsString());
+    newType.append("> ");
+    m_rewriter.ReplaceText(function->getReturnTypeSourceRange(), newType);
+    return;
+  }
+
   auto *var = result.Nodes.getNodeAs<VarDecl>("");
   if (!var) {
-    exit("Encountered not a VarDecl");
+    return;
   }
 
   if (var->hasGlobalStorage()) {
@@ -38,10 +47,19 @@ void CountCallback::run(const MatchFinder::MatchResult &result) {
   }
 
   if (var->getType()->isPointerType()) {
-    std::string newType = "std::shared_ptr<";
-    newType.append(var->getType()->getPointeeType().getAsString());
-    newType.append("> ");
+    std::string newType = "std::shared_ptr<" + var->getType()->getPointeeType().getAsString() + ">";
     m_rewriter.ReplaceText(SourceRange(var->getTypeSpecStartLoc(), var->getTypeSpecEndLoc()),
                            newType);
+
+    if (var->hasInit()) {
+      auto init = var->getInit();
+
+      auto decl_ref = dyn_cast<DeclRefExpr>(*init->child_begin());
+      auto name = decl_ref->getDecl()->getNameAsString();
+      auto type = decl_ref->getDecl()->getType().getAsString();
+
+      auto newType = "std::make_shared<" + type + ">(" + name + ")";
+      m_rewriter.ReplaceText(init->getSourceRange(), newType);
+    }
   }
 }

### 2020-11-27 proposal.tex
fix typos, add section on escape algorithms

@@
 \usepackage[hyphens]{url}
+\usepackage{tikz}
 \usepackage{hyperref}
@@
     The most capable and mature C++ parsers will naturally be the parsers used by the most popular compilers.
-    This leaves us therefore with a choice between GCC~\cite{GCC} and Clang~\cite{Clang}.
+    This leaves us therefore with a choice between GCC~\footnote{https://gcc.gnu.org/} and Clang~\footnote{https://clang.llvm.org/}.
     The literature~\cite{duffy2014} is fairly clear on the fact that the design of GCC makes it unsuitable for integration with other projects.
@@
     Clang is a front-end to the LLVM framework~\cite{Lattner2004} which transforms C++ code into the LLVM intermediate representation (IR).
+    % TODO define SSA
     The LLVM IR is a language-independent, static single assignment (SSA) representation of the program and will be the target on which we apply our analysis to identify refactoring opportunities.
@@
     It has been used in languages such as Java~\cite{Choi1999} to determine whether an object that is created in a method escapes, and if it does not, meaning it is only used locally in the method, then it can be allocated on the stack rather than the heap as a performance optimisation.
+    % TODO pin footnote to specific commit possibly
     Clang also uses escape analysis for a similar kind of optimisation~\footnote{\url{https://github.com/llvm/llvm-project/blob/master/llvm/lib/Analysis/CaptureTracking.cpp}} and also defines the term \emph{pointer capture}.
@@
         \inputminted{c++}{code/capture-example.cpp}
-        \caption{The pointer is captured in isEven since the pointer is used for a value that outlives its call, but not captured since the value of the pointer is not stored anywhere outside the function or thread}
+        \caption{The pointer is captured in isOdd since the pointer is used for a value that outlives its call, but does not escape since the value of the pointer is not stored anywhere outside the function or thread}
         \label{lst:capture-example}
@@
     \subsubsection{Escape Analysis Algorithms}\label{subsubsec:escape-analysis-algorithms}
+
+    Given how we have defined how escape analysis can be used to identify areas of refactoring, we now need to identify a suitable algorithm.
+
+    There has been extensive research on escape analysis for Java, where escape analysis is used to determine whether a object escapes the current method, as then if it can be proven that the object does not escape the method, then it is local to the method invocation and it can be allocated on the stack rather than as a heap variable (the default for all Java objects).
+
+    The escape analysis algorithm described by \citet{Choi1999} is used as the default escape analysis algorithm in the Oracle Java Virtual Machine implementation~\footnote{https://docs.oracle.com/en/java/javase/15/vm/java-hotspot-virtual-machine-performance-enhancements.html} and is widely cited so forms the basis of our investigation.
+    Due to how similar C++ and Java are in terms of their Object-Oriented features, will analyse their algorithm as well as the implementations of escape analysis in the LLVM and Go compilers.
+
+    \citet{Choi1999} introduce the abstraction of a \emph{connection graph} (CG) which we will see is also used in other implementations.
+    The connection graph captures relationships between heap-allocated objects and object references.
+    The graph is used to perform reachability analysis to determine if an is local to a thread or method.
+    Most importantly, this abstraction allows for a more powerful interprocedural analysis to be performed.
+
+    The CG is a directed graph $CG = (N_{o} \cup N_{r}, E_{p} \cup E_{d} \cup E_{f})$ where $N_{o}$ is the set of objects.
+    $N_{r}$ is the set of reference nodes which include local variables, parameters, fields and global reference variables.
+    $E_{p}$ is the set of points-to edges which exists between a reference node which points to object node.
+    $E_{d}$ is the set of deferred edges, where a deferred edge from one node $p$ to node $q$ signifies it points to what $q$ points to.
+    $E_{f}$ is the set of of all edges between objects that are a field of another object.
+    We can see in \autoref{fig:connection-graph-choi} how a connection graph would be constructed for simple Java statements.
+
+    % TODO replace with tikz graphics
+
+    \begin{figure}
+        \centering
+        \begin{tikzpicture}[scale=0.5]
+            \draw (0, 0) circle (1) node{a};
+            \draw[->, line width=1pt] (1, 0) -- node[above] {P} (3, 0);
+            \draw (3, -1) rectangle node{S1} (5, 1);
+
+            \draw (1, -5) circle (1) node{f};
+            \draw (4, -5) circle (1) node{g};
+            \draw (7, -5) circle (1) node{h};
+            \draw[->, line width=1pt] (4, -1) -- node[above] {F} (1, -4);
+            \draw[->, line width=1pt] (4, -1) -- node[above left] {F} (4, -4);
+            \draw[->, line width=1pt] (4, -1) -- node[above] {F} (7, -4);
+
+            \node[text width=4.5cm] at (0, 3) {S1: T a = new T(\ldots)};
+        \end{tikzpicture}
+        \begin{tikzpicture}[scale=0.5]
+            \draw (-4, 0) circle (1) node{b};
+            \draw[->, line width=1pt, dashed] (-3, 0) -- node[above] {D} (-1, 0);
+            \draw (0, 0) circle (1) node{a};
+            \draw[->, line width=1pt] (1, 0) -- node[above] {P} (3, 0);
+            \draw (3, -1) rectangle node{S1} (5, 1);
+
+            \draw (1, -5) circle (1) node{f};
+            \draw (4, -5) circle (1) node{g};
+            \draw (7, -5) circle (1) node{h};
+            \draw[->, line width=1pt] (4, -1) -- node[above] {F} (1, -4);
+            \draw[->, line width=1pt] (4, -1) -- node[above left] {F} (4, -4);
+            \draw[->, line width=1pt] (4, -1) -- node[above] {F} (7, -4);
+
+            \node[text width=4.5cm] at (0, 3) {S2: T b = a};
+        \end{tikzpicture}
+        \caption{A simple connection graph.
+Boxes indicate object nodes and circles indicate reference nodes (including field reference nodes).
+Solid edges indicate points-to edge, dashed edges indicate deferred edges, and edges from boxes to circles indicate field edges~\cite{Choi1999}}
+        \label{fig:connection-graph-choi}
+    \end{figure}
+
+    For each node, we associate an escape set, of either \emph{NoEscape}, \emph{ArgEscape} or \emph{GlobalEscape}.
+    \emph{NoEscape} means that the object does not escape the method in which it was created
+    \emph{ArgEscape}, with respect to a method, means that the object escapes that method via the method arguments or return value,
+    but does not escape the thread in which it is created.
+    Finally, \emph{GlobalEscape} means that the object is regarded as escaping globally.
+    The initial state for each global node is \emph{GlobalEscape} and \emph{NoEscape} for all other nodes, unless otherwise stated.
+
+    The escape analysis is executed through their defined construction of intraprocedural graphs for each method call.
+    We then proceed with the interprocedural by using the CG of the callee to update the CG of the caller.
+
+    At the completion of the escape analysis, all objects which are marked NoEscape and cannot be reached by any node whose state is not \emph{NoEscape} do not escape and are local to the thread in which they are created.
+    We show in \autoref{subsec:Interprocedural-analysis} how we can adapt this algorithm to check if the invarients defined in \autoref{subsec:pointer-analysis} are met.
+
+    % TODO move this to the referenced section - Key insight is that we can use defined analysis to check if object escapes.
+but if each node has at most one object incoming node, it means they are captured at most once
+
+    Go is another language that uses a escape analysis to allocate objects on the stack rather than the heap.
+    Their escape analysis algorithms is not defined in the specification of the language, however we can look at the source~\footnote{https://github.com/golang/go/blob/master/src/cmd/compile/internal/gc/escape.go} to determine what algorithm they use.
+    Fortunately the comments give a summary of their algorithm, which involves constructing a weighted graph where vertices represent variables, and edges represent assignments between variables.
+    They walk the graph, looking for assignment paths that violate the invariants of a variable being allocated on the stack and mark those as requiring heap allocation
+    We can see that this approach is highly similar to the use of connection graphs by \citet{Choi1999}, however there is lacking documentation on how they perform interprocedural analysis so their analysis is perhaps weaker.
+
+    Finally, LLVM also uses escape analysis to remove any unnecessary heap allocations.
+    The escape analysis algorithm is also not formally defined, however we can also look at the source code~\footnote{https://github.com/llvm/llvm-project/blob/master/llvm/lib/Analysis/CaptureTracking.cpp}.
+    We see that the \mintinline{c++}{PointerMayBeCaptured} function uses the fact that the single static assignment (SSA) in the LLVM IR allows for efficient creation of definition-usage chains.
+    For any defined value, such as a value initialising a variable, we can traverse all its usages, such as copying to another variable.
+    To determine if an object escapes, we analyse all its usages, and all their usages and so on.
+    For each usage, we then determine if it is an instruction which would cause the object to escape.
+    This leads to a roughly similar analysis as the work by \citet{Choi1999} when considering the intraprocedural analysis alone, as we informally explore a graph of the nodes that represent objects and nodes that obtain references to objects by using the original value in a copy instruction.
+    However, its assumption about different instructions are conservative, and it also has no interprocedural analysis, making it the weakest analysis so far.
 
@@
 
-    \subsubsection{Garbage Languages}
+    \subsubsection{Garbage Collected Languages}
 
     Garbage collected languages encompass some of the most popular programming languages such as Java and Python.
-    They are considered to be easier languages than C++ since they remove the need for memory management.
-    In C++, if memory is properly freed, it gets leaked and can eventually cause out of memory errors.
-    Garbage collectors instead run periodically to identify leaked memory and free it.
+    They are considered to be easier languages than C++ since they remove the need for manual memory management by the programmer.
+    In C++, if memory is improperly freed, it gets leaked and can eventually cause out of memory errors.
     Garbage collectors instead run periodically and scan the memory of the program to find objects in memory which were allocated but are no longer used anymore, i.e leaked and then free the leaked objects.

### 2020-12-04 proposal.tex
Add proposed approach and workplan section

@@
 \usepackage{tikz}
+\usepackage{pgfgantt}
+\usepackage{lscape}
 \usepackage{hyperref}
@@
     \counterwithin{listing}{section}
+    \counterwithin{figure}{section}
 
@@
     If this number for a given raw pointer is one, then we can go about converting all usages of that raw pointer into a unique pointer.
-    This will involve changing the type of the initial declaration, its definition as well as removing all deletions of the pointer, which can be seen in \autoref{lst:manual-leak} and \autoref{lst:first-example-unique}.
     If it is more than one then the invariants of a unique pointer cannot be satisfied, and we instead have to use the less performant shared smart pointer.
@@
         \item First, a pointer escapes, as then we do not know where or how the pointer may be used, so we must take a conservative approach and assume the pointer no longer has at most a single unique owner.
-        \item Secondly, if a pointer is captured more than once, as we can model a single capture as a move of the unique pointer.
+        \item Secondly, if a pointer is captured at most once, as we can model a single capture as a move of the unique pointer.
     \end{itemize}
@@
 
-    % TODO replace with tikz graphics
-
     \begin{figure}
@@
             \draw[->, line width=1pt] (4, -1) -- node[above] {F} (1, -4);
-            \draw[->, line width=1pt] (4, -1) -- node[above left] {F} (4, -4);
+            \draw[->, line width=1pt] (4, -1) -- node[below left] {F} (4, -4);
             \draw[->, line width=1pt] (4, -1) -- node[above] {F} (7, -4);
@@
     At the completion of the escape analysis, all objects which are marked NoEscape and cannot be reached by any node whose state is not \emph{NoEscape} do not escape and are local to the thread in which they are created.
-    We show in \autoref{subsec:Interprocedural-analysis} how we can adapt this algorithm to check if the invarients defined in \autoref{subsec:pointer-analysis} are met.
-
-    % TODO move this to the referenced section - Key insight is that we can use defined analysis to check if object escapes.
-but if each node has at most one object incoming node, it means they are captured at most once
+    We show in \autoref{subsubsec:interprocedural-analysis} how we can adapt this algorithm to check if the invariants defined in \autoref{subsec:pointer-analysis} are met.
 
@@
 
-    State how you propose to solve the software development problem.
-    Show that your proposed approach is feasible, but identify any risks.
+    % TODO rename section to something better
+
+    \subsection{Current Progress}\label{subsec:current-progress}
+
+    As mentioned in \autoref{subsec:clang-and-llvm}, we plan on using Clang for parsing C++ code into an AST and LLVM IR for the analysis.
+    We can employ the visitor pattern to traverse the AST and look for nodes of interest, such as variable declaration nodes \emph{(varDecl)}.
+    Given a node in the AST, we can find its equivalent declaration in the IR\@.
+    This is mainly possible since we can see from line 7 in \autoref{lst:code-to-ast-ir} that the variable names can matched from to AST to the IR\@.
+
+    \begin{listing}
+        \inputminted[fontsize=\footnotesize, breaklines]{llvm}{code/code-to-ast-summary.ll}
+        \caption{The LLVM IR representation of the program in \autoref{lst:code-to-ast-example}. Listing has been summarized for clarity.}
+        \label{lst:code-to-ast-ir}
+    \end{listing}
+
+    We can then apply escape analysis for this particular value, which will give us a boolean \mintinline{c++}{true} or \mintinline{c++}{false} result.
+    Based on the returned value, we know if the value does not escape and so we can refactor all the usages of this value.
+    This will involve changing the type of the initial declaration, its definition, as well as removing all deletions of the pointer, which can be seen in \autoref{lst:manual-leak} and \autoref{lst:first-example-unique}.
+    This is repeated for all usages of the value as well, such as being used as a formal argument.
+    The diagram in \autoref{fig:design} shows a visual representation of this proposed design.
+
+    \begin{figure}
+        \centering
+        \begin{tikzpicture}[scale=0.33, every node/.style={scale=0.75}]
+            \draw (0,0) rectangle node{program.cpp} (6, 3);
+            \draw (10, 0) rectangle node{parser} (16, 3);
+            \draw[->, line width=1pt] (6, 1.5) -- node[above] {input} (10, 1.5);
+
+            \draw (19, 3) rectangle node{AST} (25, 6);
+            \draw (19, 0) rectangle node{IR} (25, -3);
+            \draw[->, line width=1pt] (16, 1.5) -- (19, 4.5);
+            \draw[->, line width=1pt] (16, 1.5) -- (19, -1.5);
+
+            \draw (28, 0) rectangle node{Analysis} (34, 3);
+            \draw[->, line width=1pt] (25, 4.5) -- (28, 1.5);
+            \draw[->, line width=1pt] (25, -1.5) -- (28, 1.5);
+
+            \draw (37, 3) rectangle node{Rewriter} (43, 6);
+            \draw[->, line width=1pt] (25, 4.5) --  (37, 4.5);
+            \draw[->, line width=1pt] (34, 1.5) -- node[below right, text width=3cm] {refactoring decision} (37, 4.5);
+
+            \draw (46, 0) rectangle node{refactored.cpp} (52, 3);
+            \draw[->, line width=1pt] (43, 4.5) -- node[above right] {output} (46, 1.5);
+
+            \draw (10, -5) rectangle (43, 8);
+            \node[text width=0cm] at (10, 6) {Ptr-tidy};
+
+        \end{tikzpicture}
+        \caption{Proposed design of Ptr-tidy tool.
+Parser is the Clang C++ parser which produces an AST and IR representation of the input code.
+Analysis is the chosen escape analysis algorithm.
+It produces a decision of whether the pointer is suitable for refactoring, and passes this to the refactoring component which rewrites the code using smart pointers to produce the safer output program.}
+        \label{fig:design}
+    \end{figure}
+
+    This design mirrors the modular design of LLVM where because the analysis is performed on the IR, it could be reused by the frontend of a different language and not necessarily Clang as we do.
+    Secondly this also enables us to flexibly test different algorithms as the C++ specific parsing and refactoring code is not affected by the choice or power of the analysis.
+    This design has been already used to create a prototype of the tool~\footnote{https://github.com/a-usov/ptr-tidy/tree/c44e67022ba5fa2b3a541f4e730230c1e378e46e}.
+    The prototype shows that combining the AST and IR is viable and uses the already implemented LLVM escape analysis on variable declarations.
+    It also shows how Clang library allows for easy refactoring since the AST stores all source text and line numbers, making text substitution easy.
+    Substitution of AST nodes rather than source text substitution has not been explored, however this is as documentation states that generally the AST should remain immutable.
+
+    The prototype does not currently combine the analysis with the refactoring, which will be the next logical step.
+    The source refactoring is also not fully implemented, and is currently only a proof-of-concept.
+    However, \autoref{lst:prototype} shows an example output of running the prototype on a small code block, and it shows how the core idea of the project is achievable.
+
+    \begin{listing}
+        \inputminted{c++}{code/prototype-input.cpp}
+        \inputminted{text}{code/prototype-output.txt}
+        \caption{Output from using prototype tool on a chunk of code.
+We see the analysis of the variables in the code, as well as refactoring to use smart pointers.
+Analysis and refactoring not yet combined as b should not be refactored as analysis identifies it as captured, due to not being able to work across function boundaries.}
+        \label{lst:prototype}
+    \end{listing}
+
+    \subsection{Proposed Further Work}\label{subsec:proposed-further-work}
+
+    As mentioned in \autoref{subsec:current-progress}, the next work will be on completing the prototype to fully working tool.
+    This will involve completing the code refactoring component and combining this with the analysis.
+    This initial step will allow us to use the tool for initial benchmarking, as described in \autoref{subsec:benchmarking}, and this will also likely expose unhandled edge-cases, allowing us to refine the design before moving on to completing more complex analyses, as this means we can quickly complete the minimum viable product.
+    Afterwards, the focus will be on improving the analysis, by creating an interprocedural analysis, as we envision that the existing LLVM analysis will give us poor results, as pointers are often used across function boundaries, as otherwise the programmer could have used a stack variable instead.
+
+    \subsubsection{Interprocedural Analysis}\label{subsubsec:interprocedural-analysis}
+
+    We propose using a modified version of the escape analysis algorithm by \citet{Choi1999} to identify if the proposed invariants in \autoref{subsec:pointer-analysis} are violated.
+    The original algorithm already tells us whether a pointer escapes or not, however it will not tell us what the maximum number of unique owners a pointer will have throughout the execution of the program.
+    We can reuse their idea of the connection graph and add an extra condition.
+    If we look at \autoref{fig:connection-graph-choi}, then we can see that then when we copy a pointer, a new reference node with a deferred edge is created, pointing to the original pointer.
+    If each reference node has at most one incoming edge, then we know that each pointer is copied at most once, and therefore only gives ownership once, meaning that it will have at most one unique owner throughout program execution.
+    This relies on careful modeling where using pointers as function arguments is modelled as the creation of new reference nodes, which is valid in C++ as it has pass-by-value semantics.
+
+    Such an algorithm should give us better results as it will be able to determine whether pointers which are returned from functions escape, as this is a case which the intraprocedural analysis cannot determine and has to take a conservative approach in assuming it escapes.
 
     \subsection{Shared Pointer Cycle Analysis}\label{subsec:shared-pointer-cycle-analysis}
+
+    There is an emphasis in the project that the refactored program should not differ in runtime behaviour.
+    The implementation of shared pointers in the C++ standard library however cause a possible issue to this aim.
+    First of all, shared pointers are implemented using reference counting, so there is already a performance detriment if we they are used instead of raw pointers.
+    We choose to omit performance detriments, as they are likely to be marginal, however we will benchmark if any exists.
+
+    The issue that we are concerned with are functional requirements of programs, specifically that in some cases, shared pointers will fail to free memory.
+    This occurs when shared pointers are used to create reference cycles, such as can be seen in \autoref{lst:reference-cycle}.
+    This would also go against our aim of improving memory safety.
+
+    We therefore propose to have an analysis stage for checking for reference cycles.
+    If we identify a cycle, we do not perform any refactoring and leave the code as it was.
+    To identify cycles, we create a type graph of all the user defined types, and we can use depth first traversal to identify a cycle.
+    We do this by marking visited nodes in the graph as visited, and if during the traversal an adjacent node to the current node being analysed is already marked as visited, then a cycle exists.
+
+    \begin{listing}
+        \inputminted{c++}{code/reference-cycle.cpp}
+        \caption{The Person class is used to create a reference cycle.
+To deallocate the person field in bob, the alice object must be first deallocated.
+To do so, the bob object in must be deallocated first and so on, meaning the memory will never get deallocated.}
+        \label{lst:reference-cycle}
+    \end{listing}
+
+    \subsection{Benchmarking}\label{subsec:benchmarking}
+
+    There are several different benchmarks and tests we will perform to analyse the performance and correctness of the tool.
+    Firstly, it was mentioned in \autoref{subsec:shared-pointer-cycle-analysis}, we will benchmark our refactored programs against the originals for performance, so see if our refactoring efforts cause a significant detriments.
+    If the tools is effective enough, and depending on the code base, the refactorings will only include unique pointers, which will cause no performance detriments compared to using raw pointers.
+
+    Secondly, where automated test exist for the refactored programs, we will use them to ensure that the logic of the program is unchanged after refactoring.
+
+    Thirdly, we want to analyse how powerful our analysis is in identifying all opportunities in upgrading pointers to unique pointers, as our tool has to be conservative in some of its assumptions so will not be able to identify all opportunities.
+    We can measure the amount of unique pointers compared to the number of all pointers.
+    However, this number will fluctuate massively depending on what program it is run on, such as one handling resources such as files.
+    Therefore, the method we propose is to measure the performance of our tool against human refactored test cases, to see how much worse our analysis is compared to human effort.
+
+    % TODO - reference SPEC - find more better benchmarks
+    The programs that will be used for the benchmarking will be the subset of C++ programs in the SPEC CPU benchmark suite, as they will usually be pointer heavy, as well as allowing us to measure any possible performance detriments.
 
@@
 
+    \begin{landscape}
+        \begin{figure}
+            \begin{ganttchart}{1}{14}
+                \gantttitle{Work Plan}{14} \ganttnewline
+                \gantttitlelist{1,...,14}{1} \ganttnewline
+                \ganttgroup{Complete Prototype}{1}{3} \ganttnewline
+                \ganttbar{Finish Rewriting Component}{1}{1} \ganttnewline
+                \ganttbar{Combine all components into final prototype}{2}{2} \ganttnewline
+                \ganttbar{Evaluate final prototype and find edge-cases}{3}{3} \ganttnewline
+                \ganttgroup{Benchmark and Data Acquisition}{2}{3}  \ganttnewline
+                \ganttbar{Find, setup and test benchmark suite}{2}{3} \ganttnewline
+                \ganttgroup{Analysis Algorithm Exploration}{4}{7} \ganttnewline
+                \ganttbar{Implement proposed interprocedural analysis}{4}{5}\ganttnewline
+                \ganttlink{elem3}{elem7}
+                \ganttbar{Evaluate tool with interprocedural analysis}{6}{6} \ganttnewline
+                \ganttbar{Implement cycle analysis and evaluate}{7}{7} \ganttnewline
+                \ganttgroup{Result Consolidation}{8}{14} \ganttnewline
+                \ganttbar{Research Paper Writeup}{8}{14} \ganttnewline
+            \end{ganttchart}
+            \caption{[A Gantt chart demonstrating the project schedule, broken down over the first
+            14 weeks of the year from Monday 11th January until Friday 16th April.
+            ]}
+            \label{fig:gantt}
+        \end{figure}
+    \end{landscape}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-12-04 code/prototype-input.cpp

@@ -1,0 +1,10 @@
+int test(int *b) {
+    return 2;
+}
+
+int main() {
+    int a = 9;
+    int b = 2;
+    int *c = &b;
+    test(&b);
+}

### 2020-12-04 code/reference-cycle.cpp

@@ -1,0 +1,10 @@
+struct Person {
+    std::shared_ptr<Person> partner;
+};
+
+int main(){
+    std::shared_ptr<Person> alice = std::make_shared<Person>();
+    std::shared_ptr<Person> bob = std::make_shared<Person>();
+    alice->partner = bob;
+    bob->partner = alice;
+}

### 2020-12-12 proposal.tex
add risk assessment & more background, fix review comments

@@
     \title{ptr-tidy: Automatic refactoring of raw pointers in C++}
-    \author{Artem Usov}
+    \author{Artem Usov\\ 2296905U}
     \date{18 December 2020}
@@
     Memory safety errors occur when languages are used that place the tasks of memory management with the programmer, such as C and C++.
-    However simply not using these languages is not an option since their inherent danger makes them highly performant, and therefore the only option for a project such as a web browser.
+    However simply not using these languages is not an option since their unmanaged nature makes them highly performant, and therefore the best option for a system such as a web browser.
 
@@
     Rust is a systems language that offers the same performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model also guarantee memory safety.
-    The rewriting of a program in a new language is a colossal undertaking, especially in language that has a reputation for being difficult to learn.
+    The rewriting of a program in a new language is a colossal undertaking, especially in a language that has a reputation for being difficult to learn.
     We therefore propose an alternative partial solution to this problem by using a concept from Rust.
@@
     Generally pointers are used to give the program access to a resource that cannot be directly included in the program itself, such as a file or a thread.
-    However Stroustrup claims that pointers to objects allocated on the free store is dangerous and a \emph{plain old pointer}, or raw pointer as we will refer to them, should not be used to represent ownership~\cite{Stroustrup2018}.
+    However Stroustrup claims that pointers to objects allocated on the free store are dangerous and a \emph{plain old pointer}, or raw pointer as we will refer to them, should not be used to represent ownership~\cite{Stroustrup2018}.
 
@@
     In \autoref{lst:manual-leak} we can see that when using a raw pointer, there are two cases in which the programmer will never free the allocated memory.
-    Instead we can \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)} to create resource handles which automatically eliminate resource leaks with no added overhead.
-    Firstly there are \emph{unique pointers} which are a pointer that has only a single unique owner at any time.
+    Instead we can use \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)} to create resource handles which automatically eliminate resource leaks with no added overhead.
+    Firstly there are \emph{unique pointers} which are pointers that have only a single unique owner at any time.
     We can therefore fix the cases in which we leak the memory using a unique pointer which can be seen in in \autoref{lst:first-example-unique}.
-    In the comments, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will insert delete statements, now ensuring that the resource will never get leaked.
+    In the comments, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will automatically insert delete statements, now ensuring that the resource will never get leaked.
     The ownership of a unique pointer can be transferred between variables using \mintinline{c++}{std::move} which moves the ownership from one variable to another.
@@
     Reference counting is used to ensure that after the last owner loses access to the resource, it will finally be freed.
-    Shared pointers are suitable to be used in most situations, except in structures which have cyclic references \autoref{subsec:shared-pointer-cycle-analysis}.
+    Shared pointers are suitable to be used in most situations, except in structures which have cyclic references \autoref{subsec:shared-pointer-cycle-analysis} and such reference counted pointers are used by default for all values in Python and Swift.
 
@@
 
+    % TODO find refernce of how much old code is in production
+
     We therefore propose to create an automatic refactoring tool \emph{(ptr-tidy)}, that statically analyses C++ code and aims to refactor usage of raw pointers into smart pointers.
     It will involve a conservative analysis that first and foremost tries to identify raw pointers that only have a single unique owner at any point in the program.
+    These will be refactored into the previously mentioned unique pointers.
     If this cannot be guaranteed, then a more conservative refactoring to a shared pointer will be applied.
+    Given that the language standards are mostly backwards-compatible~\cite{Winters2020}, we can apply these automatic refactoring without changing the runtime behaviour compared to the original programs.
 
-    Given that the language standards are mostly backwards-compatible~\cite{Winters2020}, we can apply these automatic refactoring without changing the runtime behaviour compared to the original programs.
+    We use the term refactoring as this is one most familiar to programmers to convey a change in the code that does not change the behaviour, but improves the quality or maintainability of the code.
+    However, our work aligns more closely with the term \emph{source code rejuvenation}, as defined by \citet{Pirkelbauer2010}.
+    This is because we exactly aim to replace outdated coding patterns with newer, higher-level abstractions, as well fitting the notion that this tool would be applied once to a codebase, rather than be a reoccurring task.
+    The two terms can be used interchangeably in the context of this paper, however we will exclusively use the term refactoring.
 
@@
 
-    \emph{"Can the translation of raw pointers into smart pointers using escape analysis be used to make existing C++ code more modern and extendable?"}
+    \emph{"Can the translation of raw pointers into smart pointers be used to make existing C++ code more modern and extendable?"}
 
-    \emph{"Can the tool identify situations in which unique pointers should be used reliably?"}
+    \emph{"Can the tool reliably identify situations in which unique pointers should be used?"}
 
@@
     The most capable and mature C++ parsers will naturally be the parsers used by the most popular compilers.
-    This leaves us therefore with a choice between GCC~\footnote{https://gcc.gnu.org/} and Clang~\footnote{https://clang.llvm.org/}.
-    The literature~\cite{duffy2014} is fairly clear on the fact that the design of GCC makes it unsuitable for integration with other projects.
-    Clang on the other hand allows us access to the Clang abstract syntax tree (AST) and parser internals through consumers and visitors.
+    This leaves us therefore with a choice between GCC~\footnote{\url{https://gcc.gnu.org/}} and Clang~\footnote{\url{https://clang.llvm.org/}}.
+    The literature~\cite{Duffy2014} is fairly clear on the fact that the design of GCC makes it unsuitable for integration with other projects.
+    Clang on the other hand allows us access to the Clang abstract syntax tree (AST) and parser internals through standard design patterns such as consumers and visitors.
 
     Clang is a front-end to the LLVM framework~\cite{Lattner2004} which transforms C++ code into the LLVM intermediate representation (IR).
-    % TODO define SSA
-    The LLVM IR is a language-independent, static single assignment (SSA) representation of the program and will be the target on which we apply our analysis to identify refactoring opportunities.
+    The LLVM IR is a language-independent, static single assignment (SSA)~\cite{Rosen1988} representation of the program and will be the target on which we apply our analysis to identify refactoring opportunities.
     We believe this choice will lead to several benefits in terms of the analysis power and future usage of the project:
@@
         \item First, the usage of SSA allows us to efficiently generate a definition-usage graph, therefore allowing an efficient analysis of all the usages of a pointer to identify its number of owners.
-        \item Secondly, by separating the parsing and analysis, we make our analysis stage reusable by any other LLVM frontends.
+        \item Secondly, by separating the parsing and analysis, we make our analysis stage reusable by any other LLVM frontends as the analysis is source language agnostic.
     \end{itemize}
@@
         \centerline{\includegraphics{images/code-to-ast}}
-        \caption{Example of converting C++ code to Clang AST. All diagrams show equivalent representation of a small C++ code snippet}
+        \caption{Example of converting C++ code to Clang AST. All diagrams show equivalent representation of a small C++ code snippet.}
         \label{lst:code-to-ast-example}
@@
 
-    The escape analysis algorithm described by \citet{Choi1999} is used as the default escape analysis algorithm in the Oracle Java Virtual Machine implementation~\footnote{https://docs.oracle.com/en/java/javase/15/vm/java-hotspot-virtual-machine-performance-enhancements.html} and is widely cited so forms the basis of our investigation.
+    The escape analysis algorithm described by \citet{Choi1999} is used as the default escape analysis algorithm in the Oracle Java Virtual Machine implementation~\footnote{\url{https://docs.oracle.com/en/java/javase/15/vm/java-hotspot-virtual-machine-performance-enhancements.html}} and is widely cited so forms the basis of our investigation.
     Due to how similar C++ and Java are in terms of their Object-Oriented features, will analyse their algorithm as well as the implementations of escape analysis in the LLVM and Go compilers.
@@
     Go is another language that uses a escape analysis to allocate objects on the stack rather than the heap.
-    Their escape analysis algorithms is not defined in the specification of the language, however we can look at the source~\footnote{https://github.com/golang/go/blob/master/src/cmd/compile/internal/gc/escape.go} to determine what algorithm they use.
+    Their escape analysis algorithms is not defined in the specification of the language, however we can look at the source~\footnote{\url{https://github.com/golang/go/blob/master/src/cmd/compile/internal/gc/escape.go}} to determine what algorithm they use.
     Fortunately the comments give a summary of their algorithm, which involves constructing a weighted graph where vertices represent variables, and edges represent assignments between variables.
@@
     Finally, LLVM also uses escape analysis to remove any unnecessary heap allocations.
-    The escape analysis algorithm is also not formally defined, however we can also look at the source code~\footnote{https://github.com/llvm/llvm-project/blob/master/llvm/lib/Analysis/CaptureTracking.cpp}.
+    The escape analysis algorithm is also not formally defined, however we can also look at the source code~\footnote{\url{https://github.com/llvm/llvm-project/blob/master/llvm/lib/Analysis/CaptureTracking.cpp}}.
     We see that the \mintinline{c++}{PointerMayBeCaptured} function uses the fact that the single static assignment (SSA) in the LLVM IR allows for efficient creation of definition-usage chains.
@@
     Ada or its subset language SPARK are used extensively in certain industries to create verifiable programs for critical systems such as in avionics.
-    It is perhaps most well-know for being used in the F-22 fighter aircraft~\cite{malas1997f}, however the use of Ada for those systems show its other weakness in productivity time as it was the cause for several delays in the project, and so it largely remains unused in areas which do not call for such meticulous safety.
+    It is perhaps most well-know for being used in the F-22 fighter aircraft~\cite{Malas1997}, however the use of Ada for those systems show its other weakness in productivity time as it was the cause for several delays in the project, and so it largely remains unused in areas which do not call for such meticulous safety.
 
-    Rust has been the most language in attractive widespread adoption as a systems language.
+    Rust has been the most popular language in attractive widespread adoption as a systems language.
     Its popularity could be quantified by considering that it is being considered for usage in writing Linux kernel modules~\footnote{\url{https://lore.kernel.org/lkml/CAKwvOdmuYc8rW_H4aQG4DsJzho=F+djd68fp7mzmBp3-wY--Uw@mail.gmail.com/T}}, which cannot be said for C++.
@@
     Garbage collected languages encompass some of the most popular programming languages such as Java and Python.
-    They are considered to be easier languages than C++ since they remove the need for manual memory management by the programmer.
+    They are considered to be simpler languages than C++ since they remove the need for manual memory management by the programmer.
     In C++, if memory is improperly freed, it gets leaked and can eventually cause out of memory errors.
@@
 
-    Garbage collection does also exist for C++ such as the Boehm-Demers-Weiser conservative garbage collector~\cite{boehm2002garbage}.
+    Garbage collection does also exist for C++ such as the Boehm-Demers-Weiser conservative garbage collector~\cite{Boehm2002}.
     So why is garbage collection not used in every language?
@@
     This has led to garbage collected languages being unsuitable for use in systems programs such as OS's, as for example it would be undesirable to have seemingly random latency when interacting with the system.
-    There has been research in modern garbage collection algorithms such as ZGC~\cite{ZGC} and Shenandoah~\cite{flood2016shenandoah} for Java which aim to reduce the number and duration of pauses where the garbage collector runs, increasing the responsiveness of the Java virtual machine.
-    However, even with much reduced duration of pauses, it may still be unsuitable for some systems with strict latency requirements such as OS's or financial systems.
+    There has been research in modern garbage collection algorithms such as ZGC~\cite{Liden2018} and Shenandoah~\cite{Flood2016} for Java which aim to reduce the number and duration of pauses where the garbage collector runs, increasing the responsiveness of the Java virtual machine.
+    However, even with much reduced duration of pauses, it may still be unsuitable for some systems with strict latency requirements such as OS's or any other real time systems.
+
+    \subsection{Existing Tools and Research}\label{subsec:existing-tools-and-research}
+
+    As we have mentioned in \autoref{subsec:do-smart-pointer-solve-memory-management?}, our work closely aligns with the research done around \emph{source code rejuvenation}.
+    We believe research in this area is going to increase, as languages move and evolve much quicker than can be supported in enterprise environments, which is supported by the fact that other relevant work has been completed on this subject.
+
+    First and foremost, we have evidence that this is a problem that actually exists in industry through a paper by \citet{Wright2013} which describes a tool used in Google for running such rejuvenations across a large codebase and shows an example of running a simple rejuvenation of upgrading an API call.
+    It only uses Clang for the parsing of code into an AST, and traverses the tree to perform simple refactorings.
+    This allows for less complex refactorings, however it allows for a flexible design to implement different refactorings using the same tool, rather than the focus we have for a single specific refactoring goal.
+
+    Secondly, there has been separate academic research into rejuvenation tools such as the work by \citet{Huck2016} and \citet{Kumar2012}.
+    The former also uses the Clang and AST driven approach that we have described in \autoref{subsec:clang-and-llvm} and by \citet{Wright2013}, whilst the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
+    We have not evaluated IPR as a representation to use instead of the Clang AST, but we can however comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation that Clang has, as well as it being the choice of several other successful projects leads us to believe that we will not lose any tool efficiency in our use of it.
 
@@
     Secondly this also enables us to flexibly test different algorithms as the C++ specific parsing and refactoring code is not affected by the choice or power of the analysis.
-    This design has been already used to create a prototype of the tool~\footnote{https://github.com/a-usov/ptr-tidy/tree/c44e67022ba5fa2b3a541f4e730230c1e378e46e}.
+    This design has been already used to create a prototype of the tool~\footnote{\url{https://github.com/a-usov/ptr-tidy/tree/c44e67022ba5fa2b3a541f4e730230c1e378e46e}}.
     The prototype shows that combining the AST and IR is viable and uses the already implemented LLVM escape analysis on variable declarations.
@@
 
-    Show how you plan to organize your work, identifying intermediate deliverables and dates.
+    \subsection{Risk Analysis}\label{subsec:risk-analysis}
+
+    We have performed a risk analysis for the project, our chosen implementation as well as the work plan we envision for the completion of this project.
+    \begin{itemize}
+        \item First, the frameworks we have chosen to use, namely LLVM and Clang are mature and enterprise-backed open-source frameworks, and therefore we see no risk in using them for current and any future development,
+        \item Second, the work plan has been structured in such a way that given any unforeseen setbacks in the project, our existing prototype can be finished without the integration of more complex analyses to provide a minimum viable product which can be used in experimentation and evaluation.
+        This ensures that we can answer our research questions for this project.
+    \end{itemize}
+
+    % TODO - finish this
+
+    \subsection{Gantt Chart}\label{subsec:gantt-chart}
 
@@
             \end{ganttchart}
-            \caption{[A Gantt chart demonstrating the project schedule, broken down over the first
+            \caption{A Gantt chart demonstrating the project schedule, broken down over the first
             14 weeks of the year from Monday 11th January until Friday 16th April.
-            ]}
+            }
             \label{fig:gantt}

### 2020-12-14 proposal.tex
Fix rotation of workplan and leaving extra empty page
Full review and edit of proposal

@@
 \usepackage{pgfgantt}
-\usepackage{lscape}
+\usepackage{rotating}
 \usepackage{hyperref}
@@
 
-    % TODO insert info about what memory management/errors are
+    Systems programming often involves memory management, that is requesting memory for the program to use by the operating system.
+    This is done when the amount of memory that we need cannot be known at the time of compilation of the program.
+    However doing so creates the opportunity for a memory safety error~\cite{Dhurjati2003} and this means that correct memory management is a challenging problem.
 
-    Correct memory management is a difficult problem.
+    \emph{How challenging one might ask?}
 
-    \emph{How difficult one might ask?}
-
-    In a presentation by Matt Miller, a security engineer at Microsoft~\cite{Miller2019}, it is shown that around 70\% of vulnerabilities that are addressed through a security updates are due to memory safety issues.
-    In another presentation at the Linux Security Summit~\cite{Gaynor2020} it is shown that in several other top projects such as Firefox, macOS, Ubuntu and Android all had over half of their CVEs attributed to issues with memory safety.
+    In a presentation by Matt Miller, a security engineer at Microsoft, it is shown that around 70\% of their vulnerabilities that are addressed through a security updates are due to memory safety issues~\cite{Miller2019}.
+    In another presentation at the Linux Security Summit it is shown that in several other top projects such as Firefox, macOS, Ubuntu and Android all had over half of their CVEs~\footnote{\url{https://cve.mitre.org/}} attributed to issues with memory safety~\cite{Gaynor2020}.
     This is at industry-leading companies who are renowned for hiring top talent.
@@
 
-    Memory safety errors occur when languages are used that place the tasks of memory management with the programmer, such as C and C++.
+    Memory safety errors occur in languages that place the task of memory management with the programmer, such as C and C++.
     However simply not using these languages is not an option since their unmanaged nature makes them highly performant, and therefore the best option for a system such as a web browser.
@@
     Rust is a systems language that offers the same performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model also guarantee memory safety.
-    The rewriting of a program in a new language is a colossal undertaking, especially in a language that has a reputation for being difficult to learn.
-    We therefore propose an alternative partial solution to this problem by using a concept from Rust.
+    The rewriting of a program in a new language, namely from C++ to Rust, is a colossal undertaking, especially given that Rust has a reputation for being difficult to learn.
+    We therefore propose an alternative partial solution to this problem.
 
@@
 
-    Generally pointers are used to give the program access to a resource that cannot be directly included in the program itself, such as a file or a thread.
+    Generally pointers are used to give the program access to a resource that cannot be directly included in the program itself, such as allocated memory or a file.
     However Stroustrup claims that pointers to objects allocated on the free store are dangerous and a \emph{plain old pointer}, or raw pointer as we will refer to them, should not be used to represent ownership~\cite{Stroustrup2018}.
@@
         \inputminted{c++}{code/manual-leak.cpp}
-        \caption{Example of memory leaks using manual management.}
+        \caption{Example of memory leaks using manual management.
+Lines 4 and 5 show how the memory allocated on line 2 will never be freed since the control flow of the function is interrupted before we execute line 7.}
         \label{lst:manual-leak}
@@
     In \autoref{lst:manual-leak} we can see that when using a raw pointer, there are two cases in which the programmer will never free the allocated memory.
-    Instead we can use \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)} to create resource handles which automatically eliminate resource leaks with no added overhead.
-    Firstly there are \emph{unique pointers} which are pointers that have only a single unique owner at any time.
-    We can therefore fix the cases in which we leak the memory using a unique pointer which can be seen in in \autoref{lst:first-example-unique}.
-    In the comments, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will automatically insert delete statements, now ensuring that the resource will never get leaked.
-    The ownership of a unique pointer can be transferred between variables using \mintinline{c++}{std::move} which moves the ownership from one variable to another.
+    Instead we can use \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create resource handles which automatically eliminate resource leaks with no added overhead.
+    The first type of smart pointer are \emph{unique pointers} which have only a single unique owner at any time.
+    We can therefore fix the previous example in which we leak memory using a unique pointer, as can be seen in \autoref{lst:first-example-unique}.
+    In the comments of \autoref{lst:first-example-unique}, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will automatically insert delete statements, now ensuring that the resource will never get leaked.
+    The ownership of a unique pointer can be transferred between variables using \mintinline{c++}{std::move} which moves the ownership from one variable to another, making the original owner invalid.
 
@@
         \inputminted{c++}{code/first-example-unique.cpp}
-        \caption{Example of using a unique pointer to manage memory.}
+        \caption{Example of using a unique pointer to manage memory.
+Note how the programmer no longer has to specify to manually delete the pointer as in \autoref{lst:manual-leak}, and how the compiler can find all cases where the resource needs to be freed in lines 4, 5 and 7, and will automatically insert delete statements there on compilation.}
         \label{lst:first-example-unique}
@@
 
-    There also exists a \emph{shared pointer} for resources which may not have a single unique owner.
-    Reference counting is used to ensure that after the last owner loses access to the resource, it will finally be freed.
-    Shared pointers are suitable to be used in most situations, except in structures which have cyclic references \autoref{subsec:shared-pointer-cycle-analysis} and such reference counted pointers are used by default for all values in Python and Swift.
+    The second type of smart pointer are \emph{shared pointers} which are used for resources which may not necessarily have a single unique owner and so several owners \emph{share} the resource.
+    Reference counting~\cite{Collins1960} is used to ensure that after the last owner loses access to the resource, it will finally be freed.
+    Shared pointers are suitable to be used in most situations and such reference counted pointers are used by default for all values in the Python and Swift languages.
 
@@
 
-    It would appear then that smart pointers could solve most of our memory management issues, however they were only introduced in the C++11 standard.
-    The C++ dialects till then did not have any of these language features, with programs being written predominantly in the C++03 standard.
-    There remain many such so called \emph{legacy} programs which have not upgraded to newer standards, and even then, it is up to the programmer to properly utilise smart pointers in their code.
+    It would appear then that smart pointers could solve most of our memory management issues, but they were only introduced in the C++11 standard.
+    The C++ standards (\textit{versions}) till then did not have these language features, with older programs being written predominantly in the C++03 standard.
+    There remain many such called \emph{legacy} programs which have not upgraded to newer standards, and even then the programmer can choose not to utilise smart pointers in their code.
 
-    % TODO find refernce of how much old code is in production
+    \subsubsection{Research Aim}\label{subsubsec:reseach-aim}
+
+    % TODO find reference of how much old code is in production
 
@@
     If this cannot be guaranteed, then a more conservative refactoring to a shared pointer will be applied.
-    Given that the language standards are mostly backwards-compatible~\cite{Winters2020}, we can apply these automatic refactoring without changing the runtime behaviour compared to the original programs.
+    Given that the language standards are mostly backwards-compatible~\cite{Winters2020}, we aim to apply automatic refactoring without changing the runtime behaviour compared to the original programs.
 
-    We use the term refactoring as this is one most familiar to programmers to convey a change in the code that does not change the behaviour, but improves the quality or maintainability of the code.
+    We use the term refactoring as this is one most familiar to programmers to convey a change in code that does not change its behaviour, but improves its the quality or maintainability.
     However, our work aligns more closely with the term \emph{source code rejuvenation}, as defined by \citet{Pirkelbauer2010}.
-    This is because we exactly aim to replace outdated coding patterns with newer, higher-level abstractions, as well fitting the notion that this tool would be applied once to a codebase, rather than be a reoccurring task.
-    The two terms can be used interchangeably in the context of this paper, however we will exclusively use the term refactoring.
+    This is because we exactly aim to replace outdated coding patterns with newer, higher-level abstractions, as well as fitting the notion that this tool would be applied once to a codebase, rather than be a reoccurring task.
+    The two terms can be used interchangeably in the context of this paper, however we will exclusively use the term refactoring for simplicity.
 
@@
 
-    \subsection{The Research Questions}\label{subsec:the-research-questions}
+    \subsection{Research Questions}\label{subsec:research-questions}
 
-    \emph{"Can the translation of raw pointers into smart pointers be used to make existing C++ code more modern and extendable?"}
+    We outline below the research questions we would like so answer alongside developing the tool.
+    Namely we would like answer the question of whether such a tool can truly improve C++ code by seeing if the tool can be used outside small examples on larger existing programs, as well as if the refactored results produce code which can be compiled and can be judged to be better.
+    Secondly, if the tool is able to do this, we want to evaluate how well the tool performs by analysing the performance of the analysis algorithms.
 
-    \emph{"Can the tool reliably identify situations in which unique pointers should be used?"}
+    \emph{RQ1: Can the translation of raw pointers into smart pointers be used to make existing C++ code more modern and extendable?}
+
+    \emph{RQ2: Can the tool reliably identify situations in which unique pointers should be used?}
 
@@
 
-    In this section we will look into the frameworks and algorithms that we will use to enable us to read and refactor C++ code, and secondly identify the locations where we can apply our refactorings.
+    In this section we will look into frameworks, design patterns and algorithms that we will use to enable us to read and refactor C++ code and to identify the locations where we can apply our refactorings.
+    We also look into existing memory safe languages, garbage collection as a solution for automated memory management as well as related existing research.
 
@@
 
-    C++ as a language is massive, and the first capability our tool will need is to be able to parse C++ code.
+    C++ as a language is enormous, and the first capability our tool will need is to be able to parse (\textit{read}) C++ code.
     Doing this ourselves would be an undertaking that would likely span longer than the time span of the whole project, therefore it was decided early to use an existing parser.
@@
 
-    Clang is a front-end to the LLVM framework~\cite{Lattner2004} which transforms C++ code into the LLVM intermediate representation (IR).
+    Clang is a front-end to the LLVM framework~\cite{Lattner2004} and transforms C++ code into the LLVM intermediate representation (IR).
     The LLVM IR is a language-independent, static single assignment (SSA)~\cite{Rosen1988} representation of the program and will be the target on which we apply our analysis to identify refactoring opportunities.
@@
         \centerline{\includegraphics{images/code-to-ast}}
-        \caption{Example of converting C++ code to Clang AST. All diagrams show equivalent representation of a small C++ code snippet.}
+        \caption{Example of converting C++ code to Clang AST.
+All diagrams show equivalent representation of a small C++ code snippet.
+The textual representation of the AST shows how information rich each node in the diagrammatic representation is.}
         \label{lst:code-to-ast-example}
@@
 
-    The main capability of the tool is to be able to identify what the maximum number of owners a pointer has at all points of the program execution, for all pointers in any given program.
+    The main capability of the tool is to be able to identify what the maximum number of owners a pointer has at all points of the program execution for all pointers in any given program.
     If this number for a given raw pointer is one, then we can go about converting all usages of that raw pointer into a unique pointer.
@@
     \begin{itemize}
-        \item First, a pointer escapes, as then we do not know where or how the pointer may be used, so we must take a conservative approach and assume the pointer no longer has at most a single unique owner.
-        \item Secondly, if a pointer is captured at most once, as we can model a single capture as a move of the unique pointer.
+        \item If a pointer escapes, then we do not know where or how the pointer may be used, so we must take a conservative approach and assume the pointer no longer has at most a single unique owner.
+        \item If a pointer is does not escape but is captured more than once, then it cannot be a unique pointer.
+        The case where a pointer is captured exactly once is a special case that we can model as a move of a unique pointer.
     \end{itemize}
@@
 
-    Given how we have defined how escape analysis can be used to identify areas of refactoring, we now need to identify a suitable algorithm.
+    Now that we have defined how escape analysis can be used to identify areas of refactoring, we now need to identify a suitable algorithm.
 
-    There has been extensive research on escape analysis for Java, where escape analysis is used to determine whether a object escapes the current method, as then if it can be proven that the object does not escape the method, then it is local to the method invocation and it can be allocated on the stack rather than as a heap variable (the default for all Java objects).
+    There has been extensive research on escape analysis for Java, where escape analysis is used to determine whether a object escapes the current method.
+    If it can be proven that the object does not escape the method, then it is local to the method invocation and it can be allocated on the stack rather than as a heap variable (the default for all Java objects).
 
-    The escape analysis algorithm described by \citet{Choi1999} is used as the default escape analysis algorithm in the Oracle Java Virtual Machine implementation~\footnote{\url{https://docs.oracle.com/en/java/javase/15/vm/java-hotspot-virtual-machine-performance-enhancements.html}} and is widely cited so forms the basis of our investigation.
-    Due to how similar C++ and Java are in terms of their Object-Oriented features, will analyse their algorithm as well as the implementations of escape analysis in the LLVM and Go compilers.
+    The escape analysis algorithm described by \citet{Choi1999} is used as the default escape analysis algorithm in the Oracle Java Virtual Machine implementation~\footnote{\url{https://docs.oracle.com/en/java/javase/15/vm/java-hotspot-virtual-machine-performance-enhancements.html}} and it is widely cited so it forms the basis of our investigation.
+    Due to how similar Java is to C++ and Go in terms of their object-oriented features, it means we can also analyse the escape algorithms used in the Clang and Go compilers.
 
-    \citet{Choi1999} introduce the abstraction of a \emph{connection graph} (CG) which we will see is also used in other implementations.
+    \citet{Choi1999} introduce the abstraction of a \emph{connection graph} which we will see is also used in other implementations.
     The connection graph captures relationships between heap-allocated objects and object references.
     The graph is used to perform reachability analysis to determine if an is local to a thread or method.
-    Most importantly, this abstraction allows for a more powerful interprocedural analysis to be performed.
+    Most importantly, this abstraction allows for a powerful interprocedural analysis to be performed.
 
@@
 
-    For each node, we associate an escape set, of either \emph{NoEscape}, \emph{ArgEscape} or \emph{GlobalEscape}.
+    For each node, we associate a state of either \emph{NoEscape}, \emph{ArgEscape} or \emph{GlobalEscape}.
     \emph{NoEscape} means that the object does not escape the method in which it was created
@@
 
-    The escape analysis is executed through their defined construction of intraprocedural graphs for each method call.
-    We then proceed with the interprocedural by using the CG of the callee to update the CG of the caller.
+    The escape analysis is executed through the construction of intraprocedural connection graphs for each method call.
+    We then proceed with creating the interprocedural graph by using the connection graph of the callee to update the connection graph of the caller.
 
@@
     Go is another language that uses a escape analysis to allocate objects on the stack rather than the heap.
-    Their escape analysis algorithms is not defined in the specification of the language, however we can look at the source~\footnote{\url{https://github.com/golang/go/blob/master/src/cmd/compile/internal/gc/escape.go}} to determine what algorithm they use.
+    Their escape analysis algorithms is not defined in the specification of the language, however we can look at the source code~\footnote{\url{https://github.com/golang/go/blob/master/src/cmd/compile/internal/gc/escape.go}} to determine what algorithm they use.
     Fortunately the comments give a summary of their algorithm, which involves constructing a weighted graph where vertices represent variables, and edges represent assignments between variables.
@@
 
-    There have been many proposed systems languages such as Rust, Swift~\footnote{\url{https://developer.apple.com/swift/}}, D~\footnote{\url{https://dlang.org/}}, Nim~\footnote{\url{https://nim-lang.org/}} or Ada (SPARK)~\footnote{\url{https://adacore.com/}}, however none of these have arguably become mainstream, with Rust and Ada coming closest.
+    There have been many proposed systems languages such as Rust~\cite{Balasubramanian2017}, Swift~\footnote{\url{https://developer.apple.com/swift/}}, D~\footnote{\url{https://dlang.org/}}, Nim~\footnote{\url{https://nim-lang.org/}} or Ada (SPARK)~\footnote{\url{https://adacore.com/}}.
+    Yet none of these have arguably become mainstream alternatives to C and C++, with Rust and Ada coming closest.
 
-    Swift remains mostly a language for the Apple MacOS and iOS environments for application creation, however it has seen some usage for system programs by companies such as IBM~\footnote{\url{https://github.com/ibm-swift}}.
+    Swift remains mostly a language for application creation in the the Apple MacOS and iOS environments, however it has seen some usage for system programs by companies such as IBM~\footnote{\url{https://github.com/ibm-swift}}.
     Ada or its subset language SPARK are used extensively in certain industries to create verifiable programs for critical systems such as in avionics.
@@
     Rust affine types~\cite{Balasubramanian2017} are semantically equivalent to C++ unique pointers, in that all values have a single unique owner at any point.
-    Therefore our work can identify which pointers are unique and which are not, and those that are not cannot be transpiled into safe Rust, instead requiring to use unsafe blocks to represent such values.
+    Therefore our work can identify which pointers are unique and which are not, and those that are not cannot be transpiled into idiomatic Rust.
 
@@
     Garbage collection is not zero-cost as the garbage collection algorithm typically stops the execution of the running program to clean up the memory.
-    This means that latency or pauses is introduced to every program, and can occur at any point in the execution of the program, such as during the hot path of the program.
-    This has led to garbage collected languages being unsuitable for use in systems programs such as OS's, as for example it would be undesirable to have seemingly random latency when interacting with the system.
+    This means that latency or pauses are introduced to every program, and can occur at any point in the execution of the program, such as during the hot path of the program.
+
+    This has led to garbage collected languages being unsuitable for use in real time systems such as operating systems, as it would be undesirable to have seemingly random latency when interacting with the system.
     There has been research in modern garbage collection algorithms such as ZGC~\cite{Liden2018} and Shenandoah~\cite{Flood2016} for Java which aim to reduce the number and duration of pauses where the garbage collector runs, increasing the responsiveness of the Java virtual machine.
-    However, even with much reduced duration of pauses, it may still be unsuitable for some systems with strict latency requirements such as OS's or any other real time systems.
+    However, even with much reduced duration of pauses, it may still be unsuitable for some systems with strict latency requirements.
 
@@
 
-    As we have mentioned in \autoref{subsec:do-smart-pointer-solve-memory-management?}, our work closely aligns with the research done around \emph{source code rejuvenation}.
+    As mentioned in \autoref{subsec:do-smart-pointer-solve-memory-management?}, our work closely aligns with the research done around \emph{source code rejuvenation}.
     We believe research in this area is going to increase, as languages move and evolve much quicker than can be supported in enterprise environments, which is supported by the fact that other relevant work has been completed on this subject.
@@
     First and foremost, we have evidence that this is a problem that actually exists in industry through a paper by \citet{Wright2013} which describes a tool used in Google for running such rejuvenations across a large codebase and shows an example of running a simple rejuvenation of upgrading an API call.
-    It only uses Clang for the parsing of code into an AST, and traverses the tree to perform simple refactorings.
-    This allows for less complex refactorings, however it allows for a flexible design to implement different refactorings using the same tool, rather than the focus we have for a single specific refactoring goal.
+    They only uses Clang for the parsing of code into an AST and traversing the tree to identify and perform simple refactorings.
+    This allows for less complex refactorings, however it also allows for a flexible design to implement different refactorings using the same tool, rather than the focus we have for a single specific refactoring goal.
 
@@
     The former also uses the Clang and AST driven approach that we have described in \autoref{subsec:clang-and-llvm} and by \citet{Wright2013}, whilst the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
-    We have not evaluated IPR as a representation to use instead of the Clang AST, but we can however comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation that Clang has, as well as it being the choice of several other successful projects leads us to believe that we will not lose any tool efficiency in our use of it.
+    We have not evaluated IPR as a representation to use instead of the Clang AST, but we can however comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation that Clang has, as well as it being the choice of several other successful projects leads us to believe that we will not lose any effectiveness in our use of it.
 
@@
     As mentioned in \autoref{subsec:clang-and-llvm}, we plan on using Clang for parsing C++ code into an AST and LLVM IR for the analysis.
-    We can employ the visitor pattern to traverse the AST and look for nodes of interest, such as variable declaration nodes \emph{(varDecl)}.
+    We can employ the visitor pattern to traverse the AST and look for nodes of interest, such as variable declaration nodes.
     Given a node in the AST, we can find its equivalent declaration in the IR\@.
-    This is mainly possible since we can see from line 7 in \autoref{lst:code-to-ast-ir} that the variable names can matched from to AST to the IR\@.
+    This is mainly possible since we can see from comparing \autoref{lst:code-to-ast-example} and \autoref{lst:code-to-ast-ir} that the variable names can matched from the AST to the IR\@.
 
@@
 
-    We can then apply escape analysis for this particular value, which will give us a boolean \mintinline{c++}{true} or \mintinline{c++}{false} result.
-    Based on the returned value, we know if the value does not escape and so we can refactor all the usages of this value.
+    We can then apply escape analysis for this particular value declaration, which will give us a boolean \mintinline{c++}{true} or \mintinline{c++}{false} result.
+    Based on the returned value, we know if the value does not escape and if so, we can refactor all the uses of this value.
     This will involve changing the type of the initial declaration, its definition, as well as removing all deletions of the pointer, which can be seen in \autoref{lst:manual-leak} and \autoref{lst:first-example-unique}.
-    This is repeated for all usages of the value as well, such as being used as a formal argument.
     The diagram in \autoref{fig:design} shows a visual representation of this proposed design.
@@
 
-    This design mirrors the modular design of LLVM where because the analysis is performed on the IR, it could be reused by the frontend of a different language and not necessarily Clang as we do.
+    This design is largely inspired by the componentised design of LLVM~\cite{Lattner2004} and brings several advantages.
+    Firstly the analysis is performed on the IR and so it could be reused by the frontend of a different language and not just necessarily Clang.
     Secondly this also enables us to flexibly test different algorithms as the C++ specific parsing and refactoring code is not affected by the choice or power of the analysis.
+
     This design has been already used to create a prototype of the tool~\footnote{\url{https://github.com/a-usov/ptr-tidy/tree/c44e67022ba5fa2b3a541f4e730230c1e378e46e}}.
     The prototype shows that combining the AST and IR is viable and uses the already implemented LLVM escape analysis on variable declarations.
-    It also shows how Clang library allows for easy refactoring since the AST stores all source text and line numbers, making text substitution easy.
+    It also shows how Clang library allows for easy refactoring since the AST stores all source text and line numbers, making text substitution trivial.
     Substitution of AST nodes rather than source text substitution has not been explored, however this is as documentation states that generally the AST should remain immutable.
@@
         \inputminted{text}{code/prototype-output.txt}
-        \caption{Output from using prototype tool on a chunk of code.
-We see the analysis of the variables in the code, as well as refactoring to use smart pointers.
-Analysis and refactoring not yet combined as b should not be refactored as analysis identifies it as captured, due to not being able to work across function boundaries.}
+        \caption{Output from using prototype tool on a small program.
+We see the analysis of the variables in the code, as well as source code refactoring to use smart pointers.
+Analysis and refactoring has not yet combined as $b$ should not be refactored since the analysis identifies it as being captured, due to not being able to analyse across function boundaries.}
         \label{lst:prototype}
@@
 
-    As mentioned in \autoref{subsec:current-progress}, the next work will be on completing the prototype to fully working tool.
+    As mentioned in \autoref{subsec:current-progress}, further work needs to be completed to create a fully capable first version of the tool.
     This will involve completing the code refactoring component and combining this with the analysis.
-    This initial step will allow us to use the tool for initial benchmarking, as described in \autoref{subsec:benchmarking}, and this will also likely expose unhandled edge-cases, allowing us to refine the design before moving on to completing more complex analyses, as this means we can quickly complete the minimum viable product.
-    Afterwards, the focus will be on improving the analysis, by creating an interprocedural analysis, as we envision that the existing LLVM analysis will give us poor results, as pointers are often used across function boundaries, as otherwise the programmer could have used a stack variable instead.
+    This first version will allow us to use the tool for initial benchmarking, as described in \autoref{subsec:benchmarking}, and this will also likely expose unhandled edge-cases, allowing us to refine the design before moving on to completing more complex analyses.
+    This means we can quickly achieve a minimum viable product.
+
+    Afterwards, the focus will be on improving the analysis by implementing an interprocedural analysis, as we envision that the existing LLVM analysis will give us poor results.
+    This is because pointers are more often used across function boundaries, as otherwise the programmer could have used a stack variable instead.
 
@@
     We propose using a modified version of the escape analysis algorithm by \citet{Choi1999} to identify if the proposed invariants in \autoref{subsec:pointer-analysis} are violated.
-    The original algorithm already tells us whether a pointer escapes or not, however it will not tell us what the maximum number of unique owners a pointer will have throughout the execution of the program.
+    The original algorithm already determines whether a pointer escapes or not, however it cannot determine what the maximum number of unique owners a pointer will have throughout the execution of the program.
     We can reuse their idea of the connection graph and add an extra condition.
-    If we look at \autoref{fig:connection-graph-choi}, then we can see that then when we copy a pointer, a new reference node with a deferred edge is created, pointing to the original pointer.
-    If each reference node has at most one incoming edge, then we know that each pointer is copied at most once, and therefore only gives ownership once, meaning that it will have at most one unique owner throughout program execution.
-    This relies on careful modeling where using pointers as function arguments is modelled as the creation of new reference nodes, which is valid in C++ as it has pass-by-value semantics.
+
+    In \autoref{fig:connection-graph-choi} we can see that when we copy a pointer, a new reference node with a deferred edge is created, pointing to the original pointer.
+    Therefore if each reference node has at most one incoming edge, then we know that each pointer is copied at most once, and therefore only transfers ownership once.
+    This therefore means that the value will have at most one unique throughout program execution.
+    This relies on careful modeling where using pointers as function arguments is modelled as the creation of new reference nodes, which can be correctly modelled in C++ as it has pass-by-value semantics.
 
@@
     The implementation of shared pointers in the C++ standard library however cause a possible issue to this aim.
-    First of all, shared pointers are implemented using reference counting, so there is already a performance detriment if we they are used instead of raw pointers.
+    First of all, shared pointers are implemented using reference counting~\cite{Collins1960}, so there is already a performance detriment compared to raw pointers.
     We choose to omit performance detriments, as they are likely to be marginal, however we will benchmark if any exists.
@@
     This occurs when shared pointers are used to create reference cycles, such as can be seen in \autoref{lst:reference-cycle}.
-    This would also go against our aim of improving memory safety.
+    This would go against our aim of improving memory safety.
 
@@
     Firstly, it was mentioned in \autoref{subsec:shared-pointer-cycle-analysis}, we will benchmark our refactored programs against the originals for performance, so see if our refactoring efforts cause a significant detriments.
-    If the tools is effective enough, and depending on the code base, the refactorings will only include unique pointers, which will cause no performance detriments compared to using raw pointers.
+    Depending on the benchmark program and the effectiveness of the program, the refactoring could only include new unique pointers, which would result in identically performing programs.
 
@@
 
-    Thirdly, we want to analyse how powerful our analysis is in identifying all opportunities in upgrading pointers to unique pointers, as our tool has to be conservative in some of its assumptions so will not be able to identify all opportunities.
+    Thirdly, we want to analyse how powerful our analysis is in identifying all opportunities in upgrading pointers to unique pointers.
+    Our tool has to be conservative in some of its assumptions so might not be able to identify all opportunities.
     We can measure the amount of unique pointers compared to the number of all pointers.
@@
 
+    We have performed a risk analysis for the project, our design as well as our work plan, seen in \autoref{fig:gantt}, which outlines a schedule for the completion of this project:
+
     \subsection{Risk Analysis}\label{subsec:risk-analysis}
 
-    We have performed a risk analysis for the project, our chosen implementation as well as the work plan we envision for the completion of this project.
     \begin{itemize}
-        \item First, the frameworks we have chosen to use, namely LLVM and Clang are mature and enterprise-backed open-source frameworks, and therefore we see no risk in using them for current and any future development,
-        \item Second, the work plan has been structured in such a way that given any unforeseen setbacks in the project, our existing prototype can be finished without the integration of more complex analyses to provide a minimum viable product which can be used in experimentation and evaluation.
+        \item Firstly, the frameworks we have chosen to use, namely LLVM and Clang are mature and enterprise-backed open-source frameworks, and therefore we see no risk in using them for current and any future development,
+        \item Secondly, the work plan has been structured in such a way that given any unforeseen setbacks in the project, our existing prototype can be finished without the integration of more complex analyses to provide a minimum viable product which can be used in experimentation and evaluation.
         This ensures that we can answer our research questions for this project.
@@
 
-    % TODO - finish this
+    \newpage
 
@@
 
-    \begin{landscape}
-        \begin{figure}
+    \begin{figure}[!htb]
+        \rotatebox{90}{
         \begin{ganttchart}{1}{14}
@@
             \ganttbar{Research Paper Writeup}{8}{14} \ganttnewline
-            \end{ganttchart}
+        \end{ganttchart}}
         \caption{A Gantt chart demonstrating the project schedule, broken down over the first
@@
     \end{figure}
-    \end{landscape}
 

### 2020-12-17 paper.tex
init

\documentclass{paper}

\begin{document}

    \title{ptr-tidy: Automatic refactoring of raw pointers in C++}
    \author{Artem Usov}
    \matricnum{2296905U}

    \maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{abstract}
    \end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \section{Introduction}\label{sec:introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \section{Background}\label{sec:background}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \section{Evaluation}\label{sec:evaluation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \section{Conclusions}\label{sec:conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \vskip8pt \noindent
    \textbf{Acknowledgments.}

    \bibliographystyle{abbrv}
    \bibliography{paper}

\end{document}

### 2020-12-18 proposal.tex
Fix typos, clarify sections

@@
 \usepackage{listing}
-\usepackage[labelfont=bf]{caption}
-\usepackage[outputdir=out]{minted}
+\usepackage[labelfont=bf, font=small]{caption}
+\usepackage{minted}
 \usepackage{chngcntr}
@@
 
-\setminted{frame=lines, linenos, autogobble}
+\setminted{frame=lines, linenos, autogobble, breaklines}
 
@@
 
-    Systems programming often involves memory management, that is requesting memory for the program to use by the operating system.
+    Systems programming often involves memory management, that is requesting memory from operating system to be used and managed by the program.
     This is done when the amount of memory that we need cannot be known at the time of compilation of the program.
-    However doing so creates the opportunity for a memory safety error~\cite{Dhurjati2003} and this means that correct memory management is a challenging problem.
+    However doing so creates the opportunity for memory safety errors~\cite{Dhurjati2003} which are notoriously challenging to avoid.
 
@@
     Such memory safety vulnerabilities can be exploited, and due to new regulations such as the GDPR, these issues are more commonly exposed to the general public and punished by regulators.
-    Such an example is a fine of £20m fine for a British Airways data breach by the British Information Commissioner's Office~\cite{ICO2020}.
+    An example is a fine of £20m fine for a British Airways data breach by the British Information Commissioner's Office~\cite{ICO2020}.
 
@@
     Memory safety errors occur in languages that place the task of memory management with the programmer, such as C and C++.
-    However simply not using these languages is not an option since their unmanaged nature makes them highly performant, and therefore the best option for a system such as a web browser.
+    However simply not using these languages is not an option since their unmanaged nature makes them highly performant, and therefore the best option for systems such as a web browser.
 
@@
     Engineers at both Microsoft~\cite{Thomas2019} and Mozilla~\cite{Hostfelt2019} converge on Rust~\cite{Balasubramanian2017} as a possible solution.
-    Rust is a systems language that offers the same performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model also guarantee memory safety.
+    Rust is a systems language that offers similar performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model also guarantee memory safety.
     The rewriting of a program in a new language, namely from C++ to Rust, is a colossal undertaking, especially given that Rust has a reputation for being difficult to learn.
@@
         \caption{Example of memory leaks using manual management.
-Lines 4 and 5 show how the memory allocated on line 2 will never be freed since the control flow of the function is interrupted before we execute line 7.}
+Lines 4 and 5 show how the memory allocated on line 2 will never be freed since the control flow of the function is interrupted before we reach line 7 and free the memory}
         \label{lst:manual-leak}
@@
     In \autoref{lst:manual-leak} we can see that when using a raw pointer, there are two cases in which the programmer will never free the allocated memory.
-    Instead we can use \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create resource handles which automatically eliminate resource leaks with no added overhead.
+    Instead we can use \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create resource handles which automatically eliminate resource leaks with no, or very little added overhead.
     The first type of smart pointer are \emph{unique pointers} which have only a single unique owner at any time.
@@
     In the comments of \autoref{lst:first-example-unique}, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will automatically insert delete statements, now ensuring that the resource will never get leaked.
-    The ownership of a unique pointer can be transferred between variables using \mintinline{c++}{std::move} which moves the ownership from one variable to another, making the original owner invalid.
+    The ownership of a unique pointer can be transferred between variables using \mintinline{c++}{std::move} which moves the ownership from one variable to another, making the original reference invalid.
 
@@
         \caption{Example of using a unique pointer to manage memory.
-Note how the programmer no longer has to specify to manually delete the pointer as in \autoref{lst:manual-leak}, and how the compiler can find all cases where the resource needs to be freed in lines 4, 5 and 7, and will automatically insert delete statements there on compilation.}
+Note how the programmer no longer has to specify to manually delete the pointer as in \autoref{lst:manual-leak}, and how the compiler can detect that on lines 4, 5 and 7 the variable $p$ leaves the scope and therefore needs to be freed.
+The compiler will automatically insert these delete statements during compilation.}
         \label{lst:first-example-unique}
@@
 
-    The second type of smart pointer are \emph{shared pointers} which are used for resources which may not necessarily have a single unique owner and so several owners \emph{share} the resource.
+    The second type of smart pointer are \emph{shared pointers} which are used for resources which may not necessarily have a single unique owner and so several owners \emph{share} ownership to the resource.
     Reference counting~\cite{Collins1960} is used to ensure that after the last owner loses access to the resource, it will finally be freed.
@@
     It would appear then that smart pointers could solve most of our memory management issues, but they were only introduced in the C++11 standard.
-    The C++ standards (\textit{versions}) till then did not have these language features, with older programs being written predominantly in the C++03 standard.
+    The C++ standards (\textit{versions}) before then did not have these language features, with older programs being written predominantly in the C++03 standard.
     There remain many such called \emph{legacy} programs which have not upgraded to newer standards, and even then the programmer can choose not to utilise smart pointers in their code.
@@
 
-    We therefore propose to create an automatic refactoring tool \emph{(ptr-tidy)}, that statically analyses C++ code and aims to refactor usage of raw pointers into smart pointers.
+    We therefore propose to create an automatic refactoring tool \emph{(ptr-tidy)}, that performs static analysis on C++ code and aims to refactor usage of raw pointers into smart pointers.
     It will involve a conservative analysis that first and foremost tries to identify raw pointers that only have a single unique owner at any point in the program.
-    These will be refactored into the previously mentioned unique pointers.
+    These will be refactored into unique pointers.
     If this cannot be guaranteed, then a more conservative refactoring to a shared pointer will be applied.
-    Given that the language standards are mostly backwards-compatible~\cite{Winters2020}, we aim to apply automatic refactoring without changing the runtime behaviour compared to the original programs.
+    We aim to apply automatic refactoring without changing the runtime behaviour compared to the original programs, which we believe will be possible given that C++ language standards are backwards-compatible to a high degree~\cite{Winters2020}.
 
@@
 
-    % TODO reword section title - or move to proposed approach section
-
     \subsection{Research Questions}\label{subsec:research-questions}
 
-    We outline below the research questions we would like so answer alongside developing the tool.
-    Namely we would like answer the question of whether such a tool can truly improve C++ code by seeing if the tool can be used outside small examples on larger existing programs, as well as if the refactored results produce code which can be compiled and can be judged to be better.
-    Secondly, if the tool is able to do this, we want to evaluate how well the tool performs by analysing the performance of the analysis algorithms.
+    We outline below the research questions we would like to answer alongside developing the tool.
+    Namely we would like to answer the question of whether such a tool can truly improve C++ code by seeing if the tool can be used outside small examples on larger existing programs, as well as if the refactored results produce code which can be compiled and can be judged to be better.
+    Secondly, if the tool is able to do this, we want to evaluate how well the tool performs by analysing the effectiveness of the analysis algorithms.
 
-    \emph{RQ1: Can the translation of raw pointers into smart pointers be used to make existing C++ code more modern and extendable?}
+    \emph{RQ1: Can the translation of raw pointers into smart pointers be used to make existing C++ code more modern, safe and extendable?}
 
@@
     C++ as a language is enormous, and the first capability our tool will need is to be able to parse (\textit{read}) C++ code.
-    Doing this ourselves would be an undertaking that would likely span longer than the time span of the whole project, therefore it was decided early to use an existing parser.
+    Doing this ourselves would be an undertaking that would surely span longer than the time span of the whole project, therefore it was decided early to use an existing parser.
     The most capable and mature C++ parsers will naturally be the parsers used by the most popular compilers.
@@
     The literature~\cite{Duffy2014} is fairly clear on the fact that the design of GCC makes it unsuitable for integration with other projects.
-    Clang on the other hand allows us access to the Clang abstract syntax tree (AST) and parser internals through standard design patterns such as consumers and visitors.
+    Clang on the other hand allows us access to the Clang abstract syntax tree (AST) and parser internals via a standard API such as consumers and visitors.
 
@@
 
-    The main capability of the tool is to be able to identify what the maximum number of owners a pointer has at all points of the program execution for all pointers in any given program.
+    The main capability of the tool is to be able to identify the maximum number of owners a pointer has at all points of the program execution for all pointers in any given program.
     If this number for a given raw pointer is one, then we can go about converting all usages of that raw pointer into a unique pointer.
@@
     Escape analysis is used to determine whether an object \emph{escapes} or is accessible from outside the method or thread that created the object~\cite{Choi1999}.
-    It has been used in languages such as Java~\cite{Choi1999} to determine whether an object that is created in a method escapes, and if it does not, meaning it is only used locally in the method, then it can be allocated on the stack rather than the heap as a performance optimisation.
-    % TODO pin footnote to specific commit possibly
+    It has been used in languages such as Java~\cite{Choi1999} to determine whether an object which is created within a method escapes.
+    If it does not, meaning it is only used locally in the method, then it can be allocated on the stack rather than the heap as a performance optimisation.
     Clang also uses escape analysis for a similar kind of optimisation~\footnote{\url{https://github.com/llvm/llvm-project/blob/master/llvm/lib/Analysis/CaptureTracking.cpp}} and also defines the term \emph{pointer capture}.
     A pointer value is captured if the function makes a copy of any part of the pointer that outlives the call.
-    We can then see that any pointer that escapes must also be captured since there is no way to refer to an object without being able to determine its address, but not visa versa.
+    We can then see that any pointer that escapes must also be captured since there is no way to refer to an object without being able to determine its address, but not every captured pointer also escapes.
     In \autoref{lst:capture-example} we can see an example where a pointer may be captured but does not escape and in \autoref{lst:escape-example} we see an example of an escaped pointer.
@@
         \inputminted{c++}{code/escape-example.cpp}
-        \caption{The pointer is both captured and escapes since the pointer and its value outlive the call of the function and it is accessible by any other method or thread since it is copied to a global variable}
+        \caption{The pointer is both captured and escapes since the pointer and the value it points to outlive the call of the function and become accessible by any other method or thread as it is copied to a global variable}
         \label{lst:escape-example}
@@
 
-    Now that we have defined how escape analysis can be used to identify areas of refactoring, we now need to identify a suitable algorithm.
-
-    There has been extensive research on escape analysis for Java, where escape analysis is used to determine whether a object escapes the current method.
-    If it can be proven that the object does not escape the method, then it is local to the method invocation and it can be allocated on the stack rather than as a heap variable (the default for all Java objects).
+    Now that we have defined how escape analysis can be used to identify areas of refactoring, we need to identify a suitable algorithm.
 
@@
     The connection graph captures relationships between heap-allocated objects and object references.
-    The graph is used to perform reachability analysis to determine if an is local to a thread or method.
+    The graph is used to perform reachability analysis to determine if an object is local to a thread or method.
     Most importantly, this abstraction allows for a powerful interprocedural analysis to be performed.
@@
     $N_{r}$ is the set of reference nodes which include local variables, parameters, fields and global reference variables.
-    $E_{p}$ is the set of points-to edges which exists between a reference node which points to object node.
+    $E_{p}$ is the set of points-to edges which exist between a reference node that points to object node.
     $E_{d}$ is the set of deferred edges, where a deferred edge from one node $p$ to node $q$ signifies it points to what $q$ points to.
@@
 
-    Rust has been the most popular language in attractive widespread adoption as a systems language.
-    Its popularity could be quantified by considering that it is being considered for usage in writing Linux kernel modules~\footnote{\url{https://lore.kernel.org/lkml/CAKwvOdmuYc8rW_H4aQG4DsJzho=F+djd68fp7mzmBp3-wY--Uw@mail.gmail.com/T}}, which cannot be said for C++.
+    Rust has been the most popular language in attracting widespread adoption as a systems language.
+    Its popularity could be quantified by considering that it is being considered for use in writing Linux kernel modules~\footnote{\url{https://lore.kernel.org/lkml/CAKwvOdmuYc8rW_H4aQG4DsJzho=F+djd68fp7mzmBp3-wY--Uw@mail.gmail.com/T}}, which cannot be said for C++.
     Following the advice given by many that most systems programs should be written in Rust from now on, we see a potential future usage of our work in the creation of a C++ to Rust transpiler.
 
-    Whilst the automatic refactoring of raw pointers into smart pointers solves some memory issues, the language still offers other ways to invalidate memory safety.
+    Whilst the automatic refactoring of raw pointers into smart pointers solves some memory issues, the language still offers other ways to break memory safety.
     Equivalent idiomatic Rust would be able to find some of these other errors at compile time.
@@
 
-    Garbage collected languages encompass some of the most popular programming languages such as Java and Python.
+    Garbage collected languages encompass some of the most popular programming languages such as Java or Go.
     They are considered to be simpler languages than C++ since they remove the need for manual memory management by the programmer.
     In C++, if memory is improperly freed, it gets leaked and can eventually cause out of memory errors.
-    Garbage collectors instead run periodically and scan the memory of the program to find objects in memory which were allocated but are no longer used anymore, i.e leaked and then free the leaked objects.
+    Garbage collectors instead run periodically and scan the memory of the program to find objects in memory which were allocated but are no longer used anymore and then free the leaked objects.
     Languages such as Java~\cite{Tauro2012} use this fully by allocating almost every object on the heap and not freeing any object, instead relying on the garbage collector to clean up.
@@
     First and foremost, we have evidence that this is a problem that actually exists in industry through a paper by \citet{Wright2013} which describes a tool used in Google for running such rejuvenations across a large codebase and shows an example of running a simple rejuvenation of upgrading an API call.
-    They only uses Clang for the parsing of code into an AST and traversing the tree to identify and perform simple refactorings.
+    They only use Clang for the parsing of code into an AST and traversing the tree to identify and perform simple refactorings.
     This allows for less complex refactorings, however it also allows for a flexible design to implement different refactorings using the same tool, rather than the focus we have for a single specific refactoring goal.
@@
     Secondly, there has been separate academic research into rejuvenation tools such as the work by \citet{Huck2016} and \citet{Kumar2012}.
-    The former also uses the Clang and AST driven approach that we have described in \autoref{subsec:clang-and-llvm} and by \citet{Wright2013}, whilst the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
+    The former also uses the Clang and AST driven approach that we have described in \autoref{subsec:clang-and-llvm} and by \citet{Wright2013}, while the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
     We have not evaluated IPR as a representation to use instead of the Clang AST, but we can however comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation that Clang has, as well as it being the choice of several other successful projects leads us to believe that we will not lose any effectiveness in our use of it.
@@
 
-    % TODO rename section to something better
-
     \subsection{Current Progress}\label{subsec:current-progress}
@@
     As mentioned in \autoref{subsec:clang-and-llvm}, we plan on using Clang for parsing C++ code into an AST and LLVM IR for the analysis.
-    We can employ the visitor pattern to traverse the AST and look for nodes of interest, such as variable declaration nodes.
+    We can traverse the AST and look for nodes of interest, such as variable declaration nodes.
     Given a node in the AST, we can find its equivalent declaration in the IR\@.
-    This is mainly possible since we can see from comparing \autoref{lst:code-to-ast-example} and \autoref{lst:code-to-ast-ir} that the variable names can matched from the AST to the IR\@.
+    This is mainly possible since we can see in \autoref{lst:code-to-ast-ir} that the variable names can be matched from the code and therefore AST to the IR\@.
 
     \begin{listing}
-        \inputminted[fontsize=\footnotesize, breaklines]{llvm}{code/code-to-ast-summary.ll}
-        \caption{The LLVM IR representation of the program in \autoref{lst:code-to-ast-example}. Listing has been summarized for clarity.}
+        \inputminted{c++}{code/code-to-ast.cpp}
+        \inputminted[fontsize=\footnotesize]{llvm}{code/code-to-ast-summary.ll}
+        \caption{The LLVM IR representation of a small C++ code snippet. Listing has been summarized for clarity.}
         \label{lst:code-to-ast-ir}
@@
     It also shows how Clang library allows for easy refactoring since the AST stores all source text and line numbers, making text substitution trivial.
-    Substitution of AST nodes rather than source text substitution has not been explored, however this is as documentation states that generally the AST should remain immutable.
+    Substitution of AST nodes rather than source text substitution has not been explored, however this because Clang documentation states that generally the AST should remain immutable.
 
@@
     Therefore if each reference node has at most one incoming edge, then we know that each pointer is copied at most once, and therefore only transfers ownership once.
-    This therefore means that the value will have at most one unique throughout program execution.
+    This therefore means that the value will have at most one unique owner throughout program execution.
     This relies on careful modeling where using pointers as function arguments is modelled as the creation of new reference nodes, which can be correctly modelled in C++ as it has pass-by-value semantics.

### 2020-12-18 proposal.tex
final few fixes

@@
     Such memory safety vulnerabilities can be exploited, and due to new regulations such as the GDPR, these issues are more commonly exposed to the general public and punished by regulators.
-    An example is a fine of £20m fine for a British Airways data breach by the British Information Commissioner's Office~\cite{ICO2020}.
+    An example is a fine of £20m for a British Airways data breach by the British Information Commissioner's Office~\cite{ICO2020}.
 
@@
     The first type of smart pointer are \emph{unique pointers} which have only a single unique owner at any time.
-    We can therefore fix the previous example in which we leak memory using a unique pointer, as can be seen in \autoref{lst:first-example-unique}.
-    In the comments of \autoref{lst:first-example-unique}, we can see all the locations where \emph{p} exits the scope of the function and where the compiler will automatically insert delete statements, now ensuring that the resource will never get leaked.
+    We can therefore fix the previous example in which we leak memory by using a unique pointer, as can be seen in \autoref{lst:first-example-unique}.
+    In the comments of \autoref{lst:first-example-unique}, we can see all the locations where the control flow exits the scope where \emph{p} is defined and where the compiler will automatically insert delete statements, now ensuring that the resource will never get leaked.
     The ownership of a unique pointer can be transferred between variables using \mintinline{c++}{std::move} which moves the ownership from one variable to another, making the original reference invalid.
@@
     It would appear then that smart pointers could solve most of our memory management issues, but they were only introduced in the C++11 standard.
-    The C++ standards (\textit{versions}) before then did not have these language features, with older programs being written predominantly in the C++03 standard.
+    The C++ standards (\textit{versions}) before then did not have these language features, with older programs being written predominantly in the C++98 or C++03 standard.
     There remain many such called \emph{legacy} programs which have not upgraded to newer standards, and even then the programmer can choose not to utilise smart pointers in their code.
@@
     The AST that it produces is also rich in information and closely resembles the original code, as can be seen in \autoref{lst:code-to-ast-example}.
-    Information such as the line and column numbers are included, which makes programmatically rewriting source code extremely easy.
+    Information such as the line and column numbers are included, which makes programmatically rewriting source code easy.
 
@@
     For each node, we associate a state of either \emph{NoEscape}, \emph{ArgEscape} or \emph{GlobalEscape}.
-    \emph{NoEscape} means that the object does not escape the method in which it was created
+    \emph{NoEscape} means that the object does not escape the method in which it was created.
     \emph{ArgEscape}, with respect to a method, means that the object escapes that method via the method arguments or return value,

### 2021-03-05 paper.tex
Add Introduction

@@
-\documentclass{paper}
+\documentclass{mpaper}
+
+\usepackage[numbers]{natbib}
+\usepackage{xurl}
 
@@
 
-    \title{ptr-tidy: Automatic refactoring of raw pointers in C++}
+    \title{ptr-tidy: Automatic Rejuvenation of Raw Pointers in C++}
     \author{Artem Usov}
@@
 
+    Most ordinary computer users demand that the machine they are using provides them with a responsive, secure and productive environment to complete their tasks.
+    The programs that are most responsible for this are complex systems programs such as the underlying operating system, device drivers and web browsers.
+    Systems programming often involves memory management, that is requesting memory from the operating system to be used and managed by the program.
+    This is done when the amount of memory that we need cannot be know at the time of compilation of the program.
+    However, doing so creates the opportunity for memory safety errors~\cite{Dhurjati2003} which are notoriously challenging to avoid.
+    If there are memory safety errors in a program, there is the possibility that it could be exploited by a malicious party to crash the program, expose sensitive data or gain remote code execution capabilities~\cite{Afek2007}.
+
+    Historically systems programs are mostly written in C, C++ or a combination of the two and remain the dominant languages.
+    They are lower level languages that involve manual memory management by the programmer.
+    It is even more challenging to ensure that C and C++ programs contain no memory safety errors as evidenced by empirical evidence in industry, where in a presentation at the Linux Security Summit it is shown that in several popular projects such as Firefox, macOS, Ubuntu and Android all had over half of their CVEs attributed to issues with memory safety~\cite{Gaynor2020}.
+    Similarly, in a presentation by Matt Miller, a security engineer at Microsoft, it is shown that around 70\% of their vulnerabilities that are addressed through a security updates are due to memory safety issues~\cite{Miller2019}.
+    This is at industry-leading companies who are renowned for hiring top talent and top developer education programs.
+
+    However the continued use of C and C++, alongside industry mottos such as \emph{``Move fast and break things''}~\cite{Vardi2018} do not help by possibly diminishing the danger of memory safety error exploitation.
+    However due to new regulations such as the GDPR and increasing oversight by regulators, any such exploitations which do expose sensitive data are more commonly exposed to the general public and punished by regulators.
+    An example is a fine of £20m for a British Airways data breach by the British Information Commissioner's Office~\cite{ICO2020}.
+
+    However simply not using these languages is not an option since their unmanaged nature makes them highly performant, and therefore the best option for systems such as a web browser.
+    Engineers at both Microsoft~\cite{Thomas2019} and Mozilla~\cite{Hostfelt2019} converge on Rust~\cite{Balasubramanian2017} as a possible solution.
+    Rust is a systems language that offers similar performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model also guarantee memory safety.
+    The rewriting of a program in a new language, namely from C++ to Rust, is a colossal undertaking, especially given that Rust has a reputation for being difficult to learn.
+    Instead, modern C++ guidelines encourage the use of \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create memory or resource handles which automatically eliminate entire classes of memory safety errors.
+    Yet, we encounter further problems as C++ programs either do not actively follow this advice, or are \emph{legacy} programs that were written before these features were introduced to the language.
+
+    We therefore propose a solution in the form of a tool, \emph{ptr-tidy}, to analyse and automatically refactor programs to follow the recommended use of smart pointers where the analysis can determine that such a change is syntactically correct and does not change the runtime behaviour of the program.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+    \section{Related Work}\label{sec:related-work}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \section{Background}\label{sec:background}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+    \section{Implementation}\label{sec:implementation}
 
@@
 
-    \bibliographystyle{abbrv}
+    \bibliographystyle{abbrvnat}
     \bibliography{paper}

### 2021-03-07 src/PtrRewriter.cpp
First implementation of adding gets to smart pointer usages

@@ -37,6 +37,7 @@ template <typename Tool> void PtrRewriter<Tool>::rewrite(const VarDecl *var) {
     rewriteInit(var);
     removeDelete(var);
     rewriteFunctionReturn(var);
+    addSmartGet(var);
 
     if (!m_headerAdded) {
       auto top = var->getASTContext().getTranslationUnitDecl();
@@ -148,6 +149,24 @@ template <typename Tool> void PtrRewriter<Tool>::rewriteFunctionReturn(const Var
     finder.addMatcher(returnStmt().bind(""), &callback);
     m_tool.run(clang::tooling::newFrontendActionFactory(&finder).get());
   }
+}
+
+template <typename Tool> void PtrRewriter<Tool>::addSmartGet(const clang::VarDecl *var) {
+  auto f = [&](const MatchFinder::MatchResult &result) {
+    auto call = result.Nodes.getNodeAs<CallExpr>("");
+    for (auto i = 0; i < call->getNumArgs(); ++i) {
+      auto declRef = dyn_cast<DeclRefExpr>(call->getArg(i)->IgnoreImpCasts());
+      if (declRef && declRef->getDecl()->getID() == var->getID()) {
+        m_rewriter.ReplaceText(SourceRange(declRef->getBeginLoc(), declRef->getEndLoc()),
+                               declRef->getNameInfo().getName().getAsString() + ".get()");
+      }
+    }
+  };
+
+  LambdaCallback callback(f);
+  MatchFinder finder;
+  finder.addMatcher(callExpr(hasAnyArgument(ignoringParenImpCasts(declRefExpr()))).bind(""), &callback);
+  m_tool.run(clang::tooling::newFrontendActionFactory(&finder).get());
 }
 
 template class PtrRewriter<clang::tooling::ClangTool>;

### 2021-03-07 src/PtrRewriter.h

@@ -25,6 +25,7 @@ template <typename Tool> class PtrRewriter {
   void rewriteDeferredInit(const clang::VarDecl *var);
 
   void removeDelete(const clang::VarDecl *var);
+  void addSmartGet(const clang::VarDecl *var);
 
 public:
   explicit PtrRewriter(Tool &tool) : m_tool(tool) {}

### 2021-03-07 test/testPtrRewriter.t.cpp

@@ -208,3 +208,12 @@ TEST_F(PtrRewriterTest, HandleStructCallInitValue) {
             "struct A {int m_a; A(int a) : m_a(a) {}}; std::shared_ptr<A> test(){std::shared_ptr<A> a = "
             "std::make_shared<A>(1); return a;}");
 }
+
+TEST_F(PtrRewriterTest, HandleAddingGetToDeclRef) {
+  tool.code = "void foo(int *a){} int *test(){int *a = new int(2); foo(a); return a;}";
+
+  runToolOnCode(factory->create(), tool.code);
+  ASSERT_EQ(compareRewriterOutput(rewriter.getRewriter()),
+            "void foo(std::shared_ptr<int> a){} std::shared_ptr<int> test(){std::shared_ptr<int> a = "
+            "std::make_shared<int>(2); foo(a.get()); return a;}");
+}

### 2021-04-09 src/PtrRewriter.cpp
Fix clang-tidy error in LLVM 11.1.0, Update Readme

@@ -154,7 +154,7 @@ template <typename Tool> void PtrRewriter<Tool>::rewriteFunctionReturn(const Var
 template <typename Tool> void PtrRewriter<Tool>::addSmartGet(const clang::VarDecl *var) {
   auto f = [&](const MatchFinder::MatchResult &result) {
     auto call = result.Nodes.getNodeAs<CallExpr>("");
-    for (auto i = 0; i < call->getNumArgs(); ++i) {
+    for (unsigned int i = 0; i < call->getNumArgs(); ++i) {
       auto declRef = dyn_cast<DeclRefExpr>(call->getArg(i)->IgnoreImpCasts());
       if (declRef && declRef->getDecl()->getID() == var->getID()) {
         m_rewriter.ReplaceText(SourceRange(declRef->getBeginLoc(), declRef->getEndLoc()),

### 2021-04-11 paper.tex
Add rest of paper from overleaf

@@
 
+\usepackage{minted}
 \usepackage[numbers]{natbib}
-\usepackage{xurl}
+\usepackage[font=bf]{caption}
+
+\providecommand*{\listingautorefname}{Listing}
+
+\setminted{frame=lines, autogobble, breaklines}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
     \begin{abstract}
+    Systems programming involves dynamically requesting memory that in the C++ language has to be manually managed by the programmer. 
+    However, this leaves the opportunity for exploitable memory safety errors.
+    Modern C++ recommends that smart pointer are used to automatically manage resources such as dynamic memory.
+    However, many legacy programs may have been written before these features were introduces to the language.
+    We developed the ptr-tidy tool in order to automatically rejuvenate such programs to use smart pointers.
+    The tool uses the Clang and LLVM libraries in order to create a componentised and powerful design which could be reused by other refactoring or rejuvenation tools. 
+    We present a case study and and apply the tool on a number of open-source C++ libraries to show that the tool can successfully rejuvenate non-trivial libraries and that the tool can be used on libraries from many areas of computer science.
+    We discuss in which circumstances the tool performs well and the current limitations.
     \end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
     The programs that are most responsible for this are complex systems programs such as the underlying operating system, device drivers and web browsers.
-    Systems programming often involves memory management, that is requesting memory from the operating system to be used and managed by the program.
-    This is done when the amount of memory that we need cannot be know at the time of compilation of the program.
+    Systems programming involves memory management, that is dynamically requesting memory from the operating system to be used and managed by the program.
+    This is done when the amount of memory that we need at runtime cannot be know at the time of compilation of the program.
     However, doing so creates the opportunity for memory safety errors~\cite{Dhurjati2003} which are notoriously challenging to avoid.
+    A memory safety error can be one of:
+    \begin{itemize}
+        \item \textbf{Dangling pointers}, prematurely freeing an object that is still in use.
+        \item \textbf{Double free}, deleting an object multiple times.
+        \item \textbf{Never free}, never deleting an object.
+    \end{itemize}
+    There are other categories of memory safety errors such as buffer overflows which we do not consider in this paper.
     If there are memory safety errors in a program, there is the possibility that it could be exploited by a malicious party to crash the program, expose sensitive data or gain remote code execution capabilities~\cite{Afek2007}.
 
-    Historically systems programs are mostly written in C, C++ or a combination of the two and remain the dominant languages.
+    Historically systems programs are mostly written in C, C++ or a combination of the two and these remain the dominant languages.
     They are lower level languages that involve manual memory management by the programmer.
@@
 
-    However the continued use of C and C++, alongside industry mottos such as \emph{``Move fast and break things''}~\cite{Vardi2018} do not help by possibly diminishing the danger of memory safety error exploitation.
-    However due to new regulations such as the GDPR and increasing oversight by regulators, any such exploitations which do expose sensitive data are more commonly exposed to the general public and punished by regulators.
-    An example is a fine of £20m for a British Airways data breach by the British Information Commissioner's Office~\cite{ICO2020}.
+    % However the continued use of C and C++, alongside industry mottos such as \emph{``Move fast and break things''}~\cite{Vardi2018} do not help by possibly diminishing the danger of memory safety error exploitation.
+    % However due to new regulations such as the GDPR and increasing oversight by regulators, any such exploitations which do expose sensitive data are more commonly exposed to the general public and punished by regulators.
+    % An example is a fine of \pounds20m for a British Airways data breach by the British Information Commissioner's Office~\cite{ICO2020}.
 
-    However simply not using these languages is not an option since their unmanaged nature makes them highly performant, and therefore the best option for systems such as a web browser.
+    Unfortunately, simply not using these languages is not an option since their unmanaged nature makes them highly performant, and therefore the best option for systems such as a web browser.
     Engineers at both Microsoft~\cite{Thomas2019} and Mozilla~\cite{Hostfelt2019} converge on Rust~\cite{Balasubramanian2017} as a possible solution.
     Rust is a systems language that offers similar performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model also guarantee memory safety.
-    The rewriting of a program in a new language, namely from C++ to Rust, is a colossal undertaking, especially given that Rust has a reputation for being difficult to learn.
+    The rewriting of a program in a new language, namely from C++ to Rust, is a colossal undertaking, especially given that Rust has a reputation for being difficult to learn~\cite{Astrauskas2019}.
     Instead, modern C++ guidelines encourage the use of \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create memory or resource handles which automatically eliminate entire classes of memory safety errors.
@@
 
-    We therefore propose a solution in the form of a tool, \emph{ptr-tidy}, to analyse and automatically refactor programs to follow the recommended use of smart pointers where the analysis can determine that such a change is syntactically correct and does not change the runtime behaviour of the program.
+    We therefore propose a solution in the form of a tool, \emph{ptr-tidy}~\cite{Usov2021}, to analyse and automatically refactor programs to follow the recommended use of smart pointers where the analysis can determine that such a change is syntactically correct and does not change the runtime behaviour of the program.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{listing}
+        \inputminted{c++}{code/leak.cpp}
+        \inputminted{c++}{code/fix-leak.cpp}
+        \caption{Example of a memory leak using manual management and how smart pointers correct this.
+Lines 4 and 5 show how interruptions in the program control flow can mistakenly cause memory leaks and how smart pointer correct this by automatically inserting all needed delete statements at compile time.
+All necessary headers are omitted for brevity}
+        \label{lst:motivating-example}
+    \end{listing}
+
+    We provide a motivating example in \autoref{lst:motivating-example} where the first version of the function could have several control flow executions where the \texttt{Circle} object will leak.
+    The example seems quite trivial, but can suddenly be quite non-trivial if the conditions in the \texttt{if} statement become calls into external libraries, which may or may not throw an exception.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    Our work closely relates to work done around the field of \emph{source code rejuvenation}, as defined by \citet{Pirkelbauer2010}. 
+    We exactly aim replace outdated coding patterns with newer, higher-level abstractions, as well as fitting the notion that this tool would be applied once to a codebase, rather than be a reoccurring task.
+    We will however use the terms rejuvenation and refactoring interchangeably in this paper as it is a term that most developers are familiar with.
+
+    We believe research in this area is going to increase, as languages move and evolve much quicker than can be supported in enterprise environments, resulting in accumulating technical debt. 
+    This is supported by the fact that other relevant work has been completed on this subject.
+
+    First and foremost, we have evidence that this is a problem that exists in industry through a paper by \citet{Wright2013} which describes a tool used at Google for running such rejuvenations across a large codebase.
+    \citet{Wright2013} show an example of running a simple rejuvenation of upgrading an API call.
+    They use Clang for the parsing of code into an AST and traverse the tree to identify and perform simple refactorings.
+    Their flexible design allows them to implement many different simple refactorings using the same tool, rather than the focus we have for a single specific refactoring goal.
+
+    Secondly, there has been separate academic research into rejuvenation tools such as the work by \citet{Huck2016} and \citet{Kumar2012}.
+    The former also uses the Clang and AST driven approach like \citet{Wright2013}, while the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
+    We have not evaluated IPR as a representation compared to the Clang AST, but we can however comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation that Clang has, as well as it being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
+
+    There has also been industry interest in automatic rejuvenation tools between languages rather than within the same language. 
+    This is done either to increase the use of a new, improved language by creating creating tools to aid transition, or in order to reuse existing compilers.
+    For example, the Kotlin language team created the J2K~\footnote{\url{https://github.com/JetBrains/kotlin/tree/master/j2k}} transpiler in order to entice Java developers to switch to a language with improved language features such as null safety~\cite{Mateus2019}.
+    Similarly, several new languages such as Typescript and Dart have been recently developed which were designed not to be used by themselves, but to be transpiled to the popular Javascript language.
+    These languages aim to solve shortcomings of the output language, whilst reusing the existing mature and efficient compilers or interpreters.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \subsection{Using Smart Pointers for Memory Management}\label{subsec:using-smart-pointers-for-memory-management}
 
-    \section{Implementation}\label{sec:implementation}
+    Generally pointers are used to give the program access to a resource that cannot be directly included in the program itself, such as allocated memory or a file.
+    Pointers can also be used to pass large resources to functions without needing to create a copy, which would have a performance impact.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    Pointers can be created for objects both on \emph{the stack and the heap}~\cite{Stroustrup2000}. 
+    The stack is an small memory store automatically managed for us by the compiler and is suitable for placing small, fixed sized parameters and arguments, such as perhaps loop counter variables.
+    However, when objects must live beyond the lifetime of the current function call or if we do not know the size of an object until runtime, then we must storage these objects on the heap, occasionally called the \emph{free store}.
+    Unlike the stack, the memory on the heap must be explicitly managed by the programmer.
+
+    However Stroustrup claims that pointers to objects allocated on the free store are dangerous and a \emph{plain old pointer}, or raw pointer as we will refer to them, should not be used to represent ownership~\cite{Stroustrup2018}.
+
+    Instead we can use \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create resource handles which automatically eliminate resource leaks with no, or very little added overhead.
+    RAII means that the management of a resource, namely its acquisition and release, is bound to its lifetime. 
+    In the simplest case, this means that when a resource exists the scope of a function or method, it is released automatically.
+
+    The first type of smart pointer are \emph{unique pointers} which have only a single unique owner at any time and therefore are ideal for resources which may not be copied, such as perhaps a global session dictionary.
+    We can see how unique pointers and RAII work from out motivating example in  \autoref{lst:motivating-example}.
+    Even though the \texttt{Circle} object is placed on the heap, the compiler can manage the smart pointer for us so that in all cases where the lifetime of the object ends, the object will be freed and ensures the resource will not leak.
+    The ownership of a unique pointer can be transferred between variables using \texttt{std::move} which moves the ownership from one variable to another, making the original reference invalid.
+    % TODO insert example diagram possibly
+
+    The second type of smart pointer are \emph{shared pointers} which are used for resources which may not necessarily have a single unique owner and so several owners \emph{share} ownership to the resource.
+    Reference counting~\cite{Collins1960} is used to ensure that after the last owner loses access to the resource, it will finally be freed.
+    Shared pointers are suitable to be used in most situations and such reference counted pointers are used by default for all values in the Python and Swift languages~\footnote{\url{https://docs.swift.org/swift-book/LanguageGuide/AutomaticReferenceCounting.html}}
+
+    \subsection{Clang and LLVM}\label{subsec:clang-and-llvm}
+
+    \begin{figure*}
+        \centering
+        \includegraphics[scale=0.75]{images/llvm.pdf}
+        \caption{LLVM design diagram showing how language front-ends create LLVM IR that can be optimised and then used by an architecture back-end, such as x86 to create a machine code executable.}
+        \label{fig:llvm-design}
+    \end{figure*}
+
+    As seen in \autoref{sec:background}, much similar research uses Clang and LLVM.
+    LLVM~\cite{Lattner2004} is a compiler infrastructure project consisting of several compiler and toolchain technologies. 
+    LLVM is centrally designed around a language independent intermediate representation (IR), which can be created by a number of language front-ends and is used by a number of architecture back-ends to generate machine code.
+    Program optimisations can then be applied by portable and reusable transformations on the IR.
+    Any improvement in back-end code generation, as well as any new or improved IR optimisers can then benefit any of the supported language front-ends.
+    It is therefore visible that this allows for a very modular and extensible design which can be seen in \autoref{fig:llvm-design}.
+
+    Clang~\cite{Lattner2008} is such an LLVM front-end for C, C++ and a variety of other similar languages such as OpenGL.
+    By using the LLVM infrastructure, it allows for faster and more efficient compilation of C++ compared to existing, older compilers such as the GNU Compiler Collection~\footnote{\url{https://gcc.gnu.org/}} (GCC)~\cite{Lattner2008}.
+    However, most importantly for our research, the design of GCC makes it unsuitable for integration with other projects, such as analysis tools, whereas Clang allows access to several analysis tools as well as the Clang abstract syntax tree (AST) and parser internals via a standard API such as consumers and visitors~\cite{Duffy2014}.
+
+    \begin{figure}
+        \centering
+        \inputminted{c++}{code/code-to-ast.cpp}
+        \includegraphics{images/code-to-ast.pdf}
+        \caption{Example of converting C++ code to Clang AST. All diagrams show equivalent representation of a small C++ code snippet.}
+        \label{fig:code-to-ast-example}
+    \end{figure}
+
+    Further more, Clang also explicitly provides the \emph{LibTooling} C++ library for the creation of tools that leverage the Clang parsing front-end and AST.
+    The Clang AST is different from ASTs produced by some other compilers in that it closely resembles both the written C++ code and the C++ standard~\footnote{\url{https://clang.llvm.org/docs/IntroductionToTheClangAST.html}}.
+    For example, we can see in \autoref{fig:code-to-ast-example} that the AST stores much syntactic as well as semantic information in the form of source code locations and variable initialisation styles.
+    The LLVM authors remark that this makes it very suitable for refactoring tools.
+
+    The LLVM IR represents a language-independent, static single assignment (SSA)~\cite{Rosen1988} representation of the program.
+    Variables in in SSA form can only be assigned once, which allows us to efficiently generate compact definition-usage graphs, which enable developers to simplify many dataflow optimizations and enables fast, flow-insensitive algorithms~\cite{Lattner2004}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+    \section{Methodology}\label{sec:methodology}
+
+    \subsection{Parsing}
+
+    The first capability our tool would need is to be able to parse C++ code.
+    Doing this ourselves would be an undertaking that would take a significant amount of time in order to create a complete parser due to the size and complexity of the C++ language, therefore it was decided early to use an existing parser.
+    The information in \autoref{subsec:clang-and-llvm} outlines why Clang is the most suitable parser to use, with the other benefit that it brings such as \emph{LibTooling}.
+
+    \subsection{Analysis}\label{subsec:methodology-analysis}
+
+    Next, our tool would need to be able to identify specifically which pointers are suitable for refactoring.
+    An \emph{unique pointer} is a smart pointer object that represent a single unique owner of a region of memory at any time. 
+    We will therefore need to develop a static analysis algorithm to identify the maximum number of owners a region of memory has throughout all points of a program. 
+    If a memory region has at most a single unique owner at a time, we can refactor the initialising raw pointer to a unique pointer, otherwise the pointer can be refactored to a shared pointer.    
+    This analysis will need to ensure that its result is correct, and will inevitably have to make some conservative assumptions about the program.
+
+    \begin{figure}
+        \centering
+        \includegraphics[scale=1.0]{images/memory.pdf}
+        \caption{Both the variables \texttt{a} and \texttt{c} end up pointing to memory address \texttt{0001}.
+Similarly variables \texttt{b} and \texttt{d} point to address \texttt{0003}.
+The memory at \texttt{0001} is used through two different pointers on line 5, meaning that it is not uniquely owned by a single owner.
+However, address \texttt{0003} does have a unique single owner at any time in the program, shown by the single incoming arrow to the memory region.
+Therefore the initialisation of variable \texttt{d} can be modelled as an ownership transfer from \texttt{b}}
+        \label{fig:memory-pointers}
+    \end{figure}
+
+    A key intuition is that a region of memory, subsequent to its initialisation, is only able to gain additional owners if access to the region of memory is \emph{shared} to another variable, that is if a pointer to this memory is copied to a another pointer, at which point the original owner is longer the unique owner.
+    If we copy the pointer to a global variable, the variable could suddenly be read by any other running threads in the program, so we have to take a conservative decision that the memory no longer has a single unique owner.
+    If we copy the pointer to a local variable, then if the original variable remains in use, the memory will have two unique owners. 
+    However if the original pointer is not used beyond being copied, this could actually be modelled as a \emph{transfer} of ownership, and the memory will still have a single unique owner at all points of the program.
+    We show this intuition in \autoref{fig:memory-pointers}, where the memory region \texttt{0001} is not suitable for being managed by a unique pointer, whilst region \texttt{0003} is.
+
+    We can formalise our key intuition by introducing the concept of \emph{escape analysis}.
+    Escape analysis is used to determine whether an object \emph{escapes} or is accessible from outside the method or thread that created the object~\cite{Choi1999}.
+    It has been used in languages such as Java~\cite{Choi1999} to determine whether an object which is created within a method escapes.
+    If it does not escape, meaning it is only used locally in the method, then it can be allocated on the stack rather than the heap as a performance optimisation.
+    LLVM also uses escape analysis for a similar kind of optimisation~\footnote{\url{https://github.com/llvm/llvm-project/blob/main/llvm/lib/Analysis/CaptureTracking.cpp}}.
+
+    It is now visible that the invariant of the unique pointer are only violated when:
+    \begin{itemize}
+        \item a pointer escapes, as then we do not know where or how the pointer may be used, so we must take a conservative approach and assume the memory region no longer has at most a single unique owner.
+        \item a pointer is does not escape but access to the memory is shared rather than transferred to a local variable. 
+    \end{itemize}
+
+    \subsection{Analysis Target}\label{subsec:methodology-analysistarget}
+
+    The existing escape analysis in LLVM is implemented to operate on LLVM IR code, as the IR representation, as mentioned in \autoref{subsec:clang-and-llvm}, is well suited for efficient analysis algorithms.
+    However, much of the existing work from \autoref{sec:related-work} which use Clang perform both their analyses and their source code refactoring on the Clang AST.
+    We however propose an improved novel design which uses the Clang AST for parsing and refactoring and the LLVM IR for simple and efficient analysis.
+
+    For the same input file, we can use the \emph{LibTooling} library to generate an AST and IR in-memory representation.
+    Then, for any variable declaration (\emph{VarDecl}) in the Clang AST which allocates new memory on the heap, we locate this declaration in the IR.
+    This is possible as we can instruct Clang to not discard variable names using a \texttt{fno-discard-value-names} flags, and as the combination of function signature and variable name in the translation unit (\emph{cpp file}) must be unique according to the One-definition rule~\cite{C++17}.
+
+    \begin{listing}
+        \inputminted{llvm}{code/code-to-ir.ll}
+        \caption{The LLVM IR module representation of the C++ code in \autoref{fig:code-to-ast-example}.
+Listing has been summarised for clarity.
+The variable on line 5 can be matched to the variable \texttt{a} in \autoref{fig:code-to-ast-example} by matching the function signature and variable name.}
+        \label{lst:code-to-ir}
+    \end{listing}
+
+    In \autoref{fig:code-to-ast-example}, given the variable declaration for the variable \texttt{a}, then as it is not a global variable, we can walk its parent nodes until we reach a \emph{FunctionDecl}.
+    We are now able to use the function name to find the corresponding function declaration in the IR module, which can be seen in  \autoref{lst:code-to-ir}.
+    Note that although the \texttt{main} function name corresponds exactly, other function names in the IR module will have mangled names~\footnote{\url{https://www.ibm.com/support/knowledgecenter/en/ssw_ibm_i_72/rzarg/name_mangling.html}}.
+    We again use Clang library functions to mangle function names if necessary so they can be found in the IR. 
+    Given a function in the IR, we can use its symbol table to use the declaration name to find the IR value on which we will perform our analysis.
+
+    \subsection{Choice of Escape Analysis Algorithm}
+
+    The initial choice of escape analysis algorithm is the existing LLVM algorithm mentioned in \autoref{subsec:methodology-analysis}.
+    As it is already developed for use on LLVM IR, it is therefore trivial to integrate into the tool which allows us to quickly create a minimal viable product. 
+    This allowed for rapid successful prototyping of the tool and therefore reduces risk of the chosen design being unsuitable by following rapid development processes~\cite{Martin1991}.
+    We show in \autoref{subsec:methodology-systemdesign} that the design of our tool means that the initial choice of algorithm will not place a limit on the final accuracy or power of the tool.
+
+    Investigation of the algorithm shows that it is a conservative analysis with some limits on the analysis complexity so that it does not greatly increase Clang compilation time.
+    The analysis uses the definition-usage graph of the value being analysed to add all the instructions that use the value to a work queue.
+    Each item in the work queue is then processed and the algorithm dispatches on the type (\emph{opcode}) of the IR instruction.
+    It is then checked whether each instruction within the context that it is being used in could cause the value to escape.
+    The key limitation that we see of this algorithm is that it is a intraprocedural analysis.
+    Pointers are very often used across function boundaries and the conservative decisions the algorithm will take in these situations might limit the effectiveness of the tool in identifying all possible refactoring opportunities.
+
+    \subsection{Source code rewriting}
+
+    The analysis gives a result of whether the IR value escapes, a boolean \texttt{true} or \texttt{false}.
+    For a \texttt{false} result, the value declaration is refactored into a unique pointer, otherwise a shared pointer.
+    However, the rewriting is identical for both cases with just the string tokens that we replace for the original type differing.
+    The rewriting was developed using a rule based approach, using test driven development (TDD) as this allowed rapid but correct development by continuously being able to check that additional created rules still cause all the tests to pass, causing no regressions.
+    The five cases of code (\emph{or rewriting rules}) that needs to be rewritten are:
+    \begin{itemize}
+        \item Declaration types.
+        \item Variable initialisation and assignment expressions.
+        \item Pointer delete statements.
+        \item Function return types.
+        \item Pointer Usages.
+    \end{itemize}
+
+    Rewriting declaration types are the easiest of all. We can directly replace the type of the variable declaration node that we already have access to. 
+    For example for \mintinline{c++}{int *ptr}, we can extract the type of the variable, and format it into the string \mintinline{c++}{"std::shared_ptr<{}>"} or \mintinline{c++}{"std::unique_ptr<{}>"}, where the \texttt{\{\}} represent the replacement field.
+
+    Initialisation and assignment expressions are less simple due to two reason.
+    First of all, the initialisation of a variable does not necessarily have to occur immediately at the point of declaration.
+    Therefore we need to find all the points of initialisation or value assignment to a variable in the AST.
+    Secondly, there are many different ways to initialise a variable~\cite[p.~196]{C++17}, with the most common initialisation methods using:
+    \begin{itemize}
+        \item Braced-lists \mintinline{c++}{std::string s{}}.
+        \item Equal sign \mintinline{c++}{std::string s = "hello"}
+        \item parentheses expression list \mintinline{c++}{std::string s("hello")} 
+    \end{itemize}
+    For pointer variables, this is convoluted even more depending on if the pointer gets initialised using an existing pointer \mintinline{c++}{int* a = nullptr; int *b = a;}, or if it gets initialised using a new statement \mintinline{c++}{auto p = new Circle{Point{0,0},10}}.
+    In the case of a new statement, we can initialise the rewritten smart pointer using a value initialiser \mintinline{c++}{std::make_unique<>()} or the shared pointer equivalent. 
+    Otherwise, we can create a smart pointer from a pointer value using by the default equal sign copy initialiser or parentheses expression list.
+
+    Rewriting pointer deletion statements also require us to traverse the AST to find \mintinline{c++}{delete} statements for the variable declaration node that we are rewriting.
+    However, they do not require any complicated logic, and we can upon finding relevant statements remove them from the source code.
+
+    If the variable declaration is a variable that gets returned at the end of the function, then the return type of the function signature will need to be rewritten too. 
+    We explicitly check for each variable declaration that we handle if it gets returned, and if so, we can rewrite the function return type the same way that we handle variable declaration types.
+
+    Lastly, smart pointer have overloaded operators for the arrow \texttt{->} and dereference \texttt{*} operators so that they behave identically as if the value was a raw pointer.
+    This simplifies our task as it reduces the amount of rewriting we have to do for each usage of a specific pointer value.
+    However, there is no such overloaded operator for passing smart pointer as function arguments. 
+    In this case, the compiler will find a type error, when it might expect a type such as \mintinline{c++}{int *}, but actually receives \mintinline{c++}{std::unique_ptr<int>}. 
+    Therefore, we again have to traverse the AST to find all function call expressions which involve the variable declaration that we are currently handling, and we need to add \mintinline{c++}{get()} calls to the smart pointer values to get a raw pointer to the managed objects.
+
+    We use the Clang \emph{Rewriter}\footnote{\url{https://clang.llvm.org/doxygen/classclang_1_1Rewriter.html}} class as a high level abstraction compared to handling files and text buffers ourselves, which would likely be error-prone for buffer overflows.
+
+    \begin{figure}
+        \centering
+        \begin{minted}{c++}
+        char *test() {
+            char* a = new char[2];
+            a[0] = 'a';
+            printf("\%s", a);
+            return a;
+        }
+        \end{minted}
+
+        \includegraphics{images/arrow.pdf}
+
+        \begin{minted}{c++}
+        std::shared_ptr<int> test() {
+            std::shared_ptr<char> a = std::shared_ptr<char>(new char[2]);
+            a.get()[0] = 'a';
+            printf("\%s", a.get());
+            return a;
+        }
+        \end{minted}
+
+        \caption{An example of all the rewriting rules except pointer delete statements.
+This example solely shows rewriting and does not necessarily follow respect any smart pointer viability analyses.
+We see how both the declaration and equal sign copy initialisation of \texttt{a} have been rewritten to into a \texttt{shared\_ptr}.
+All the uses of \texttt{a}  that require a raw pointer have a \texttt{get()} call added to obtain the managed memory.}
+        \label{fig:rewriting}
+    \end{figure}
+
+    In \autoref{fig:rewriting}, we show an example that incorporates all the cases in which code needs to be rewritten except pointer delete statements.
+    It is visible how in the majority of cases, the first two rules will be applicable as every new smart pointer need to have the correct type and be initialised to the right type of object. 
+    The fifth rule will also likely be applicable in most cases as variables tend to not be unused and variables which were pointers will be used in functions expecting pointers, as they otherwise could be stack variables.
+
+    \subsection{System Design}\label{subsec:methodology-systemdesign}
+
+    \begin{figure*}
+        \centering
+        \includegraphics[scale=0.65]{images/design.pdf}
+        \caption{System design diagram of the ptr-tidy tool.}
+        \label{fig:system-design}
+    \end{figure*}
+
+    We show the full system design diagram in \autoref{fig:system-design}, showing the full sequence of operations that are executed in order to transform and refactor the input file containing raw pointers into the output file that uses smart pointers.
+    The input file is parsed into a Clang AST, which is traversed to find variable declarations.
+    The corresponding variable declaration is found in the IR, and used as input to the analysis algorithm.
+    The output of the analysis is used in the rewriting component, after which all the changes made in rewriting the program are consolidated and output as the new, improved program.
+
+    We believe this design is noteworthy as it closely follows the system design of the general LLVM infrastructure in \autoref{fig:llvm-design}, by having a Clang parser front-end, an analysis middle-end, and a rewriter back-end.
+    We also roughly follow the pipeline architectural design, as described by \citet{Schmidt2013}, by having a series of transformation steps, which each have an input and one or more outputs.
+    A series of transformation steps are composed until we obtain our end output file.
+    This allows us to conform to the Single-responsibility principle as proposed by \citet{Martin2002}.
+    This gives us a highly flexible design where different analysis algorithms can be very easily added and trialled for the analysis component, meaning that potentially more powerful analyses can be tested without needing any changed to the parsing or rewriting components.
+    Similarly, given that our analysis is executed on the language-agnostic LLVM IR, the analysis component could be reused by rejuvenation tools for any other languages, allowing our work to be used in future research.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    Through our research, we would like to answer the question of whether our implemented tool can truly improve C++ code by seeing if the tool can be used outside small examples on larger existing programs, as well as if the refactored results produce code which can be compiled and can be judged to be better.
+    Secondly, if the tool is able to do this, we want to evaluate how well the tool performs by analysing the effectiveness of the analysis algorithms.
+
+    In order to evaluate the objectives above, hypotheses were made:
+
+    \begin{itemize}
+        \item \textbf{Hypothesis 1} The automatic translation of raw pointers into smart pointers can be used to make existing C++ code more modern, safe and extendable.
+        \item \textbf{Hypothesis 2} The tool can reliably identify situations in which unique pointers should be used.
+    \end{itemize}
+
+    Each of these hypotheses will require experimentation to be verified.
+    We will answer hypothesis 1 through a case study and we will answer hypothesis 2 through the gathering of relevant metrics on several more projects.
+    This section will detail the justification, process, and results for these experiments.
+
+    We run our experiments on a Windows 10 Pro x64 machine using the Windows Subsystem for Linux, with an AMD Ryzen 5 PRO 4650U and 16 GiB RAM, using TinyXML-2 v8.0.0.
+    The ptr-tidy tool was compiled with GCC v10.2.0 using CMake release profile, using Clang and LLVM v11.1.0 libraries.
+
+    \subsection{TinyXML-2 Case Study}\label{subsec:case-study}
+
+    In order to answer hypothesis 1, it was decided to run a case study instead of applying the tool in large on a range of open source projects. 
+    This was decided as an analysis into a well chosen target project will be able to show relevant success and failure cases that we can extrapolate onto more general C++ projects of the same kind. 
+    If the impact of the success cases outweighs the the faults of the failure cases, then we will be able to judge the our hypothesis as being true.
+
+    We decided on using the \emph{TinyXML-2} library as our chosen case study~\footnote{\url{https://github.com/leethomason/tinyxml2}}.
+    TinyXML-2 is a C++ XML parsing library that can parse an XML document and build from that a Document Object Model that can be read, modified, and saved.
+    TinyXML-2 has over three thousand \emph{GitHub stars} which puts it in the top 1\% of projects on the hosting website, which can be used as a rough metric for its popularity and good code quality~\cite{Sanatinia2016}.
+    It is also for example used in the \emph{libigl}~\cite{Libigl} geometry processing library,  which itself is used by a variety of large industry companies.
+    It is released under a \emph{zlib} license~\footnote{\url{https://opensource.org/licenses/Zlib}} which allows us to use and alter the software.
+    What makes library a particularly attractive choice is the simplicity of the library.
+    The library does not make use of exceptions, RTTI or the STL~\cite{C++17}, and therefore most importantly does not make use of smart pointers. 
+    The library is also around three thousand lines of code in length, making it large enough to likely have a variety of edge cases, whilst remaining small enough for manual inspection.
+
+    \begin{listing}
+        \inputminted{cpp}{code/tinyxml2-example.cpp}
+        \caption{Example of a method in the TinyXML-2 library which could be refactored to use smart pointers as it returns an owning pointer to a newly created object.}
+        \label{lst:tinyxml-refactor-example}
+    \end{listing}
+
+    In particular, what makes this suitable for the target of our case study over perhaps other similarly small and simple libraries is that due to parsing varying documents at runtime, the library is particularly heavy in pointer logic and memory allocations.
+    We can in \autoref{lst:tinyxml-refactor-example} an example of a method from the library which would be well suited for refactoring to use smart pointers.
+    The heavy use of pointers and memory allocation is what leads us to believe that the case study into this library will generalise well onto all other C++ projects.
+
+    \subsubsection{Experiment}
+
+    \begin{listing}
+        \inputminted{json}{code/compile_commands.json}
+        \caption{Compilation Database for TinyXML-2}
+        \label{lst:tinyxml-compilation-database}
+    \end{listing}
+
+    We first generate a simple \emph{compilation database}~\footnote{\url{https://clang.llvm.org/docs/JSONCompilationDatabase.html}}, as seen in \autoref{lst:tinyxml-compilation-database}, for the TinyXML-2 library, as this is what our tool uses to resolve compilation options.
+    More complex libraries or applications would have their compilation options and any include and link libraries included in compilation command.
+    The user of our tool can easily generate a compilation database for the project that they wish to rejuvenate using their build tool, such as \emph{CMake}.
+
+    We can then run the tool on any file which has a compilation command defined in the compilation database, such as the \texttt{tinyxml2.cpp} file from our chosen library.   
+    We include the result of the analysis as debug messages during the running of the tool, in order to aid our evaluation of the case study, such as:
+    \begin{minted}{text}
+    </home/artem/tinyxml2/tinyxml2.cpp:207:70, col:75> Variable curLineNumPtr escapes
+    </home/artem/tinyxml2/tinyxml2.cpp:213:5, col:19> Variable start does not escape
+    \end{minted}
+
+    The final rejuvenated version of the TinyXML-2 is included as reference~\footnote{\url{https://gist.github.com/a-usov/155e55afc5d375070c52ad3e4da19072}}.
+
+    \subsubsection{Results and Discussion}
+
+    \begin{listing}
+        \centering
+        \begin{minted}{c++}
+        std::shared_ptr<XMLText>  XMLDocument::NewText( const char* str )
+        {
+            std::shared_ptr<XMLText>  text = std::shared_ptr<XMLText>( CreateUnlinkedNode<XMLText>( _textPool ));
+            text->SetValue( str );
+            return text;
+        }
+        \end{minted}
+        \caption{Rejuvenated version of the method in \autoref{lst:tinyxml-refactor-example}.
+The \texttt{text} variable has been identified as being eligible to be a \texttt{shared\_ptr}.
+The first, second and fourth of our rewriting rules are applied in this example to the \texttt{text} variable.}
+        \label{lst:tinyxml-refactor-example-refactored}
+    \end{listing}
+
+    We can first and foremost see in \autoref{lst:tinyxml-refactor-example-refactored} the result of the rejuvenation our candidate method in \autoref{lst:tinyxml-refactor-example}.
+
+    We can see that we achieve a similar refactoring as in \autoref{fig:rewriting}. 
+    The smart pointer is initialised using using the pointer copy initialiser.
+    We see identical refactoring in the other \texttt{XMLDocument:New...} methods.
+
+    \begin{figure}
+        \centering
+        \begin{minted}{c++}
+        XMLAttribute* XMLElement::CreateAttribute()
+        {
+            TIXMLASSERT( sizeof( XMLAttribute ) == _document->_attributePool.ItemSize() );
+            XMLAttribute* attrib = new (_document->_attributePool.Alloc() ) XMLAttribute();
+            TIXMLASSERT( attrib );
+            attrib->_memPool = &_document->_attributePool;
+            attrib->_memPool->SetTracked();
+            return attrib;
+        }
+        \end{minted}
+
+        \includegraphics{images/arrow.pdf}
+
+        \begin{minted}{c++}
+        std::shared_ptr<XMLAttribute>  XMLElement::CreateAttribute()
+        {
+            TIXMLASSERT( sizeof( XMLAttribute ) == _document->_attributePool.ItemSize() );
+            std::shared_ptr<XMLAttribute>  attrib = std::make_shared<XMLAttribute>();
+            TIXMLASSERT( attrib );
+            attrib->_memPool = &_document->_attributePool;
+            attrib->_memPool->SetTracked();
+            return attrib;
+        }
+        \end{minted}
+        \caption{Rejuvenation of the \texttt{CreateAttribute} method.
+The \texttt{attrib} variable is copy initialised using the default constructor rather than a copy initialisation using an already allocated pointer.
+Therefore we identified that the new smart pointer can be initialised using the \texttt{make\_shared} function.}
+        \label{fig:rewriting-example2}
+    \end{figure}
+
+    The only case where our tool was able to successfully refactor the TinyXML-2 library to use the recommended and more efficient \texttt{make\_shared}~\footnote{\url{https://isocpp.github.io/CppCoreGuidelines/CppCoreGuidelines\#Rr-make_shared}} function is shown in \autoref{fig:rewriting-example2}.
+    We see the reason for this being that the TinyXML-2 uses memory allocation abstractions such as memory pools rather than creating objects directly through their constructors. 
+    Such more detailed refactoring would require human effort.
+
+    \begin{figure}
+        \centering
+        \begin{minted}{c++}
+        void nonescapingFunc(int *p) {
+            *p += 99;
+        }
+        \end{minted}
+        \begin{minted}{llvm}
+        ; Function Attrs: noinline nounwind optnone sspstrong uwtable
+        define void @_Z15nonescapingFuncPi(i32* %p) #0 {
+        entry:
+          %p.addr = alloca i32*, align 8
+          store i32* %p, i32** %p.addr, align 8
+          %0 = load i32*, i32** %p.addr, align 8
+          %1 = load i32, i32* %0, align 4
+          %add = add nsw i32 %1, 99
+          store i32 %add, i32* %0, align 4
+          ret void
+        }
+        \end{minted}
+        \caption{C++ and equivelent LLVM IR snippet showing how function arguments are not considered \texttt{noescape} or \texttt{noalias}}
+        \label{fig:function-arguments}
+    \end{figure}
+
+    Not a single function argument was refactored to use smart pointers, even though arguments are included in the analysis of the program, since the \texttt{VarDecl Matcher} also matches on arguments (\texttt{ParmVarDecl}) as it is a subclass~\footnote{\url{https://clang.llvm.org/doxygen/classclang_1_1ParmVarDecl.html}}.
+    Upon investigation, this remains largely an issue due to poor memory aliasing guarantees in the language~\cite{C++17}. 
+    Parameters are by default always not marked \texttt{noescape} and \texttt{noalias}~\footnote{\url{https://clang.llvm.org/docs/AttributeReference.html\#noalias}} in the generated LLVM IR, as can be seen in \autoref{fig:function-arguments}.
+    If we enable Clang compiler optimisations when generating the IR, the compiler correctly identify that \texttt{p} cannot escape within the function, giving it a signature in the IR of 
+    \mintinline{llvm}{
+    (i32* nocapture %p)
+    }.
+    However this would break the AST to IR correspondence that we established in \autoref{subsec:methodology-analysistarget} as variables that exist in the AST may find themselves optimised and removed from the IR or functions can that exist in the AST may find themselves inlined in the IR.
+    Also, we still do not have non-aliasing guarantees so that we know that the object has not escaped before entering the function.
+    What is possible is that we can manually mark a variable in the code as being non-capture and non-alias, such as:
+    \begin{minted}{c++}
+    void nonescapingFunc(__attribute__((noescape)) int * __restrict__ p) {
+        *p += 100;
+    }
+    \end{minted}
+    This relies however on the Clang compiler specific implementation of the \texttt{noescape} attribute which does not exist in the GCC compiler, as well as the non-standard, compiler specific implementation of the \texttt{restrict} attribute.
+    Such code therefore is not portable and not standard conforming, making it very unlikely unacceptable for most projects.
+    This last unlikely case is not currently correctly handled by the LLVM escape analysis to produce a more correct refactoring. 
+    The automatic generation of such front-end code however could be possible through the integration of Clang compiler optimisations and related research such as by \citet{Campos2015} for integration into our tool for better performance.
+    The impact this on our tool is that the extent of our analysis is not as complete as we would like it to be, however \autoref{lst:tinyxml-refactor-example-refactored} and \autoref{fig:rewriting-example2} show that we can still achieve desirable results through the analysis of local variables.
+
+    The tool is particularly effective at refactoring factory methods~\cite{Gamma1995} such as in \autoref{lst:tinyxml-refactor-example-refactored} and other cases where resources are created at the top rather than at the bottom of the function call stack and passed through return values.
+
+    The rejuvated version of the TinyXML-2 library that the tool produces is not directly compilable into an object file, mainly due to edge cases in the rewriting component. 
+    These mainly stem from outdated C++ practises that must be explicitly handled in modern code that has more strict type checking. 
+    For example, the initialisation of a pointer using the value \texttt{0}, such as
+    \mintinline{c++}{XMLNode* returnNode = 0;} is not recommended as they should be initialised with \mintinline{c++}{nullptr}.
+    The rejuvenation of the example into
+    \mint{c++}|std::shared_ptr<XMLNode>  returnNode = std::shared_ptr<XMLNode>(0);|
+    therefore causes a type error during compilation.
+
+    All of the compilation issues currently stem from unimplemented behaviour in the rewriting component rather than the strength or performance of the analysis.
+    The size and complexity of the language makes it difficult to ensure that all edge cases are handled by the rewriting component, which can in future development be continuously improved.
+    The fixes that need to be applied can be done by the programmer, with all of the identified fixes in the TinyXML-2 library being a single line in length.
+
+    We do not think this however detracts from the value or usefulness of our tool.
+    Our tool provides fully rejuvenated and compilable programs without any human input needed for simpler inputs, using modern styles of C++.
+    For larger, more complex, older input programs some human input is currently still needed, however the work of our tool still allows the programmer to make their programs more modern and safer in by performing the complex analysis and guiding the programmer towards the needed fixes through compiler errors.
+    This results in the rejuvenated output being obtained easier by removing the need for the programmer to manually understand the structure and style of the programmer and the often complex ownership relations between the various variables in the program.
+
+    We therefore believe that the analysis of our tool on the case study of the TinyXML-2 library, even with the identified failure cases, the success of the automated translation and identified success cases allow us to confirm our first hypothesis.
+
+    \subsection{Analysis of performance on a corpora of projects}
+
+    In order to answer hypothesis 2, it was decided to run analysis of the tool on more various open source projects and gather metrics on the performance of the tool in terms of its performance in successfully identifying pointers, how extensive the rejuvenation in context of the whole program, and some more general runtime characteristics.
+
+    It was decided that whilst gathering more projects to test, we would continue to focus on simple, single file libraries such as TinyXML-2 in order to ease the cost of setting up each project to be used by the tool.
+    We used a list of single file, public-domain/open source libraries with minimal dependencies~\cite{Barrett2019} by acknowledged game and software developer Sean Barret~\footnote{\url{https://www.mobygames.com/developer/sheet/view/developerId,4966/}} to discover as many as possible candidate libraries.
+    This list also ensures that we only select projects which we are able to use and modify as well as including project from many areas of software engineering such as graphics and geometry to file parsers.
+    This ensures that we run the tool across a wide variety of projects which will have different distributions and usages of memory allocations and pointer use, giving greater validity that our tool is able to generalise to any type of project.
+    We filtered projects from the list that are C++ only, do not already use smart pointers and have at least one dynamic memory allocation and are not header-only.
+    With this, we arrived at the following list of selected projects:
+
+    \begin{itemize}
+        \item \textbf{Clipper} for clipping and offsetting lines and polygons~\footnote{\url{http://www.angusj.com/delphi/clipper.php}}.
+        \item \textbf{HappyHTTP} for issuing HTTP requests and processing responses~\footnote{\url{http://scumways.com/happyhttp/happyhttp.html}}.
+        \item \textbf{MicroPather} for path finding and A* solving.~\footnote{\url{http://www.grinninglizard.com/MicroPather/}}.
+        \item \textbf{Jzon} for JSON parsing~\footnote{\url{https://github.com/Zguy/Jzon}}.
+        \item \textbf{Poisson Disk Points Generator}~\footnote{\url{https://github.com/corporateshark/poisson-disk-generator}}.
+        \item \textbf{PolyPartition} for polygon partition and triangulation~\footnote{\url{https://github.com/ivanfratric/polypartition}}.
+        \item \textbf{TinyXML-2} for XML parsing~\footnote{\url{https://github.com/leethomason/tinyxml2}}.
+        \item \textbf{Xatlas} for generating unique texture coordinates suitable for baking lightmaps or texture painting~\footnote{\url{https://github.com/jpcy/xatlas}}.
+    \end{itemize}
+
+    We first modified the tool to include a simple counter for how many pointers that are analysed are identified as escaping and how many as non-escaping.
+    We also disabled any output of debug messages in order to not skew the tool runtime with writing large amounts of text, other than the output rejuvenated program, as this is the essential function of the tool.
+    We measure the runtime of the tool on each project as the median of five runs, on a machine with no other programs running, recording the pointer analysis statistics as well.
+    We measure the number of lines changed by using the output of the Linux \texttt{diff} command on the rejuvenated and original files. 
+    We show the raw data of the gathered metrics in \autoref{table:data}.
+
+    \begin{table*}
+    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
+    \hline
+                                   & \textbf{clipper} & \textbf{happyhttp} & \textbf{jzon} & \textbf{micropather} & \textbf{poisson} & \textbf{polypartition} & \textbf{tinyxml2} & \textbf{xatlas} \\ \hline \hline
+    \textbf{Runtime (ms)}          & 146.521          & 10.540             & 3.140                & 2.429         & 4.471            & 11.885                 & 16.055            & 105.212         \\ \hline
+    \textbf{Escaped Pointer}       & 94               & 22                 & 2                    & 39            & 7                & 46                     & 164               & 235             \\ \hline
+    \textbf{Non-Escaped Pointers}  & 132              & 16                 & 3                    & 34            & 3                & 22                     & 84                & 111             \\ \hline
+    \textbf{Lines of Code Changed}   & 205             & 27                 & 3                    & 57            & 6                & 31                     & 142               & 154             \\ \hline
+    \textbf{Total Lines of Code} & 5035             & 1273               & 1324                 & 1587          & 523              & 2270                   & 5366              & 10318           \\ \hline
+    \end{tabular}    
+    \caption{Data gathered from the rejuvenation of each of these projects.
+Total lines of code include both the implementation and header file of each project.
+Runtime measured in milliseconds as the median of five runs.}
+    \label{table:data}
+    \end{table*}
+
+    \begin{figure*}
+        \centering
+        \includegraphics[scale=0.55]{images/pointers.png}
+        \caption{First graph shows the fraction of non-escaped, and therefore refactored, pointers for various projects.
+This varies from a minimum of 30\% up to a maximum of 60\%, with a mean of 42\%.
+The second graph shows the number of total pointers compared to the number of total lines in each project.}
+        \label{fig:ratio-pointer}
+    \end{figure*}
+
+    \begin{figure*}
+        \centering
+        \includegraphics[scale=0.8]{images/lines.png}
+        \caption{Graph showing the fraction of modified lines by the tool compared to the total number of lines in each project, which is the combination of implementation and header files.
+This varies from a minimum of 0.22\% up to a maximum of 4.07\%, with a mean of 2.18\%.}
+        \label{fig:lines}
+    \end{figure*}
+
+    \begin{figure*}
+        \centering
+        \includegraphics[scale=0.55]{images/time.png}
+        \caption{Graphs of the tool runtime versus the total lines of code in each project and the total number of pointers that was analysed.
+The x axis for both graphs are plotted on a log scale.
+This is for clarity purposes in order to emphasise the two outliers consisting of the Clipper and Xatlas libraries.
+Time was measured in microsecond using the Bash \texttt{time} command and measuring the median real execution time of 5 runs.}
+        \label{fig:time}
+    \end{figure*}
+
+    \subsubsection{Results and Discussion}
+
+    We see first of all in \autoref{fig:ratio-pointer} that our tool has consistent behaviour across the various projects, with 4 projects having between 30\% and 40\% of pointers identified for refactoring and 4 other projects having between 40\% and 60\% of pointers identified.
+    The mean of 42\% is shows that we are able to identify a non-trivial amount of pointers, however it is unclear this is not higher because of a fault in the analysis or whether this is the typical performance for a conservative analysis.
+    The second graph in \autoref{fig:ratio-pointer} also shows how the amount of pointers in a project as the size of the project increases is close to linear. 
+    This is a surprising result as we expected that different project in different areas of computer science of the same length would utilise vastly different amounts of memory allocations and use vastly different amounts of pointers.
+    There seems to be no relationship between the 4 projects that had over 40\% of pointers identified and their lengths.
+    We therefore conclude that the length of a project that the tool has to rejuvenate has no impact on its performance.
+
+    \autoref{fig:lines} the number of lines that were modified in the rejuvenated version of the project compared to the original. 
+    We can much more clearly see our previous expectation that some projects utilise pointer more and other less. 
+    The two projects with the lowest amount of modified lines, Jzon and the Poisson Disk Points Generator both only had ten or less pointers to analyse in total, with the Jzon library in particular only having three changed lines from the original.
+    Manual investigation of the Jzon library shows that even though it is a parsing library like TinyXML-2, the amount of code limited to memory allocations and pointers is very minimal. 
+    We therefore again conclude that it seems like the tool is able to generalise well to different input programs.
+
+    Lastly, \autoref{fig:time} shows two graphs, plotting the runtime of the tool on the selected projects versus the lines of code and number of pointers that they contain.
+    Unlike usual refactoring tools, our tool as a rejuvenation tool is only intended to be run once on a code base, compared to a tool such as \texttt{clang-tidy}~\footnote{\url{https://clang.llvm.org/extra/clang-tidy/}} which is usually invoked many times.
+    Therefore, we do not have a strict requirement on a fast processing time. 
+    However, we can see that for the majority of projects, the processing remains under twenty milliseconds, even as the number of lines of code increases to over 5000. 
+    We see a linear relationship in both of the graphs, except for the outliers of the Clipper and Xatlas libraries.
+    We were unable to establish why these libraries take an exceptionally longer time to process than the TinyXML-2 library which has the same amount of lines of code and analysed pointers as the Clipper library.
+
+    To conclude, we believe we can partially confirm our second hypothesis.
+    We show that the tool generalises to a range of other libraries apart from the TinyXML-2 library selected in our case study and show that our analysis achieves similar results across these libraries.
+    However, what we cannot conclude is the accuracy of our tool in terms of the best case performance. 
+    We were not able to find in the literature an estimate for the fraction of code that is used for memory management, nor are we able to confirm that our analysis achieves the best possible result out of any possible analysis that could be implemented.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    In this paper, we present the problem of \emph{manual memory management} in C++ and how it can create the opportunity for exploitable \emph{memory safety errors}. 
+    We show how modern C++, that is the C++ 11 standard and newer, encourage the use of \emph{smart pointers} instead for memory management. 
+    Smart pointer create memory or resource handles which automatically get released when the handle exists the scope of a function or method. 
+    This referred to as \emph{RAII} (\textit{Resource Acquisition is Initialisation}).
+    We also highlight the eventual increasing need for automatic rejuvenation, the replacement of outdated coding patterns with newer, higher-level abstractions, of maintained projects as languages move and evolve much quicker than can be supported in enterprise environments.
+    The learning and application of new language requires a significant time investment that may not be able to be allocated.
+
+    We then introduce our \emph{ptr-tidy} tool which was designed automatically rejuvenate code that does not use smart pointers, thereby solving our first problem.
+    We show the design and implementation of the tool, which uses the Clang and LLVM libraries. 
+    The tool uses Clang and the Clang abstract syntax tree (AST) in order to obtain a rich and closely resembling representation of an input program.
+    The tool then uses the equivalent representation of the input program in the lower-level LLVM intermediate representation (IR) in order to analyse which pointers are suitable to be \emph{unique pointers} or \emph{shared pointers}.
+    The AST and output of the analysis are then input to a rewriting component which creates and outputs our final rejuvenated program. 
+    This design is the solution to our second problem, as we believe this design is both generic, reusable and powerful enough to be used for a variety of same language refactoring or rejuvenation or different language transpiling.
+    The design of the LLVM IR allows for powerful and efficient analyses to be performed, whilst its language agnostic nature allows a variety of possible parsing front-ends and output back-ends to be used, such as existing mature implementations available for C++ or Rust.
+
+    Our evaluation showed that our tool can effectively be used to rejuvenate a set of simple open-source libraries, with a specific in depth analysis into the TinyXML-2 library that showed correct rejuvenations in object factory methods.
+    We observe that the tool is particularly effective at rejuvenating specific cases in programs where dynamic resources are created at the top of the function call stack and passed down the stack through return values.
+    We also discovered that due to limitations in the specification of the C++ language, we are unable to rejuvenate function parameters statically, which meant the extend of our rejuvenation had limits.
+    The successful application of the tool across a varied set of open-source projects allowed us to conclude that the analysis and design of the tool were indeed suitable for refactoring generic C++ programs.
+
+    \subsection{Future Work}
+
+    The escape analysis algorithm described by \citet{Choi1999} is used as the default escape analysis algorithm in the Oracle Java Virtual Machine implementation~\footnote{\url{https://docs.oracle.com/en/java/javase/16/vm/java-hotspot-virtual-machine-performance-enhancements.html}} and is widely cited~\cite{Wang2020, Stadler2014}.
+    Most importantly, the authors present an interprocedural analysis which we hypothesise could bring greater effectiveness to our tool and therefore we believe that it could lead to improved analysis performance compared to our implemented intraprocedural analysis.
+    The similarity of Java and C++ as object oriented languages allowed for a direct application of the algorithm to C++.
+
+    Secondly, we mention in \autoref{subsec:case-study} that we are unable to enable optimising compiler invocations for generating the LLVM IR as it breaks our AST to IR correspondence. 
+    However, it remains to investigate whether these optimisations can be called dynamically during the running of our tool in order to gain information about conservative assumptions that we currently make.
+
+    In terms of the evaluation of the tool, we would like to extend it with more complex, multi-file projects and determine whether the translation unit barrier, where an analysis pass can only access the current implementation file, has an impact on analysis performance.
+    An inter-modular analysis, similar to existing LTO optimisations in compilers~\footnote{\url{https://llvm.org/docs/LinkTimeOptimization.html}}, can then be evaluated whether it brings greater analysis performance.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
     \vskip8pt \noindent
-    \textbf{Acknowledgments.}
+    \textbf{Acknowledgements.}
+
+    I would like to thank my supervisor Dr Jeremy Singer for his guidance and feedback throughout the duration of the project.
+    I would also like to thank my family for their continued support.
 

### 2021-04-11 code/fix-leak.cpp

@@ -1,0 +1,8 @@
+void foo(int x) {
+    std::unique_ptr<Circle> p = std::make_unique<Circle>(Point{0,0},10);
+    // ...
+    if (x<0) throw Bad_x{}; // delete p inserted
+    if (x==0) return; // delete p inserted
+    // ...
+    // delete p inserted
+}

### 2021-04-11 code/leak.cpp

@@ -1,0 +1,8 @@
+  void foo(int x) {
+    Shape *p = new Circle{Point{0,0},10};
+    // ...
+    if (x<0) throw Bad_x{}; // potential leak
+    if (x==0) return; // potential leak
+    // ...
+    delete p;
+}

### 2021-04-11 code/tinyxml2-example.cpp

@@ -1,0 +1,6 @@
+XMLText* XMLDocument::NewText( const char* str )
+{
+    XMLText* text = CreateUnlinkedNode<XMLText>( _textPool );
+    text->SetValue( str );
+    return text;
+}

### 2021-04-13 paper.tex
Go over finished draft

@@
     \begin{abstract}
-    Systems programming involves dynamically requesting memory that in the C++ language has to be manually managed by the programmer. 
+    Systems programming involves dynamically requesting memory which in the C++ language has to be manually managed by the programmer. 
     However, this leaves the opportunity for exploitable memory safety errors.
-    Modern C++ recommends that smart pointer are used to automatically manage resources such as dynamic memory.
-    However, many legacy programs may have been written before these features were introduces to the language.
-    We developed the ptr-tidy tool in order to automatically rejuvenate such programs to use smart pointers.
-    The tool uses the Clang and LLVM libraries in order to create a componentised and powerful design which could be reused by other refactoring or rejuvenation tools. 
-    We present a case study and and apply the tool on a number of open-source C++ libraries to show that the tool can successfully rejuvenate non-trivial libraries and that the tool can be used on libraries from many areas of computer science.
-    We discuss in which circumstances the tool performs well and the current limitations.
+    Modern C++ recommends that smart pointers are used to automatically manage resources such as dynamic memory.
+    However, many legacy programs may have been written before these features were introduced to the language.
+    We develop the ptr-tidy tool in order to automatically rejuvenate such programs to use smart pointers.
+    The tool uses the Clang and LLVM libraries in order to create a componentised and powerful design which could be used by other refactoring or rejuvenation tools. 
+    We present a case study and also apply the tool on a number of open-source C++ libraries from a variety of areas in computer science to show that the tool can successfully rejuvenate non-trivial libraries.
+    We discuss the circumstances in which the tool performs well and the current limitations.
     \end{abstract}
@@
     Systems programming involves memory management, that is dynamically requesting memory from the operating system to be used and managed by the program.
-    This is done when the amount of memory that we need at runtime cannot be know at the time of compilation of the program.
-    However, doing so creates the opportunity for memory safety errors~\cite{Dhurjati2003} which are notoriously challenging to avoid.
+    This is done as the amount of memory that we need at runtime cannot be know at the time of compilation of the program.
+    However, doing so creates the opportunity for memory safety errors~\cite{Dhurjati2003}.
     A memory safety error can be one of:
     \begin{itemize}
-        \item \textbf{Dangling pointers}, prematurely freeing an object that is still in use.
-        \item \textbf{Double free}, deleting an object multiple times.
-        \item \textbf{Never free}, never deleting an object.
+        \item \textbf{Dangling pointers}, prematurely freeing memory that is still in use.
+        \item \textbf{Double free}, freeing memory multiple times.
+        \item \textbf{Never free}, never freeing memory.
     \end{itemize}
@@
 
-    Historically systems programs are mostly written in C, C++ or a combination of the two and these remain the dominant languages.
+    Historically, systems programs are mostly written in C, C++ or a combination of the two and these remain the dominant languages.
     They are lower level languages that involve manual memory management by the programmer.
-    It is even more challenging to ensure that C and C++ programs contain no memory safety errors as evidenced by empirical evidence in industry, where in a presentation at the Linux Security Summit it is shown that in several popular projects such as Firefox, macOS, Ubuntu and Android all had over half of their CVEs attributed to issues with memory safety~\cite{Gaynor2020}.
-    Similarly, in a presentation by Matt Miller, a security engineer at Microsoft, it is shown that around 70\% of their vulnerabilities that are addressed through a security updates are due to memory safety issues~\cite{Miller2019}.
-    This is at industry-leading companies who are renowned for hiring top talent and top developer education programs.
+    It is challenging to ensure that C and C++ programs contain no memory safety errors as evidenced by empirical data in industry, where in a presentation at the Linux Security Summit it was shown that in several popular projects such as Firefox, macOS, Ubuntu and Android all had over half of their CVEs~\footnote{\url{https://cve.mitre.org/}} attributed to issues with memory safety~\cite{Gaynor2020}.
+    Similarly, in a presentation by Matt Miller, a security engineer at Microsoft, it is shown that around 70\% of their vulnerabilities that are addressed through security updates are due to memory safety issues~\cite{Miller2019}.
+    % This is at industry-leading companies who are renowned for hiring top talent and with top developer education programs.
 
@@
 
-    Unfortunately, simply not using these languages is not an option since their unmanaged nature makes them highly performant, and therefore the best option for systems such as a web browser.
+    Unfortunately, simply not using C or C++ is not an option since their unmanaged nature allows for high performance, and therefore the best option for systems with strict or high requirements.
     Engineers at both Microsoft~\cite{Thomas2019} and Mozilla~\cite{Hostfelt2019} converge on Rust~\cite{Balasubramanian2017} as a possible solution.
-    Rust is a systems language that offers similar performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model also guarantee memory safety.
+    Rust is a systems language that offers similar performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model guarantee memory safety.
     The rewriting of a program in a new language, namely from C++ to Rust, is a colossal undertaking, especially given that Rust has a reputation for being difficult to learn~\cite{Astrauskas2019}.
-    Instead, modern C++ guidelines encourage the use of \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create memory or resource handles which automatically eliminate entire classes of memory safety errors.
-    Yet, we encounter further problems as C++ programs either do not actively follow this advice, or are \emph{legacy} programs that were written before these features were introduced to the language.
+    Instead, modern C++ guidelines encourage the use of \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create compiler managed resource handles.
+    Their correct use eliminates the three previously identified classes of memory safety errors.
+    Yet, we encounter further problems as C++ programs either do not actively follow this advice, or are \emph{legacy} programs which were written before these features were introduced to the language.
 
-    We therefore propose a solution in the form of a tool, \emph{ptr-tidy}~\cite{Usov2021}, to analyse and automatically refactor programs to follow the recommended use of smart pointers where the analysis can determine that such a change is syntactically correct and does not change the runtime behaviour of the program.
+    We therefore propose a solution in the form of a tool, \emph{ptr-tidy}~\cite{Usov2021}, to analyse and automatically refactor programs to follow the recommended use of smart pointers where the analysis can determine that such a change is correct and does not change the runtime behaviour of the program.
 
-    \begin{listing}
+    \begin{figure}
+        \centering
         \inputminted{c++}{code/leak.cpp}
+        \includegraphics{images/arrow.pdf}
         \inputminted{c++}{code/fix-leak.cpp}
-        \caption{Example of a memory leak using manual management and how smart pointers correct this.
-Lines 4 and 5 show how interruptions in the program control flow can mistakenly cause memory leaks and how smart pointer correct this by automatically inserting all needed delete statements at compile time.
+        \caption{Example of a memory leak using manual management and how smart pointers correct this problem.
+Lines 4 and 5 show how interruptions in the program control flow can mistakenly cause memory leaks as the \texttt{delete} statement is never reached.
+Smart pointers correct this by automatically inserting delete statements at all necessary locations at compile time.
 All necessary headers are omitted for brevity}
         \label{lst:motivating-example}
-    \end{listing}
+    \end{figure}
 
-    We provide a motivating example in \autoref{lst:motivating-example} where the first version of the function could have several control flow executions where the \texttt{Circle} object will leak.
-    The example seems quite trivial, but can suddenly be quite non-trivial if the conditions in the \texttt{if} statement become calls into external libraries, which may or may not throw an exception.
+    We provide a motivating example in \autoref{lst:motivating-example} where the initial version of the function has several control flow statements that could cause the \texttt{Circle} object to never be freed.
+    The example can seem trivial as all the possible executions of the program are easy to visualise, but can suddenly be rather non-trivial if the conditions in the \texttt{if} statement become calls into external libraries, which may throw an exception and cause the \texttt{delete} statement to again not be reached.
 
@@
     Our work closely relates to work done around the field of \emph{source code rejuvenation}, as defined by \citet{Pirkelbauer2010}. 
-    We exactly aim replace outdated coding patterns with newer, higher-level abstractions, as well as fitting the notion that this tool would be applied once to a codebase, rather than be a reoccurring task.
-    We will however use the terms rejuvenation and refactoring interchangeably in this paper as it is a term that most developers are familiar with.
+    We exactly aim to replace outdated coding patterns with newer, higher-level abstractions, as well as fitting the notion that this tool would be applied once to a codebase, rather than be a reoccurring task.
+    We will however use the terms rejuvenating and refactoring interchangeably in this paper as it is a term that most developers are more familiar with.
 
-    We believe research in this area is going to increase, as languages move and evolve much quicker than can be supported in enterprise environments, resulting in accumulating technical debt. 
+    We believe research in this area is going to increase, as languages move and evolve much quicker than can be supported in enterprise environments~\cite{Overbey2009, Favre2005}, resulting in accumulating technical debt.
     This is supported by the fact that other relevant work has been completed on this subject.
 
-    First and foremost, we have evidence that this is a problem that exists in industry through a paper by \citet{Wright2013} which describes a tool used at Google for running such rejuvenations across a large codebase.
+    First and foremost, we have evidence that this is a problem that exists in industry through a paper by \citet{Wright2013} which describes a tool used at Google for running such rejuvenations across a large C++ codebase.
     \citet{Wright2013} show an example of running a simple rejuvenation of upgrading an API call.
-    They use Clang for the parsing of code into an AST and traverse the tree to identify and perform simple refactorings.
+    They use Clang for the parsing of code into a syntax tree and traverse the tree to identify and perform simple refactorings.
     Their flexible design allows them to implement many different simple refactorings using the same tool, rather than the focus we have for a single specific refactoring goal.
@@
     Secondly, there has been separate academic research into rejuvenation tools such as the work by \citet{Huck2016} and \citet{Kumar2012}.
-    The former also uses the Clang and AST driven approach like \citet{Wright2013}, while the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
-    We have not evaluated IPR as a representation compared to the Clang AST, but we can however comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation that Clang has, as well as it being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
+    The former also uses the Clang driven approach like \citet{Wright2013}, while the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
+    We have not evaluated IPR as a representation compared to the use of Clang, but we can  comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation for Clang, as well as being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
 
-    There has also been industry interest in automatic rejuvenation tools between languages rather than within the same language. 
-    This is done either to increase the use of a new, improved language by creating creating tools to aid transition, or in order to reuse existing compilers.
+    There has also been industry interest in automatic rejuvenation tools \emph{between} languages rather than \emph{within} the same language. 
+    This is done either to increase the use of a new, improved language by creating tools to aid transition, or in order to reuse existing compilers.
     For example, the Kotlin language team created the J2K~\footnote{\url{https://github.com/JetBrains/kotlin/tree/master/j2k}} transpiler in order to entice Java developers to switch to a language with improved language features such as null safety~\cite{Mateus2019}.
-    Similarly, several new languages such as Typescript and Dart have been recently developed which were designed not to be used by themselves, but to be transpiled to the popular Javascript language.
+    Similarly, several new languages such as Typescript~\footnote{\url{https://www.typescriptlang.org/}} and Dart~\footnote{\url{https://dart.dev/}} have been recently developed which were designed not to be used by themselves, but to be transpiled to the popular Javascript language.
     These languages aim to solve shortcomings of the output language, whilst reusing the existing mature and efficient compilers or interpreters.
@@
 
-    Generally pointers are used to give the program access to a resource that cannot be directly included in the program itself, such as allocated memory or a file.
+    Generally, pointers are used to give a program access to a resource that cannot be directly included in the program itself, such as allocated memory or a file.
     Pointers can also be used to pass large resources to functions without needing to create a copy, which would have a performance impact.
@@
     Pointers can be created for objects both on \emph{the stack and the heap}~\cite{Stroustrup2000}. 
-    The stack is an small memory store automatically managed for us by the compiler and is suitable for placing small, fixed sized parameters and arguments, such as perhaps loop counter variables.
-    However, when objects must live beyond the lifetime of the current function call or if we do not know the size of an object until runtime, then we must storage these objects on the heap, occasionally called the \emph{free store}.
+    The stack is a small memory store automatically managed for us by the compiler and is suitable for placing object whose size is known at compile time, such as simple integers like loop counter variables.
+    However, when objects must live beyond the lifetime of the current function call or if we do not know the size of an object until runtime, then we must store these objects on the heap, also called the \emph{free store}.
     Unlike the stack, the memory on the heap must be explicitly managed by the programmer.
 
-    However Stroustrup claims that pointers to objects allocated on the free store are dangerous and a \emph{plain old pointer}, or raw pointer as we will refer to them, should not be used to represent ownership~\cite{Stroustrup2018}.
+    However Stroustrup claims that pointers to objects allocated on the free store are dangerous and a plain old pointer, or \emph{raw pointer} as we will refer to them, should not be used to represent ownership of such objects.~\cite{Stroustrup2018}.
 
-    Instead we can use \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create resource handles which automatically eliminate resource leaks with no, or very little added overhead.
+    Instead we can use \emph{smart pointers}~\cite{Dimov2003} and the concept of \emph{RAII (Resource Acquisition Is Initialisation)}~\cite{Stroustrup2000} to create resource handles with no, or very little added overhead.
     RAII means that the management of a resource, namely its acquisition and release, is bound to its lifetime. 
-    In the simplest case, this means that when a resource exists the scope of a function or method, it is released automatically.
+    In the simplest case, this means that when a resource exits the scope of a function or method, it is released automatically.
+    We can see how smart pointers and RAII work in our motivating example in  \autoref{lst:motivating-example}.
+    Even though the \texttt{Circle} object is placed on the heap, the compiler can manage the smart pointer for us so that in all the cases where the lifetime of the object ends, the object will be released.
 
-    The first type of smart pointer are \emph{unique pointers} which have only a single unique owner at any time and therefore are ideal for resources which may not be copied, such as perhaps a global session dictionary.
-    We can see how unique pointers and RAII work from out motivating example in  \autoref{lst:motivating-example}.
-    Even though the \texttt{Circle} object is placed on the heap, the compiler can manage the smart pointer for us so that in all cases where the lifetime of the object ends, the object will be freed and ensures the resource will not leak.
-    The ownership of a unique pointer can be transferred between variables using \texttt{std::move} which moves the ownership from one variable to another, making the original reference invalid.
+    The first type of smart pointer are \emph{unique pointers}, which can have only a single unique owner at any time and therefore are ideal for resources which may not be copied.
+    The resource managed by a unique pointer is released when the current owner exits the scope of the function or method it is contained in.
+    The ownership of a unique pointer can be transferred between variables using \texttt{std::move} which \emph{moves} the ownership from one variable to another, making the original owner invalid.
     % TODO insert example diagram possibly
@@
 
-    As seen in \autoref{sec:background}, much similar research uses Clang and LLVM.
+    As seen from \autoref{sec:related-work}, much similar research uses Clang and thus LLVM.
     LLVM~\cite{Lattner2004} is a compiler infrastructure project consisting of several compiler and toolchain technologies. 
-    LLVM is centrally designed around a language independent intermediate representation (IR), which can be created by a number of language front-ends and is used by a number of architecture back-ends to generate machine code.
-    Program optimisations can then be applied by portable and reusable transformations on the IR.
-    Any improvement in back-end code generation, as well as any new or improved IR optimisers can then benefit any of the supported language front-ends.
-    It is therefore visible that this allows for a very modular and extensible design which can be seen in \autoref{fig:llvm-design}.
+    LLVM is centrally designed around a language independent intermediate representation (IR), which can be created by a number of language front-ends and is used by a number of back-ends to generate machine code.
+    Programs can then be optimised by applying portable and reusable transformations on the IR.
+    Any improvement in back-end machine code generation, as well as any new or improved IR optimisations therefore benefit all of the language front-ends.
+    It is therefore visible that this allows for a very modular and powerful design which can be seen in \autoref{fig:llvm-design}.
 
-    Clang~\cite{Lattner2008} is such an LLVM front-end for C, C++ and a variety of other similar languages such as OpenGL.
+    Clang~\cite{Lattner2008} is a LLVM language front-end for C, C++ and a variety of other extensions such as OpenGL.
     By using the LLVM infrastructure, it allows for faster and more efficient compilation of C++ compared to existing, older compilers such as the GNU Compiler Collection~\footnote{\url{https://gcc.gnu.org/}} (GCC)~\cite{Lattner2008}.
-    However, most importantly for our research, the design of GCC makes it unsuitable for integration with other projects, such as analysis tools, whereas Clang allows access to several analysis tools as well as the Clang abstract syntax tree (AST) and parser internals via a standard API such as consumers and visitors~\cite{Duffy2014}.
+    However, most importantly for our research, the design of GCC makes it unsuitable for integration in other projects such as analysis tools, whereas Clang allows access to several analysis tools as well as the Clang abstract syntax tree (AST) and parser internals via a standard API such as consumers and visitors~\cite{Duffy2014}.
 
@@
         \includegraphics{images/code-to-ast.pdf}
-        \caption{Example of converting C++ code to Clang AST. All diagrams show equivalent representation of a small C++ code snippet.}
+        \caption{Example of converting C++ code to the Clang AST. The Clang AST is rich in information and could be used to reconstruct the original C++ code snippet.}
         \label{fig:code-to-ast-example}
@@
 
-    Further more, Clang also explicitly provides the \emph{LibTooling} C++ library for the creation of tools that leverage the Clang parsing front-end and AST.
-    The Clang AST is different from ASTs produced by some other compilers in that it closely resembles both the written C++ code and the C++ standard~\footnote{\url{https://clang.llvm.org/docs/IntroductionToTheClangAST.html}}.
+    Further more, Clang also provides the \emph{LibTooling} C++ library for the creation of tools that leverage the Clang parsing front-end and AST.
+    The Clang AST is different from ASTs produced by some other compilers in that it closely resembles the written C++ code~\footnote{\url{https://clang.llvm.org/docs/IntroductionToTheClangAST.html}}.
     For example, we can see in \autoref{fig:code-to-ast-example} that the AST stores much syntactic as well as semantic information in the form of source code locations and variable initialisation styles.
@@
 
-    The LLVM IR represents a language-independent, static single assignment (SSA)~\cite{Rosen1988} representation of the program.
-    Variables in in SSA form can only be assigned once, which allows us to efficiently generate compact definition-usage graphs, which enable developers to simplify many dataflow optimizations and enables fast, flow-insensitive algorithms~\cite{Lattner2004}.
+    The LLVM IR is a language-independent, static single assignment (SSA)~\cite{Rosen1988} representation of a program.
+    Variables in SSA form can only be assigned once, which allows us to efficiently generate compact definition-usage graphs, which enable developers to simplify many dataflow optimizations and enables fast, flow-insensitive algorithms~\cite{Lattner2004}.
+
+    \vskip50pt
 
@@
 
-    The first capability our tool would need is to be able to parse C++ code.
-    Doing this ourselves would be an undertaking that would take a significant amount of time in order to create a complete parser due to the size and complexity of the C++ language, therefore it was decided early to use an existing parser.
-    The information in \autoref{subsec:clang-and-llvm} outlines why Clang is the most suitable parser to use, with the other benefit that it brings such as \emph{LibTooling}.
+    The first capability our tool needs is to be able to parse C++ code.
+    Doing this ourselves would be an undertaking that would take a significant amount of time in order to create a complete parser due to the size and complexity of the C++ language.
+Therefore it was decided early to use an existing parser.
+    The information in \autoref{subsec:clang-and-llvm} outlines why Clang is the most suitable parser to use, with the other benefits that it brings such as the \emph{LibTooling} library.
 
@@
 
-    Next, our tool would need to be able to identify specifically which pointers are suitable for refactoring.
-    An \emph{unique pointer} is a smart pointer object that represent a single unique owner of a region of memory at any time. 
-    We will therefore need to develop a static analysis algorithm to identify the maximum number of owners a region of memory has throughout all points of a program. 
-    If a memory region has at most a single unique owner at a time, we can refactor the initialising raw pointer to a unique pointer, otherwise the pointer can be refactored to a shared pointer.    
+    Next, our tool would need to be able to identify specifically which pointers are suitable for refactoring into smart pointers.
+    An \emph{unique pointer} is a smart pointer that represents a single unique owner of a region of memory at any point in time.
+    We will therefore need to develop a static analysis algorithm to identify the maximum number of owners a region of memory has throughout all points of the program. 
+    If a memory region has at most a single unique owner, then we can refactor the initialising raw pointer to be a unique pointer, and otherwise to a shared pointer.    
     This analysis will need to ensure that its result is correct, and will inevitably have to make some conservative assumptions about the program.
@@
 
-    A key intuition is that a region of memory, subsequent to its initialisation, is only able to gain additional owners if access to the region of memory is \emph{shared} to another variable, that is if a pointer to this memory is copied to a another pointer, at which point the original owner is longer the unique owner.
+    A key intuition is that a region of memory, subsequent to its initialisation, is only able to gain additional owners if access to the region of memory is \emph{shared} to another variable, that is if a pointer to this memory is copied to a another pointer, at which point the original owner is no longer the unique owner.
     If we copy the pointer to a global variable, the variable could suddenly be read by any other running threads in the program, so we have to take a conservative decision that the memory no longer has a single unique owner.
@@
     It has been used in languages such as Java~\cite{Choi1999} to determine whether an object which is created within a method escapes.
-    If it does not escape, meaning it is only used locally in the method, then it can be allocated on the stack rather than the heap as a performance optimisation.
+    If it does not escape, meaning it is only used locally in the method, then the object can be allocated on the stack rather than the heap as a performance optimisation.
     LLVM also uses escape analysis for a similar kind of optimisation~\footnote{\url{https://github.com/llvm/llvm-project/blob/main/llvm/lib/Analysis/CaptureTracking.cpp}}.
@@
         \item a pointer escapes, as then we do not know where or how the pointer may be used, so we must take a conservative approach and assume the memory region no longer has at most a single unique owner.
-        \item a pointer is does not escape but access to the memory is shared rather than transferred to a local variable. 
+        \item a pointer does not escape but access to the memory is shared rather than transferred between local variables. 
     \end{itemize}
@@
     The existing escape analysis in LLVM is implemented to operate on LLVM IR code, as the IR representation, as mentioned in \autoref{subsec:clang-and-llvm}, is well suited for efficient analysis algorithms.
-    However, much of the existing work from \autoref{sec:related-work} which use Clang perform both their analyses and their source code refactoring on the Clang AST.
-    We however propose an improved novel design which uses the Clang AST for parsing and refactoring and the LLVM IR for simple and efficient analysis.
+    However, much of the existing work in \autoref{sec:related-work} use Clang and the Clang AST to perform their parsing, analyses and their source code refactoring..
+    We however propose an improved novel design which uses the Clang AST for parsing and refactoring and the LLVM IR for efficient analysis.
 
     For the same input file, we can use the \emph{LibTooling} library to generate an AST and IR in-memory representation.
-    Then, for any variable declaration (\emph{VarDecl}) in the Clang AST which allocates new memory on the heap, we locate this declaration in the IR.
-    This is possible as we can instruct Clang to not discard variable names using a \texttt{fno-discard-value-names} flags, and as the combination of function signature and variable name in the translation unit (\emph{cpp file}) must be unique according to the One-definition rule~\cite{C++17}.
+    Then, for any variable declaration (\texttt{VarDecl}) in the Clang AST which allocates new memory on the heap, we locate this declaration in the IR.
+    This is possible as we can instruct Clang to not discard variable names using a \texttt{fno-discard-value-names} flag, and as the combination of function signature and variable name in the translation unit must be unique according to the One-definition rule~\cite{C++17}.
 
@@
 
-    In \autoref{fig:code-to-ast-example}, given the variable declaration for the variable \texttt{a}, then as it is not a global variable, we can walk its parent nodes until we reach a \emph{FunctionDecl}.
-    We are now able to use the function name to find the corresponding function declaration in the IR module, which can be seen in  \autoref{lst:code-to-ir}.
+    In \autoref{fig:code-to-ast-example}, given the variable declaration for the variable \texttt{a}, then as it is not a global variable, we can visit its parent nodes in the AST until we reach a \texttt{FunctionDecl}.
+    We can then use the name of the \texttt{FunctionDecl} to find the corresponding function declaration in the IR module, which can be seen in  \autoref{lst:code-to-ir}.
     Note that although the \texttt{main} function name corresponds exactly, other function names in the IR module will have mangled names~\footnote{\url{https://www.ibm.com/support/knowledgecenter/en/ssw_ibm_i_72/rzarg/name_mangling.html}}.
     We again use Clang library functions to mangle function names if necessary so they can be found in the IR. 
-    Given a function in the IR, we can use its symbol table to use the declaration name to find the IR value on which we will perform our analysis.
+    Given the IR function declaration, we can find the corresponding IR value of our initial variable \texttt{a} by finding the name of the variable inside its symbol table.
+    This IR value can then be used to perform our analysis.
 
@@
     The analysis gives a result of whether the IR value escapes, a boolean \texttt{true} or \texttt{false}.
-    For a \texttt{false} result, the value declaration is refactored into a unique pointer, otherwise a shared pointer.
-    However, the rewriting is identical for both cases with just the string tokens that we replace for the original type differing.
+    For a \texttt{false} result, meaning the value does not escape, the variable declaration is refactored into a unique pointer, and otherwise a shared pointer.
+    However, the rewriting of the source code is identical for both cases with just the string tokens that replace the original source code differing.
     The rewriting was developed using a rule based approach, using test driven development (TDD) as this allowed rapid but correct development by continuously being able to check that additional created rules still cause all the tests to pass, causing no regressions.
-    The five cases of code (\emph{or rewriting rules}) that needs to be rewritten are:
+    The five cases of code (\emph{or the rewriting rules}) that needs to be rewritten are:
     \begin{itemize}
-        \item Declaration types.
-        \item Variable initialisation and assignment expressions.
+        \item The type of a variable declaration
+        \item Variable initialisation or assignment expressions.
         \item Pointer delete statements.
         \item Function return types.
-        \item Pointer Usages.
+        \item Function call expressions that expect raw pointers.
     \end{itemize}
 
-    Rewriting declaration types are the easiest of all. We can directly replace the type of the variable declaration node that we already have access to. 
-    For example for \mintinline{c++}{int *ptr}, we can extract the type of the variable, and format it into the string \mintinline{c++}{"std::shared_ptr<{}>"} or \mintinline{c++}{"std::unique_ptr<{}>"}, where the \texttt{\{\}} represent the replacement field.
+    Rewriting the type of a variable declaration is the easiest of all. We can directly replace the type of the variable declaration node that we already have access to. 
+    For example for \mintinline{c++}{int *ptr}, we can extract the type of the variable, and format it into the string \mintinline{c++}{"std::shared_ptr<{}>"} or \mintinline{c++}{"std::unique_ptr<{}>"}, where \texttt{\{\}} represent the replacement field, which in this case will be formatted with the string \mintinline{c++}{int}.
 
-    Initialisation and assignment expressions are less simple due to two reason.
+    Initialisation and assignment expressions are less simple due to two reasons.
     First of all, the initialisation of a variable does not necessarily have to occur immediately at the point of declaration.
@@
     \end{itemize}
-    For pointer variables, this is convoluted even more depending on if the pointer gets initialised using an existing pointer \mintinline{c++}{int* a = nullptr; int *b = a;}, or if it gets initialised using a new statement \mintinline{c++}{auto p = new Circle{Point{0,0},10}}.
+    For pointer variables, this is convoluted even more depending on if the pointer gets initialised using an existing pointer \mintinline{c++}{int* a = nullptr; int *b = a;}, or if it gets initialised using a \mintinline{c++}{new} statement \mintinline{c++}{auto p = new Circle{Point{0,0},10}}.
     In the case of a new statement, we can initialise the rewritten smart pointer using a value initialiser \mintinline{c++}{std::make_unique<>()} or the shared pointer equivalent. 
-    Otherwise, we can create a smart pointer from a pointer value using by the default equal sign copy initialiser or parentheses expression list.
+    Otherwise, we can initialise the rewritten smart pointer from an existing pointer by using the default equal sign copy initialiser or parentheses expression list.
 
-    Rewriting pointer deletion statements also require us to traverse the AST to find \mintinline{c++}{delete} statements for the variable declaration node that we are rewriting.
-    However, they do not require any complicated logic, and we can upon finding relevant statements remove them from the source code.
+    Rewriting pointer delete statements also requires us to traverse the AST to find \mintinline{c++}{delete} statements for the variable declaration node that we are rewriting.
+    However, this does not require any complicated logic and we can upon finding relevant \mintinline{c++}{delete} statements simply remove them from the source code.
 
     If the variable declaration is a variable that gets returned at the end of the function, then the return type of the function signature will need to be rewritten too. 
-    We explicitly check for each variable declaration that we handle if it gets returned, and if so, we can rewrite the function return type the same way that we handle variable declaration types.
+    We explicitly check for each variable declaration that we handle if it gets returned, and if so, we can rewrite the function return type the same way that we handle the types of variable declarations.
 
-    Lastly, smart pointer have overloaded operators for the arrow \texttt{->} and dereference \texttt{*} operators so that they behave identically as if the value was a raw pointer.
-    This simplifies our task as it reduces the amount of rewriting we have to do for each usage of a specific pointer value.
-    However, there is no such overloaded operator for passing smart pointer as function arguments. 
-    In this case, the compiler will find a type error, when it might expect a type such as \mintinline{c++}{int *}, but actually receives \mintinline{c++}{std::unique_ptr<int>}. 
-    Therefore, we again have to traverse the AST to find all function call expressions which involve the variable declaration that we are currently handling, and we need to add \mintinline{c++}{get()} calls to the smart pointer values to get a raw pointer to the managed objects.
-
-    We use the Clang \emph{Rewriter}\footnote{\url{https://clang.llvm.org/doxygen/classclang_1_1Rewriter.html}} class as a high level abstraction compared to handling files and text buffers ourselves, which would likely be error-prone for buffer overflows.
+    Lastly, smart pointers have overloaded operators for the arrow \texttt{->} and dereference \texttt{*} operators so that they behave identically as if the value was a raw pointer.
+    This simplifies our task as it reduces the amount of rewriting we have to do for each use of the new smart pointer.
+    However, there is no such overloaded operator for passing smart pointers as function arguments. 
+    In this case, the compiler will find a type error as it might expect a type such as \mintinline{c++}{int *}, but actually receives \mintinline{c++}{std::unique_ptr<int>}. 
+    Therefore, we again have to traverse the AST to find all function call expressions which involve the current variable declaration and add \mintinline{c++}{get()} calls in order to get a raw pointer to the managed object.
+    We use the \emph{Clang Rewriter}\footnote{\url{https://clang.llvm.org/doxygen/classclang_1_1Rewriter.html}} class as a high level abstraction compared to handling files and text buffers ourselves.
 
@@
         \caption{An example of all the rewriting rules except pointer delete statements.
-This example solely shows rewriting and does not necessarily follow respect any smart pointer viability analyses.
+This example solely shows rewriting and does not necessarily respect any smart pointer viability analyses.
 We see how both the declaration and equal sign copy initialisation of \texttt{a} have been rewritten to into a \texttt{shared\_ptr}.
@@
     In \autoref{fig:rewriting}, we show an example that incorporates all the cases in which code needs to be rewritten except pointer delete statements.
-    It is visible how in the majority of cases, the first two rules will be applicable as every new smart pointer need to have the correct type and be initialised to the right type of object. 
-    The fifth rule will also likely be applicable in most cases as variables tend to not be unused and variables which were pointers will be used in functions expecting pointers, as they otherwise could be stack variables.
+    It is visible how in the majority of cases, the first two rules will always be applicable as every new smart pointer needs to have the correct type and be initialised to an object of the right type.
+    The fifth rule will also likely be applicable in most cases as variables tend to not be
+    unused and therefore will be used in functions that are expecting a raw pointer.
 
@@
 
-    Through our research, we would like to answer the question of whether our implemented tool can truly improve C++ code by seeing if the tool can be used outside small examples on larger existing programs, as well as if the refactored results produce code which can be compiled and can be judged to be better.
-    Secondly, if the tool is able to do this, we want to evaluate how well the tool performs by analysing the effectiveness of the analysis algorithms.
+    Through our research, we would like to answer the question of whether our implemented tool can truly improve C++ code by seeing if the tool can be used not only on small test examples but also on larger existing programs.
+We would also like to determine if the rejuvenation produces code which can be compiled and can be judged to be better.
+    Then, if the tool is able to do this, we want to evaluate how well the tool performs by analysing the effectiveness of the analysis algorithms.
 
@@
     \begin{itemize}
-        \item \textbf{Hypothesis 1} The automatic translation of raw pointers into smart pointers can be used to make existing C++ code more modern, safe and extendable.
+        \item \textbf{Hypothesis 1} The automatic translation of raw pointers into smart pointers can be used to make existing C++ code more modern and safe.
         \item \textbf{Hypothesis 2} The tool can reliably identify situations in which unique pointers should be used.
@@
 
-    We run our experiments on a Windows 10 Pro x64 machine using the Windows Subsystem for Linux, with an AMD Ryzen 5 PRO 4650U and 16 GiB RAM, using TinyXML-2 v8.0.0.
+    We run our experiments on a Windows 10 Pro x64 machine using the Windows Subsystem for Linux, with an AMD Ryzen 5 PRO 4650U and 16 GiB RAM.
     The ptr-tidy tool was compiled with GCC v10.2.0 using CMake release profile, using Clang and LLVM v11.1.0 libraries.
@@
 
-    In order to answer hypothesis 1, it was decided to run a case study instead of applying the tool in large on a range of open source projects. 
-    This was decided as an analysis into a well chosen target project will be able to show relevant success and failure cases that we can extrapolate onto more general C++ projects of the same kind. 
-    If the impact of the success cases outweighs the the faults of the failure cases, then we will be able to judge the our hypothesis as being true.
+    In order to answer hypothesis 1, it was decided to run a case study on a specifically chosen C++ project.
+    This was decided as an analysis into a well chosen target project will be able to show relevant success and failure cases of the tool.
+    If the impact of the success cases outweighs the faults of the failure cases, then we will be able to judge our hypothesis as being true.
 
-    We decided on using the \emph{TinyXML-2} library as our chosen case study~\footnote{\url{https://github.com/leethomason/tinyxml2}}.
+    We decided on using the \emph{TinyXML-2}~\footnote{\url{https://github.com/leethomason/tinyxml2}} library as our chosen case study.
     TinyXML-2 is a C++ XML parsing library that can parse an XML document and build from that a Document Object Model that can be read, modified, and saved.
     TinyXML-2 has over three thousand \emph{GitHub stars} which puts it in the top 1\% of projects on the hosting website, which can be used as a rough metric for its popularity and good code quality~\cite{Sanatinia2016}.
-    It is also for example used in the \emph{libigl}~\cite{Libigl} geometry processing library,  which itself is used by a variety of large industry companies.
+    It is also used in the \emph{libigl}~\cite{Libigl} geometry processing library,  which itself is used by a variety of large industry companies.
     It is released under a \emph{zlib} license~\footnote{\url{https://opensource.org/licenses/Zlib}} which allows us to use and alter the software.
-    What makes library a particularly attractive choice is the simplicity of the library.
+    What makes the library a particularly attractive choice is the simplicity of the library.
     The library does not make use of exceptions, RTTI or the STL~\cite{C++17}, and therefore most importantly does not make use of smart pointers. 
-    The library is also around three thousand lines of code in length, making it large enough to likely have a variety of edge cases, whilst remaining small enough for manual inspection.
+    The library is also around three thousand lines of code in length, making it large enough to likely have a variety of interesting features to analyse, whilst remaining small enough for manual inspection.
 
@@
     In particular, what makes this suitable for the target of our case study over perhaps other similarly small and simple libraries is that due to parsing varying documents at runtime, the library is particularly heavy in pointer logic and memory allocations.
-    We can in \autoref{lst:tinyxml-refactor-example} an example of a method from the library which would be well suited for refactoring to use smart pointers.
-    The heavy use of pointers and memory allocation is what leads us to believe that the case study into this library will generalise well onto all other C++ projects.
+    We can see in \autoref{lst:tinyxml-refactor-example} an example of a method from the library which would be well suited for refactoring to use smart pointers.
+    The heavy use of pointers and memory allocation is what leads us to believe that analysis into this library will generalise well onto other C++ projects.
 
@@
     More complex libraries or applications would have their compilation options and any include and link libraries included in compilation command.
-    The user of our tool can easily generate a compilation database for the project that they wish to rejuvenate using their build tool, such as \emph{CMake}.
+    The user of our tool can easily generate a compilation database for the project that they wish to rejuvenate using their build tool, such as \emph{CMake}~\footnote{\url{https://cmake.org/}}.
 
@@
 
-    We can first and foremost see in \autoref{lst:tinyxml-refactor-example-refactored} the result of the rejuvenation our candidate method in \autoref{lst:tinyxml-refactor-example}.
+    We can first and foremost see in \autoref{lst:tinyxml-refactor-example-refactored} the result of the rejuvenation on our candidate method in \autoref{lst:tinyxml-refactor-example}.
 
     We can see that we achieve a similar refactoring as in \autoref{fig:rewriting}. 
-    The smart pointer is initialised using using the pointer copy initialiser.
+    The smart pointer is initialised using the pointer copy initialiser.
     We see identical refactoring in the other \texttt{XMLDocument:New...} methods.
@@
     Parameters are by default always not marked \texttt{noescape} and \texttt{noalias}~\footnote{\url{https://clang.llvm.org/docs/AttributeReference.html\#noalias}} in the generated LLVM IR, as can be seen in \autoref{fig:function-arguments}.
-    If we enable Clang compiler optimisations when generating the IR, the compiler correctly identify that \texttt{p} cannot escape within the function, giving it a signature in the IR of 
+    If we enable Clang compiler optimisations when generating the IR in \autoref{fig:function-arguments}, the compiler correctly identify that \texttt{p} cannot escape within the function, giving it a signature in the IR of
     \mintinline{llvm}{
@@
     Also, we still do not have non-aliasing guarantees so that we know that the object has not escaped before entering the function.
-    What is possible is that we can manually mark a variable in the code as being non-capture and non-alias, such as:
+    What is possible is that we can manually mark a variable in the code as being non-captured and non-aliased, such as:
     \begin{minted}{c++}
@@
     This relies however on the Clang compiler specific implementation of the \texttt{noescape} attribute which does not exist in the GCC compiler, as well as the non-standard, compiler specific implementation of the \texttt{restrict} attribute.
-    Such code therefore is not portable and not standard conforming, making it very unlikely unacceptable for most projects.
+    Such code therefore is not portable and not standard conforming, making it likely unacceptable for most projects.
     This last unlikely case is not currently correctly handled by the LLVM escape analysis to produce a more correct refactoring. 
     The automatic generation of such front-end code however could be possible through the integration of Clang compiler optimisations and related research such as by \citet{Campos2015} for integration into our tool for better performance.
-    The impact this on our tool is that the extent of our analysis is not as complete as we would like it to be, however \autoref{lst:tinyxml-refactor-example-refactored} and \autoref{fig:rewriting-example2} show that we can still achieve desirable results through the analysis of local variables.
+    The impact this has on our tool is that the extent of our analysis is not as complete as we would like it to be, however \autoref{lst:tinyxml-refactor-example-refactored} and \autoref{fig:rewriting-example2} show that we can still achieve desirable results through the analysis of local variables.
 
@@
 
-    The rejuvated version of the TinyXML-2 library that the tool produces is not directly compilable into an object file, mainly due to edge cases in the rewriting component. 
-    These mainly stem from outdated C++ practises that must be explicitly handled in modern code that has more strict type checking. 
+    The rejuvenated version of the TinyXML-2 library that the tool produces is not directly compilable into an object file, mainly due to edge cases in the rewriting component, causing syntactically incorrect output code.
+    These mainly stem from the use of now outdated C++ practises.
     For example, the initialisation of a pointer using the value \texttt{0}, such as
@@
     \mint{c++}|std::shared_ptr<XMLNode>  returnNode = std::shared_ptr<XMLNode>(0);|
-    therefore causes a type error during compilation.
+    therefore causes a type error during compilation as the \mintinline{c++}{shared_ptr} constructor has stricter type checking.
 
@@
     The size and complexity of the language makes it difficult to ensure that all edge cases are handled by the rewriting component, which can in future development be continuously improved.
-    The fixes that need to be applied can be done by the programmer, with all of the identified fixes in the TinyXML-2 library being a single line in length.
+    The fixes that need to be applied in order to compile the rejuvenated program can be done manually by the programmer, with all of the identified fixes in the TinyXML-2 library being a single line in length.
 
-    We do not think this however detracts from the value or usefulness of our tool.
-    Our tool provides fully rejuvenated and compilable programs without any human input needed for simpler inputs, using modern styles of C++.
-    For larger, more complex, older input programs some human input is currently still needed, however the work of our tool still allows the programmer to make their programs more modern and safer in by performing the complex analysis and guiding the programmer towards the needed fixes through compiler errors.
-    This results in the rejuvenated output being obtained easier by removing the need for the programmer to manually understand the structure and style of the programmer and the often complex ownership relations between the various variables in the program.
+    However, We do not think this detracts from the value or usefulness of our tool.
+    Our tool provides fully rejuvenated and compilable programs without the need for any human input simpler input programs. 
+    These programs are modernised to use features such as smart pointers, which also bring stricter type checking.
+    For larger, more complex, older input programs some human input is currently still needed, however the work of our tool still allows the programmer to make their programs more modern and safer by performing the complex analysis and guiding the programmer towards the needed fixes through compiler errors.
+    This results in the rejuvenated output being obtained easier by removing the need for the programmer to manually understand the structure and style of the program and the often complex ownership relations between the various variables in the program.
 
-    We therefore believe that the analysis of our tool on the case study of the TinyXML-2 library, even with the identified failure cases, the success of the automated translation and identified success cases allow us to confirm our first hypothesis.
+    We therefore believe that from the analysis of our tool on the case study of the TinyXML-2 library, even with the identified failure cases, the success of the automated translation and identified success cases allow us to confirm our first hypothesis.
 
@@
 
-    In order to answer hypothesis 2, it was decided to run analysis of the tool on more various open source projects and gather metrics on the performance of the tool in terms of its performance in successfully identifying pointers, how extensive the rejuvenation in context of the whole program, and some more general runtime characteristics.
+        \begin{table*}
+    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
+    \hline
+                                   & \textbf{clipper} & \textbf{happyhttp} & \textbf{jzon} & \textbf{micropather} & \textbf{poisson} & \textbf{polypartition} & \textbf{tinyxml2} & \textbf{xatlas} \\ \hline \hline
+    \textbf{Runtime (ms)}          & 146.521          & 10.540             & 3.140                & 2.429         & 4.471            & 11.885                 & 16.055            & 105.212         \\ \hline
+    \textbf{Escaped Pointer}       & 94               & 22                 & 2                    & 39            & 7                & 46                     & 164               & 235             \\ \hline
+    \textbf{Non-Escaped Pointers}  & 132              & 16                 & 3                    & 34            & 3                & 22                     & 84                & 111             \\ \hline
+    \textbf{Lines of Code Changed}   & 205             & 27                 & 3                    & 57            & 6                & 31                     & 142               & 154             \\ \hline
+    \textbf{Total Lines of Code} & 5035             & 1273               & 1324                 & 1587          & 523              & 2270                   & 5366              & 10318           \\ \hline
+    \end{tabular}    
+    \caption{Data gathered from the rejuvenation of each projects.
+Total lines of code include both the implementation and header file of each project.
+Runtime measured in milliseconds as the median of five runs.}
+    \label{table:data}
+    \end{table*}
 
-    It was decided that whilst gathering more projects to test, we would continue to focus on simple, single file libraries such as TinyXML-2 in order to ease the cost of setting up each project to be used by the tool.
+    \begin{figure*}
+        \centering
+        \includegraphics[scale=0.55]{images/pointers.png}
+        \caption{First graph shows the fraction of non-escaped, and therefore refactored, pointers for each project.
+This varies from a minimum of 30\% up to a maximum of 60\%, with a mean of 42\%.
+The second graph shows the number of total pointers compared to the number of total lines in each project.}
+        \label{fig:ratio-pointer}
+    \end{figure*}
+
+    \begin{figure*}
+        \centering
+        \includegraphics[scale=0.8]{images/lines.png}
+        \caption{Graph showing the fraction of modified lines by the tool compared to the total number of lines in each project, which is the combination of implementation and header files.
+This varies from a minimum of 0.22\% up to a maximum of 4.07\%, with a mean of 2.18\%.}
+        \label{fig:lines}
+    \end{figure*}
+
+    \begin{figure*}
+        \centering
+        \includegraphics[scale=0.55]{images/time.png}
+        \caption{Graphs of the tool runtime versus the total lines of code in each project and the total number of pointers that was analysed.
+The x axis for both graphs are plotted on a log scale.
+This is for clarity purposes in order to emphasise the two outliers consisting of the Clipper and Xatlas libraries.
+Time was measured in microsecond using the Bash \texttt{time} command and measured as the median real execution time of 5 runs.}
+        \label{fig:time}
+    \end{figure*}
+
+    In order to answer hypothesis 2, it was decided to run analysis of the tool on more open source projects and gather metrics on the performance of the tool in terms of its performance in successfully identifying pointers can be refactored, how extensive the rejuvenation is in context of the whole program, and some general runtime characteristics.
+
+    It was decided that whilst gathering more projects to test, we would continue to focus on simple, single file libraries similar to TinyXML-2 in order to ease the cost of setting up each project to be used by the tool.
     We used a list of single file, public-domain/open source libraries with minimal dependencies~\cite{Barrett2019} by acknowledged game and software developer Sean Barret~\footnote{\url{https://www.mobygames.com/developer/sheet/view/developerId,4966/}} to discover as many as possible candidate libraries.
-    This list also ensures that we only select projects which we are able to use and modify as well as including project from many areas of software engineering such as graphics and geometry to file parsers.
+    This list also ensures that we select projects from many areas of software engineering such as graphics and geometry to file parsers.
     This ensures that we run the tool across a wide variety of projects which will have different distributions and usages of memory allocations and pointer use, giving greater validity that our tool is able to generalise to any type of project.
@@
 
-    \begin{table*}
-    \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
-    \hline
-                                   & \textbf{clipper} & \textbf{happyhttp} & \textbf{jzon} & \textbf{micropather} & \textbf{poisson} & \textbf{polypartition} & \textbf{tinyxml2} & \textbf{xatlas} \\ \hline \hline
-    \textbf{Runtime (ms)}          & 146.521          & 10.540             & 3.140                & 2.429         & 4.471            & 11.885                 & 16.055            & 105.212         \\ \hline
-    \textbf{Escaped Pointer}       & 94               & 22                 & 2                    & 39            & 7                & 46                     & 164               & 235             \\ \hline
-    \textbf{Non-Escaped Pointers}  & 132              & 16                 & 3                    & 34            & 3                & 22                     & 84                & 111             \\ \hline
-    \textbf{Lines of Code Changed}   & 205             & 27                 & 3                    & 57            & 6                & 31                     & 142               & 154             \\ \hline
-    \textbf{Total Lines of Code} & 5035             & 1273               & 1324                 & 1587          & 523              & 2270                   & 5366              & 10318           \\ \hline
-    \end{tabular}    
-    \caption{Data gathered from the rejuvenation of each of these projects.
-Total lines of code include both the implementation and header file of each project.
-Runtime measured in milliseconds as the median of five runs.}
-    \label{table:data}
-    \end{table*}
-
-    \begin{figure*}
-        \centering
-        \includegraphics[scale=0.55]{images/pointers.png}
-        \caption{First graph shows the fraction of non-escaped, and therefore refactored, pointers for various projects.
-This varies from a minimum of 30\% up to a maximum of 60\%, with a mean of 42\%.
-The second graph shows the number of total pointers compared to the number of total lines in each project.}
-        \label{fig:ratio-pointer}
-    \end{figure*}
-
-    \begin{figure*}
-        \centering
-        \includegraphics[scale=0.8]{images/lines.png}
-        \caption{Graph showing the fraction of modified lines by the tool compared to the total number of lines in each project, which is the combination of implementation and header files.
-This varies from a minimum of 0.22\% up to a maximum of 4.07\%, with a mean of 2.18\%.}
-        \label{fig:lines}
-    \end{figure*}
-
-    \begin{figure*}
-        \centering
-        \includegraphics[scale=0.55]{images/time.png}
-        \caption{Graphs of the tool runtime versus the total lines of code in each project and the total number of pointers that was analysed.
-The x axis for both graphs are plotted on a log scale.
-This is for clarity purposes in order to emphasise the two outliers consisting of the Clipper and Xatlas libraries.
-Time was measured in microsecond using the Bash \texttt{time} command and measuring the median real execution time of 5 runs.}
-        \label{fig:time}
-    \end{figure*}
-
     \subsubsection{Results and Discussion}
@@
     We see first of all in \autoref{fig:ratio-pointer} that our tool has consistent behaviour across the various projects, with 4 projects having between 30\% and 40\% of pointers identified for refactoring and 4 other projects having between 40\% and 60\% of pointers identified.
-    The mean of 42\% is shows that we are able to identify a non-trivial amount of pointers, however it is unclear this is not higher because of a fault in the analysis or whether this is the typical performance for a conservative analysis.
+    The mean of 42\% shows that we are able to identify a non-trivial amount of pointers, however it is unclear this is not higher because of a fault in the analysis or whether this is the typical performance for a conservative analysis.
     The second graph in \autoref{fig:ratio-pointer} also shows how the amount of pointers in a project as the size of the project increases is close to linear. 
@@
     There seems to be no relationship between the 4 projects that had over 40\% of pointers identified and their lengths.
-    We therefore conclude that the length of a project that the tool has to rejuvenate has no impact on its performance.
+    We therefore conclude that the size of a project that the tool has to rejuvenate has no impact on its performance.
 
-    \autoref{fig:lines} the number of lines that were modified in the rejuvenated version of the project compared to the original. 
-    We can much more clearly see our previous expectation that some projects utilise pointer more and other less. 
+    \autoref{fig:lines} shows the number of lines that were modified in the rejuvenated version of the project compared to the original. 
+    We can much more clearly see our previous expectation that some projects utilise pointers more and other less. 
     The two projects with the lowest amount of modified lines, Jzon and the Poisson Disk Points Generator both only had ten or less pointers to analyse in total, with the Jzon library in particular only having three changed lines from the original.
@@
     We see a linear relationship in both of the graphs, except for the outliers of the Clipper and Xatlas libraries.
-    We were unable to establish why these libraries take an exceptionally longer time to process than the TinyXML-2 library which has the same amount of lines of code and analysed pointers as the Clipper library.
+    We were unable to establish why these libraries take an exceptionally longer time to process, especially since the TinyXML-2 library has a similar amount of lines of code and analysed pointers as the Clipper library.
+    This gives us confidence that the use of our tool would scale well onto even larger projects and therefore as useful as possible.
 
     To conclude, we believe we can partially confirm our second hypothesis.
-    We show that the tool generalises to a range of other libraries apart from the TinyXML-2 library selected in our case study and show that our analysis achieves similar results across these libraries.
+    We show that the tool generalises to a range of other libraries apart from the TinyXML-2 library selected in our case study and show that our analysis achieves similar results across all these libraries.
     However, what we cannot conclude is the accuracy of our tool in terms of the best case performance. 
@@
     We show how modern C++, that is the C++ 11 standard and newer, encourage the use of \emph{smart pointers} instead for memory management. 
-    Smart pointer create memory or resource handles which automatically get released when the handle exists the scope of a function or method. 
+    Smart pointers create resource handles which automatically get released when the handle exits the scope of a function or method. 
     This referred to as \emph{RAII} (\textit{Resource Acquisition is Initialisation}).
     We also highlight the eventual increasing need for automatic rejuvenation, the replacement of outdated coding patterns with newer, higher-level abstractions, of maintained projects as languages move and evolve much quicker than can be supported in enterprise environments.
-    The learning and application of new language requires a significant time investment that may not be able to be allocated.
 
-    We then introduce our \emph{ptr-tidy} tool which was designed automatically rejuvenate code that does not use smart pointers, thereby solving our first problem.
+    We then introduce our \emph{ptr-tidy} tool in order to automatically rejuvenate code that does not use smart pointers, thereby solving our first problem.
     We show the design and implementation of the tool, which uses the Clang and LLVM libraries. 
-    The tool uses Clang and the Clang abstract syntax tree (AST) in order to obtain a rich and closely resembling representation of an input program.
+    The tool uses Clang and the Clang abstract syntax tree (AST) in order to obtain a rich and closely resembling representation of the input program.
     The tool then uses the equivalent representation of the input program in the lower-level LLVM intermediate representation (IR) in order to analyse which pointers are suitable to be \emph{unique pointers} or \emph{shared pointers}.
     The AST and output of the analysis are then input to a rewriting component which creates and outputs our final rejuvenated program. 
-    This design is the solution to our second problem, as we believe this design is both generic, reusable and powerful enough to be used for a variety of same language refactoring or rejuvenation or different language transpiling.
+    This design is the solution to our second problem, as we believe this design is both generic, reusable and powerful enough to be used for a variety of rejuvenations.
     The design of the LLVM IR allows for powerful and efficient analyses to be performed, whilst its language agnostic nature allows a variety of possible parsing front-ends and output back-ends to be used, such as existing mature implementations available for C++ or Rust.
@@
     We observe that the tool is particularly effective at rejuvenating specific cases in programs where dynamic resources are created at the top of the function call stack and passed down the stack through return values.
-    We also discovered that due to limitations in the specification of the C++ language, we are unable to rejuvenate function parameters statically, which meant the extend of our rejuvenation had limits.
+    We also discovered that due to limitations in the specification of the C++ language, we are unable to rejuvenate function parameters, which meant the extent of our rejuvenation had limits.
     The successful application of the tool across a varied set of open-source projects allowed us to conclude that the analysis and design of the tool were indeed suitable for refactoring generic C++ programs.
@@
     The escape analysis algorithm described by \citet{Choi1999} is used as the default escape analysis algorithm in the Oracle Java Virtual Machine implementation~\footnote{\url{https://docs.oracle.com/en/java/javase/16/vm/java-hotspot-virtual-machine-performance-enhancements.html}} and is widely cited~\cite{Wang2020, Stadler2014}.
-    Most importantly, the authors present an interprocedural analysis which we hypothesise could bring greater effectiveness to our tool and therefore we believe that it could lead to improved analysis performance compared to our implemented intraprocedural analysis.
-    The similarity of Java and C++ as object oriented languages allowed for a direct application of the algorithm to C++.
+    Most importantly, the authors present an interprocedural analysis which we hypothesise could bring greater effectiveness to our tool as we believe that it could lead to improved analysis performance compared to our implemented intraprocedural analysis.
+    The similarity of Java and C++ as object oriented languages should allow for a direct application of the algorithm in C++.
 
     Secondly, we mention in \autoref{subsec:case-study} that we are unable to enable optimising compiler invocations for generating the LLVM IR as it breaks our AST to IR correspondence. 
-    However, it remains to investigate whether these optimisations can be called dynamically during the running of our tool in order to gain information about conservative assumptions that we currently make.
+    However, it remains to be investigated whether these optimisations can be called dynamically during the running of our tool in order to gain insight into conservative assumptions that we currently have to make.
 

### 2021-04-13 paper.tex
Incorporate comments from ou

@@
     Engineers at both Microsoft~\cite{Thomas2019} and Mozilla~\cite{Hostfelt2019} converge on Rust~\cite{Balasubramanian2017} as a possible solution.
-    Rust is a systems language that offers similar performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model guarantee memory safety.
+    Rust is a systems language that offers similar performance as C and C++~\cite{Lin2016}, however its linear type system and memory ownership model guarantee that none of the memory errors that we are considering can occur.
     The rewriting of a program in a new language, namely from C++ to Rust, is a colossal undertaking, especially given that Rust has a reputation for being difficult to learn~\cite{Astrauskas2019}.
@@
     The former also uses the Clang driven approach like \citet{Wright2013}, while the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
-    We have not evaluated IPR as a representation compared to the use of Clang, but we can  comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation for Clang, as well as being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
+    We have not evaluated IPR as a representation compared to the use of Clang, but we can  comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation for Clang, the ongoing support for new C++ standards and non-standard features as well as being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
 
@@
     The first type of smart pointer are \emph{unique pointers}, which can have only a single unique owner at any time and therefore are ideal for resources which may not be copied.
-    The resource managed by a unique pointer is released when the current owner exits the scope of the function or method it is contained in.
+    The resource managed by a unique pointer is released when execution of the program exits the scope of the function or method that the owner of the resource is contained in.
     The ownership of a unique pointer can be transferred between variables using \texttt{std::move} which \emph{moves} the ownership from one variable to another, making the original owner invalid.
@@
     LLVM~\cite{Lattner2004} is a compiler infrastructure project consisting of several compiler and toolchain technologies.
+
     LLVM is centrally designed around a language independent intermediate representation (IR), which can be created by a number of language front-ends and is used by a number of back-ends to generate machine code.
+    The LLVM IR is a language-independent, static single assignment (SSA)~\cite{Rosen1988} representation of a program.
+    Variables in SSA form can only be assigned once, which allows us to efficiently generate definition-usage graphs, and allows for fast and optimised analysis algorithms to be written.
     Programs can then be optimised by applying portable and reusable transformations on the IR.
+
     Any improvement in back-end machine code generation, as well as any new or improved IR optimisations therefore benefit all of the language front-ends.
-    It is therefore visible that this allows for a very modular and powerful design which can be seen in \autoref{fig:llvm-design}.
+    It can be seen that this allows for a very modular and powerful design, shown in \autoref{fig:llvm-design}.
 
@@
     The LLVM authors remark that this makes it very suitable for refactoring tools.
-
-    The LLVM IR is a language-independent, static single assignment (SSA)~\cite{Rosen1988} representation of a program.
-    Variables in SSA form can only be assigned once, which allows us to efficiently generate compact definition-usage graphs, which enable developers to simplify many dataflow optimizations and enables fast, flow-insensitive algorithms~\cite{Lattner2004}.
-
-    \vskip50pt
 
@@
         \begin{minted}{c++}
-        std::shared_ptr<int> test() {
-            std::shared_ptr<char> a = std::shared_ptr<char>(new char[2]);
+        std::shared_ptr<char[]> test() {
+            std::shared_ptr<char[]> a = std::shared_ptr<char[]>(new char[2]);
             a.get()[0] = 'a';

### 2021-04-13 paper.tex
Incorporate feedback Jeremy

@@
     Modern C++ recommends that smart pointers are used to automatically manage resources such as dynamic memory.
-    However, many legacy programs may have been written before these features were introduced to the language.
+    However, many legacy programs were been written before these features were introduced to the language.
     We develop the ptr-tidy tool in order to automatically rejuvenate such programs to use smart pointers.
-    The tool uses the Clang and LLVM libraries in order to create a componentised and powerful design which could be used by other refactoring or rejuvenation tools. 
-    We present a case study and also apply the tool on a number of open-source C++ libraries from a variety of areas in computer science to show that the tool can successfully rejuvenate non-trivial libraries.
+    The tool uses the Clang and LLVM libraries in order to create a componentised and powerful framework which could be used by other refactoring or rejuvenation tools. 
+    We present a case study and also apply the tool on a number of open-source C++ libraries from a variety of application domains to show that the tool can successfully rejuvenate non-trivial libraries.
     We discuss the circumstances in which the tool performs well and the current limitations.
@@
     Systems programming involves memory management, that is dynamically requesting memory from the operating system to be used and managed by the program.
-    This is done as the amount of memory that we need at runtime cannot be know at the time of compilation of the program.
+    This is done as the amount of memory that we need at runtime cannot be known at the time of compilation of the program.
     However, doing so creates the opportunity for memory safety errors~\cite{Dhurjati2003}.
@@
     \begin{itemize}
-        \item \textbf{Dangling pointers}, prematurely freeing memory that is still in use.
+        \item \textbf{Dangling pointer}, prematurely freeing memory that is still in use.
         \item \textbf{Double free}, freeing memory multiple times.
@@
 
-    Historically, systems programs are mostly written in C, C++ or a combination of the two and these remain the dominant languages.
+    Historically, systems programs were mostly written in C, C++ or a combination of the two and these remain the dominant languages.
     They are lower level languages that involve manual memory management by the programmer.
-    It is challenging to ensure that C and C++ programs contain no memory safety errors as evidenced by empirical data in industry, where in a presentation at the Linux Security Summit it was shown that in several popular projects such as Firefox, macOS, Ubuntu and Android all had over half of their CVEs~\footnote{\url{https://cve.mitre.org/}} attributed to issues with memory safety~\cite{Gaynor2020}.
+    It is extremely challenging to ensure that C and C++ programs contain no memory safety errors as evidenced by empirical data in industry, where in a presentation at the Linux Security Summit it was shown that in several popular projects such as Firefox, macOS, Ubuntu and Android all had over half of their CVEs~\footnote{\url{https://cve.mitre.org/}} attributed to issues with memory safety~\cite{Gaynor2020}.
     Similarly, in a presentation by Matt Miller, a security engineer at Microsoft, it is shown that around 70\% of their vulnerabilities that are addressed through security updates are due to memory safety issues~\cite{Miller2019}.
@@
 Lines 4 and 5 show how interruptions in the program control flow can mistakenly cause memory leaks as the \texttt{delete} statement is never reached.
-Smart pointers correct this by automatically inserting delete statements at all necessary locations at compile time.
+Smart pointers correct this as the compiler automatically inserts delete statements at all necessary locations at compile time.
 All necessary headers are omitted for brevity}
@@
     We provide a motivating example in \autoref{lst:motivating-example} where the initial version of the function has several control flow statements that could cause the \texttt{Circle} object to never be freed.
-    The example can seem trivial as all the possible executions of the program are easy to visualise, but can suddenly be rather non-trivial if the conditions in the \texttt{if} statement become calls into external libraries, which may throw an exception and cause the \texttt{delete} statement to again not be reached.
+    The example can seem trivial as all the possible executions paths of the program are easy to visualise, but can suddenly become rather non-trivial if the conditions in the \texttt{if} statement become calls into external libraries, which may throw an exception and cause the \texttt{delete} statement to again not be reached.
 
@@
     We exactly aim to replace outdated coding patterns with newer, higher-level abstractions, as well as fitting the notion that this tool would be applied once to a codebase, rather than be a reoccurring task.
-    We will however use the terms rejuvenating and refactoring interchangeably in this paper as it is a term that most developers are more familiar with.
+    We will however use the terms rejuvenating and refactoring interchangeably in this paper as refactoring is a term that most developers are more familiar with.
 
@@
     Similarly, several new languages such as Typescript~\footnote{\url{https://www.typescriptlang.org/}} and Dart~\footnote{\url{https://dart.dev/}} have been recently developed which were designed not to be used by themselves, but to be transpiled to the popular Javascript language.
-    These languages aim to solve shortcomings of the output language, whilst reusing the existing mature and efficient compilers or interpreters.
+    These languages aim to solve shortcomings of the target language, whilst reusing the existing mature and efficient compilers or interpreters.
 
@@
 
-    Generally, pointers are used to give a program access to a resource that cannot be directly included in the program itself, such as allocated memory or a file.
+    Generally, pointers are used to give a program indirect access to a resource that cannot be directly included in the program itself, such as allocated memory or a file.
     Pointers can also be used to pass large resources to functions without needing to create a copy, which would have a performance impact.
@@
 
-    The first type of smart pointer are \emph{unique pointers}, which can have only a single unique owner at any time and therefore are ideal for resources which may not be copied.
+    The first type of smart pointer is a \emph{unique pointer}, which has only a single unique owner at any time and therefore are ideal for resources which may not be copied.
     The resource managed by a unique pointer is released when execution of the program exits the scope of the function or method that the owner of the resource is contained in.
@@
 
-    The second type of smart pointer are \emph{shared pointers} which are used for resources which may not necessarily have a single unique owner and so several owners \emph{share} ownership to the resource.
+    The second type of smart pointer is a \emph{shared pointer} which is used for resources that may not necessarily have a single unique owner and so several owners \emph{share} ownership to the resource.
     Reference counting~\cite{Collins1960} is used to ensure that after the last owner loses access to the resource, it will finally be freed.
@@
     We will therefore need to develop a static analysis algorithm to identify the maximum number of owners a region of memory has throughout all points of the program. 
-    If a memory region has at most a single unique owner, then we can refactor the initialising raw pointer to be a unique pointer, and otherwise to a shared pointer.    
+    If a memory region has at most a single owner, then we can refactor the initialising raw pointer to be a unique pointer, and otherwise to a shared pointer.    
     This analysis will need to ensure that its result is correct, and will inevitably have to make some conservative assumptions about the program.
@@
 
-    It is now visible that the invariant of the unique pointer are only violated when:
+    It is now apparent that the invariant of the unique pointer is only violated when:
     \begin{itemize}
@@
 
-    In order to evaluate the objectives above, hypotheses were made:
+    In order to evaluate the objectives above, hypotheses were formed:
 
@@
     What makes the library a particularly attractive choice is the simplicity of the library.
-    The library does not make use of exceptions, RTTI or the STL~\cite{C++17}, and therefore most importantly does not make use of smart pointers. 
+    The library does not make use of the STL~\cite{C++17}, and therefore most importantly does not make use of smart pointers. 
     The library is also around three thousand lines of code in length, making it large enough to likely have a variety of interesting features to analyse, whilst remaining small enough for manual inspection.
@@
         \end{minted}
-        \caption{C++ and equivelent LLVM IR snippet showing how function arguments are not considered \texttt{noescape} or \texttt{noalias}}
+        \caption{C++ and equivalent LLVM IR snippet showing how function arguments are not considered \texttt{noescape} or \texttt{noalias}}
         \label{fig:function-arguments}
@@
 
-    \subsection{Analysis of performance on a corpora of projects}
+    \subsection{Analysis of performance on a corpus of projects}
 
@@
     In terms of the evaluation of the tool, we would like to extend it with more complex, multi-file projects and determine whether the translation unit barrier, where an analysis pass can only access the current implementation file, has an impact on analysis performance.
-    An inter-modular analysis, similar to existing LTO optimisations in compilers~\footnote{\url{https://llvm.org/docs/LinkTimeOptimization.html}}, can then be evaluated whether it brings greater analysis performance.
+    An inter-modular analysis, similar to existing link time optimisations in compilers~\footnote{\url{https://llvm.org/docs/LinkTimeOptimization.html}}, can then be evaluated whether it brings greater analysis performance.
 

### 2021-04-14 paper.tex
add review ou 2

@@
 
-    \begin{figure}
+    \begin{figure}[t]
         \centering
@@
     Reference counting~\cite{Collins1960} is used to ensure that after the last owner loses access to the resource, it will finally be freed.
-    Shared pointers are suitable to be used in most situations and such reference counted pointers are used by default for all values in the Python and Swift languages~\footnote{\url{https://docs.swift.org/swift-book/LanguageGuide/AutomaticReferenceCounting.html}}
+    Shared pointers are suitable to be used in most situations and such reference counted pointers are used by default for all values in the Python~\footnote{\url{https://docs.python.org/3/c-api/refcounting.html}} and Swift languages~\footnote{\url{https://docs.swift.org/swift-book/LanguageGuide/AutomaticReferenceCounting.html}}.
 
@@
 
-    Not a single function argument was refactored to use smart pointers, even though arguments are included in the analysis of the program, since the \texttt{VarDecl Matcher} also matches on arguments (\texttt{ParmVarDecl}) as it is a subclass~\footnote{\url{https://clang.llvm.org/doxygen/classclang_1_1ParmVarDecl.html}}.
-    Upon investigation, this remains largely an issue due to poor memory aliasing guarantees in the language~\cite{C++17}. 
-    Parameters are by default always not marked \texttt{noescape} and \texttt{noalias}~\footnote{\url{https://clang.llvm.org/docs/AttributeReference.html\#noalias}} in the generated LLVM IR, as can be seen in \autoref{fig:function-arguments}.
-    If we enable Clang compiler optimisations when generating the IR in \autoref{fig:function-arguments}, the compiler correctly identify that \texttt{p} cannot escape within the function, giving it a signature in the IR of
+    Not a single function parameter was refactored to use smart pointers, even though parameters are included in the analysis of the program, since the \texttt{VarDecl Matcher} also matches on arguments (\texttt{ParmVarDecl}) as it is a subclass~\footnote{\url{https://clang.llvm.org/doxygen/classclang_1_1ParmVarDecl.html}}.
+    Upon investigation, this is largely an issue due memory coherence guarantees in the specification of the language~\cite[p.~66]{C++17}. 
+    Function pointer parameters are assumed by default to escape and to possibly be aliased (\textit{the memory could possibly be accessed through a different variable}).
+    Therefore by default no parameters will ever be identified by our analysis as non-escaping and will not be refactored.
+    We can see this in \autoref{fig:function-arguments}, where the variable \texttt{p} in the generated LLVM IR is not marked as \texttt{noescape} and \texttt{noalias}~\footnote{\url{https://clang.llvm.org/docs/AttributeReference.html\#noalias}}.
+
+    If we enable Clang compiler optimisations when generating the IR for the C++ code in  \autoref{fig:function-arguments}, then the compiler can now correctly identify that \texttt{p} cannot escape within the function, giving it a signature in the IR of
     \mintinline{llvm}{
@@
     However this would break the AST to IR correspondence that we established in \autoref{subsec:methodology-analysistarget} as variables that exist in the AST may find themselves optimised and removed from the IR or functions can that exist in the AST may find themselves inlined in the IR.
-    Also, we still do not have non-aliasing guarantees so that we know that the object has not escaped before entering the function.
-    What is possible is that we can manually mark a variable in the code as being non-captured and non-aliased, such as:
+    Also, we still do not have non-aliasing guarantees for the variable \texttt{p} so that we know that the object has not escaped before entering the function.
+    It is possible to manually mark a variable in the code as being non-captured and non-aliased by adding parameter attributes such as:
     \begin{minted}{c++}
@@
     Such code therefore is not portable and not standard conforming, making it likely unacceptable for most projects.
-    This last unlikely case is not currently correctly handled by the LLVM escape analysis to produce a more correct refactoring. 
-    The automatic generation of such front-end code however could be possible through the integration of Clang compiler optimisations and related research such as by \citet{Campos2015} for integration into our tool for better performance.
+    The LLVM escape analysis currently incorrectly identifies parameters with such attributes to still be escaping, as it is a very rarely occurring optimisation case.
+
+    % The automatic generation of such front-end code however could be possible through the integration of Clang compiler optimisations at the runtime of our tool and through related research such as by \citet{Campos2015} for integration into our tool for better performance.
+
     The impact this has on our tool is that the extent of our analysis is not as complete as we would like it to be, however \autoref{lst:tinyxml-refactor-example-refactored} and \autoref{fig:rewriting-example2} show that we can still achieve desirable results through the analysis of local variables.
@@
 
-    However, We do not think this detracts from the value or usefulness of our tool.
-    Our tool provides fully rejuvenated and compilable programs without the need for any human input simpler input programs. 
+    However, we do not think this detracts from the value or usefulness of our tool.
+    Our tool provides fully rejuvenated and compilable programs without the need for any human input for simpler input programs. 
     These programs are modernised to use features such as smart pointers, which also bring stricter type checking.
-    For larger, more complex, older input programs some human input is currently still needed, however the work of our tool still allows the programmer to make their programs more modern and safer by performing the complex analysis and guiding the programmer towards the needed fixes through compiler errors.
-    This results in the rejuvenated output being obtained easier by removing the need for the programmer to manually understand the structure and style of the program and the often complex ownership relations between the various variables in the program.
+    For larger, more complex and older input programs some human input is currently still needed, however the work of our tool allows the programmer to make their programs more modern and safer by performing the complex analysis and guiding the programmer towards the needed fixes through compiler errors.
+    This results in the rejuvenated output being obtained more easily by removing the need for the programmer to manually understand the structure and style of the program and the often complex ownership relations between the various variables in the program.
 

### 2021-04-14 paper.tex
Update paper.tex

@@
     We exactly aim to replace outdated coding patterns with newer, higher-level abstractions, as well as fitting the notion that this tool would be applied once to a codebase, rather than be a reoccurring task.
-    We will however use the terms rejuvenating and refactoring interchangeably in this paper as refactoring is a term that most developers are more familiar with.
+    We will however use the terms rejuvenating and refactoring interchangeably in this paper as refactoring is a term that most developers are more familiar with to convey code changes in order to improve the overall quality.
 
@@
     They use Clang for the parsing of code into a syntax tree and traverse the tree to identify and perform simple refactorings.
-    Their flexible design allows them to implement many different simple refactorings using the same tool, rather than the focus we have for a single specific refactoring goal.
+    Their flexible design allows them to implement many different simple refactorings using the same tool, rather than the focus we have for a single more complex refactoring goal.
 
@@
     The former also uses the Clang driven approach like \citet{Wright2013}, while the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
-    We have not evaluated IPR as a representation compared to the use of Clang, but we can  comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation for Clang, the ongoing support for new C++ standards and non-standard features as well as being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
+    We have not evaluated IPR as a representation compared to the use of Clang, but we can  comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation for Clang, the continuous support for new C++ standards and non-standard features as well as being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
 
@@
         \centering
-        \includegraphics[scale=0.55]{images/pointers.png}
+        \includegraphics[scale=0.45]{images/pointers.pdf}
         \caption{First graph shows the fraction of non-escaped, and therefore refactored, pointers for each project.
@@
         \centering
-        \includegraphics[scale=0.8]{images/lines.png}
+        \includegraphics[scale=0.8]{images/lines.pdf}
         \caption{Graph showing the fraction of modified lines by the tool compared to the total number of lines in each project, which is the combination of implementation and header files.
@@
         \centering
-        \includegraphics[scale=0.55]{images/time.png}
+        \includegraphics[scale=0.475]{images/time.pdf}
         \caption{Graphs of the tool runtime versus the total lines of code in each project and the total number of pointers that was analysed.

### 2021-04-15 paper.tex
Update paper.tex

@@
     The former also uses the Clang driven approach like \citet{Wright2013}, while the latter uses a representation called IPR~\cite{Dos2011}, which is a general and efficient data structure for representing C++ programs.
-    We have not evaluated IPR as a representation compared to the use of Clang, but we can  comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation for Clang, the continuous support for new C++ standards and non-standard features as well as being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
+    We have not evaluated IPR as a representation compared to the use of Clang, but we can  comment that whilst we appreciate the aim of the project to create a efficient and compiler-independent representation, the greater amount of documentation for Clang, the continuous support for new C++ standards and non-standard language extensions as well as being the choice of several other successful projects leads us to believe that it may be the more suitable and mature option for any such research.
 
@@
 
-    The first type of smart pointer is a \emph{unique pointer}, which has only a single unique owner at any time and therefore are ideal for resources which may not be copied.
-    The resource managed by a unique pointer is released when execution of the program exits the scope of the function or method that the owner of the resource is contained in.
+    The first type of smart pointer is a \emph{unique pointer}, which has only a single unique owner at any time and therefore is ideal for resources which may not be copied.
+    The resource managed by a unique pointer is released when execution of the program exits the scope that the owner of the resource is contained in.
     The ownership of a unique pointer can be transferred between variables using \texttt{std::move} which \emph{moves} the ownership from one variable to another, making the original owner invalid.
@@
     Any improvement in back-end machine code generation, as well as any new or improved IR optimisations therefore benefit all of the language front-ends.
-    It can be seen that this allows for a very modular and powerful design, shown in \autoref{fig:llvm-design}.
+    It can be seen that this allows for a very modular and powerful design, as shown in \autoref{fig:llvm-design}.
 
@@
         char *test() {
-            char* a = new char[2];
-            a[0] = 'a';
-            printf("\%s", a);
+            char* a = new char('a');
+            printf("\%c", *a);
             return a;
@@
         \begin{minted}{c++}
-        std::shared_ptr<char[]> test() {
-            std::shared_ptr<char[]> a = std::shared_ptr<char[]>(new char[2]);
-            a.get()[0] = 'a';
-            printf("\%s", a.get());
+        std::shared_ptr<char> test() {
+            std::shared_ptr<char> a = std::make_shared<char>('a');
+            printf("\%c", *a.get());
             return a;
@@
         \end{minted}
-        \caption{C++ and equivalent LLVM IR snippet showing how function arguments are not considered \texttt{noescape} or \texttt{noalias}}
+        \caption{C++ and equivalent LLVM IR snippet showing how the function argument \texttt{p} is not considered \texttt{noescape} or \texttt{noalias} by not having the correct attributes.}
         \label{fig:function-arguments}
@@
 
-    Not a single function parameter was refactored to use smart pointers, even though parameters are included in the analysis of the program, since the \texttt{VarDecl Matcher} also matches on arguments (\texttt{ParmVarDecl}) as it is a subclass~\footnote{\url{https://clang.llvm.org/doxygen/classclang_1_1ParmVarDecl.html}}.
+    Not a single function parameter was identified to be non-escaping, as parameters are included in the analysis of the program, since the \texttt{VarDecl Matcher} also matches on arguments (\texttt{ParmVarDecl}) as it is a subclass~\footnote{\url{https://clang.llvm.org/doxygen/classclang_1_1ParmVarDecl.html}}.
     Upon investigation, this is largely an issue due memory coherence guarantees in the specification of the language~\cite[p.~66]{C++17}. 
     Function pointer parameters are assumed by default to escape and to possibly be aliased (\textit{the memory could possibly be accessed through a different variable}).
-    Therefore by default no parameters will ever be identified by our analysis as non-escaping and will not be refactored.
+    Therefore by default no parameters will ever be identified by our analysis as non-escaping.
     We can see this in \autoref{fig:function-arguments}, where the variable \texttt{p} in the generated LLVM IR is not marked as \texttt{noescape} and \texttt{noalias}~\footnote{\url{https://clang.llvm.org/docs/AttributeReference.html\#noalias}}.
